---
title: "Chapter 09: Poisson GLMMs"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 09: Poisson GLMMs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

```{r load-stored}
ch09 <- readRDS(system.file(
  "extdata", "Chapter09_airbnb_poisson.rds", package = "lmebayes"
))
```

# Chapter 09: Poisson GLMMs

Everything so far has been Gaussian. This chapter switches to counts, which
changes the sampler substantially even though the interface barely moves:
replace `lmerb()` with `glmerb()` and add `family = poisson()`.

The data is `airbnb` from **bayesrules** [@JohnsonLauEtAl2022]: listings in
Chicago neighborhoods, with the number of reviews as the count outcome. The
fit takes about half a minute because every Block 1 draw runs an envelope
accept-reject sampler, so it is precomputed by
`data-raw/make_vignette_outputs.R`.

# What changes for non-Gaussian responses

| | Gaussian | Poisson |
|---|---|---|
| Block 1 draw | Conjugate normal, closed form | Envelope accept-reject |
| Joint posterior | Multivariate normal (when \(\tau^2\) fixed) | Never normal |
| Exact iid route | Available | Not available |
| Convergence rate | Certified by theorem | Local-Gaussian heuristic |
| Pilot stage | Only when \(\tau^2\) sampled | Always |

The last three rows matter. For a Gaussian target the rate \(\lambda^*\)
of Chapter 06 is a theorem; for a Poisson likelihood it is evaluated at
IRLS-style curvature weights taken at the posterior mode, which makes it a
good practical guide rather than a certificate. The engine compensates by
always running a pilot stage. The distinction is spelled out in
`vignette("Chapter-B02", package = "lmebayesCore")`, under "Beyond the
Gaussian case".

Log-concavity is what makes the Block 1 draw possible at all. The Poisson
log-likelihood with a log link is log-concave in the linear predictor, so
the likelihood-subgradient envelope of @Nygren2006 is a valid bound and
accept-reject sampling is exact — not a Metropolis approximation.

# The model

```{r data-code, eval = FALSE}
data(airbnb, package = "bayesrules")
ab <- airbnb[, c("neighborhood", "reviews", "rating", "room_type")]
ab <- ab[complete.cases(ab), ]
ab$neighborhood <- droplevels(factor(ab$neighborhood))
ab$rating_c <- ab$rating - mean(ab$rating)

form <- reviews ~ rating_c + (1 + rating_c || neighborhood)
model_setup(form, data = ab, family = poisson())
```

```{r design-out, echo = FALSE}
cat(ch09$design, sep = "\n")
```

`r ch09$n_obs` listings across `r ch09$n_groups` neighborhoods. Each
neighborhood gets its own baseline review rate and its own rating effect.
Rating is centered so the intercept means "log expected reviews for an
average-rated listing".

Eight neighborhoods are rank-deficient. For a count model that usually
means too few listings, or no variation in rating among them. They stay in
the fit and still receive coefficients — with nothing to learn from in the
deficient direction, they are shrunk to the population value — but they are
set aside when the level-2 rank condition is checked and when priors are
calibrated.

# Priors and fitting

```{r fit-code, eval = FALSE}
ps <- Prior_Setup_GLMM(form, data = ab, family = poisson(), pop.pwt = 0.01)

set.seed(2024)
fit <- glmerb(
  form,
  data         = ab,
  family       = poisson(),
  pfamily_list = pfamily_list(ps),
  n            = 200L,
  progbar      = FALSE
)
fit
```

```{r fit-out, echo = FALSE}
cat(ch09$fit_print, sep = "\n")
```

Two interface differences from `lmerb()`. There is no `dispersion_ranef`
argument — the Poisson family fixes the dispersion at one, so there is no
\(\sigma^2\) to specify. And the reference fit is `glmer()` rather than
`lmer()`, which is what `Prior_Setup_GLMM()` calibrates against.

The column header reads `ICM mode` rather than `fixef.mode`: for a
non-Gaussian likelihood the pre-sampling point estimate is the posterior
*mode*, found by iterated conditional modes, and the mode is not the mean.
That is precisely why a pilot stage is needed — a chain started at the mode
begins at an unknown positive distance from the target mean, so the
mean-error term of the convergence bound is no longer zero.

# Interpreting on the count scale

Coefficients are on the log scale, so exponentiate to get rate ratios.

```{r interpret}
pm <- unlist(ch09$popef_means)
round(rbind(log_scale = pm, rate_ratio = exp(pm)), 3)
```

The intercept exponentiates to the expected number of reviews for an
average-rated listing in an average neighborhood — about
`r round(exp(unlist(ch09$popef_means)[1]), 0)`. The `rating_c` coefficient
is a multiplicative effect per unit of rating: each additional rating point
multiplies expected reviews by roughly
`r round(exp(unlist(ch09$popef_means)[2]), 2)`.

But look at the posterior standard deviation on that slope — it is nearly
as large as the coefficient. The population-average rating effect is not
well determined, and the summary explains why.

# The full summary

```{r summary-code, eval = FALSE}
summary(fit)
```

```{r summary-out, echo = FALSE}
cat(ch09$fit_summary, sep = "\n")
```

The random-effects standard deviation for `rating_c` is larger than the
population mean effect itself. Neighborhoods genuinely differ in how rating
relates to review volume — in some, better-rated listings get many more
reviews; in others the relationship is flat or reversed. Reporting only the
population coefficient would describe none of them well. This is the case
for a random slope rather than a fixed one, and it is visible directly in
the variance components.

There is no `Residual` row in the dispersion table. Poisson has no free
dispersion parameter; the variance is determined by the mean. If your
counts are overdispersed relative to that, the random intercept absorbs
some of it, but a genuine overdispersion model is outside this package's
model class.

# Convergence

```{r convergence}
c(m_convergence = ch09$m_convergence)
```

`r ch09$m_convergence` sweeps per stored draw, and a pilot stage ran before
the main chains. Because the rate is heuristic here rather than certified,
the sensible check is empirical: the sweep history should show the
cross-chain mean flat and the cross-chain variance settled by the last
sweep, exactly as in Chapter 06. `fit$sweep_history$main` carries it.

# Cost

The Gaussian fits in earlier chapters took a second or two. This one takes
about thirty. The difference is entirely Block 1: every group, on every
sweep, on every chain, runs an envelope construction and an accept-reject
loop. `fit$groupef.iters` reports how many candidates were needed per
accepted draw — around eight here, which is normal for a Poisson model and
means the envelope is doing real work.

Practical consequences: develop with small `n` and raise it for the final
run; prefer `tv_tol` at its default rather than tightening it (each extra
sweep is a full pass over all groups and chains); and remember that
`glmbayes` `Chapter-A10` documents optional OpenCL offload for envelope
construction, which this package inherits.

# See also

- **Chapter 10** — the binomial case, where the same machinery applies.
- **Chapter 11** — `simulate = FALSE`, when you only want the mode.
- `vignette("Chapter-A05", package = "glmbayes")` — likelihood subgradient
  densities, the basis of the Block 1 draw.
- `?glmerb`
