---
title: "Chapter 05: Reading the output"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 05: Reading the output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

# Chapter 05: Reading the output

An `lmerb` fit is a list with about thirty elements. This chapter is the
map: which accessor gives which quantity, what the raw containers hold, and
where the naming departs from `lme4` in ways that will surprise you.

```{r fit, results = "hide"}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)
form <- Reaction ~ Days_c + (1 + Days_c || Subject)

ps <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01)

set.seed(1)
fit <- lmerb(
  form, data = dat,
  pfamily_list     = pfamily_list(ps),
  dispersion_ranef = ps$group.dispersion,
  n                = 500L,
  progbar          = FALSE
)
```

# Two blocks, two prefixes

Everything in the output is named after which block it belongs to.

| Prefix | Block | Contents |
|---|---|---|
| `popef.*` | Block 2 | Population coefficients \(\gamma\) and their variance components |
| `groupef.*` | Block 1 | Per-group coefficients \(\beta_j\) |
| `group.dispersion*` | — | Observation dispersion \(\sigma^2\) |

Within each prefix the suffix says what kind of quantity it is.

| Suffix | Meaning |
|---|---|
| none | The raw draws |
| `.mode` | Point estimate before sampling (exact posterior mean for Gaussian, mode otherwise) |
| `.means` / `.mean` | Average over the stored draws |
| `.iters` | Accept-reject candidates per draw (sampler efficiency) |

```{r names}
names(fit)
```

# `print()`

```{r print}
fit
```

The header line carries three facts: how many draws, and which engine
produced them. `exact iid` means the joint posterior was multivariate
normal and was sampled directly. `two-block Gibbs` means sweeps were
needed; Chapter 06 covers that case.

The table's three columns are worth checking every time. `fixef.mode` is
computed analytically before sampling; `fixef.means` is the Monte Carlo
average. They should agree to within `draws SD / sqrt(n)`. If they do not,
something is wrong with the sampler, not with your model.

# `summary()`

```{r summary}
summary(fit)
```

Three sections:

1. **Block 2 dispersion** — the variance components \(\tau^2_k\), and the
   residual variance, each against its `lmer` reference. When a component
   is `dNormal` its percentiles are constant, because it is fixed rather
   than sampled.
2. **Block 2 level-2 fixed effects** — the population parameters, with
   prior means, `lmer` reference values, posterior SD, Monte Carlo error,
   and `Pr(Prior_tail)`.
3. **Block 1 random effects** — a distributional summary across groups
   rather than a per-group listing.

`summary()` takes a `groups` argument when you want specific levels:

```{r summary-groups, eval = FALSE}
summary(fit, groups = c("308", "309"))
```

## `Pr(Prior_tail)`

This is a directional tail probability, not a p-value. It measures how far
the posterior has moved into the tail of the prior, in the direction of the
shift. Near 0.5 means the data barely moved that coefficient away from its
prior mean. Near zero means the posterior sits far out in the prior's tail
— the data disagrees with the prior, decisively. The theory is in
`vignette("Chapter-A04", package = "glmbayes")`.

# Accessors

`fixef()`, `ranef()`, and `coef()` are `lme4` generics. **lmebayes** defines
methods for them but does not re-export the generics, so call them from
**lme4**.

## `ranef()`

```{r ranef}
re <- lme4::ranef(fit)
head(re$Subject)
```

**These are not `lme4` random effects.** They are the full per-subject
coefficients \(\beta_j\) — subject 308's intercept is their expected
reaction time, not an offset from the population mean. See Chapter 02.
The `lme4` quantity is \(\beta_j\) minus the population mean:

```{r ranef-deviations}
head(re$Subject) - rep(unlist(fit$popef.means), each = 6)
```

Two arguments matter. `type` chooses between the chain mean (default) and
the ICM point estimate, and `condVar` attaches posterior variances:

```{r ranef-args}
re_mode <- lme4::ranef(fit, type = "mode")
head(re_mode$Subject, 3)

re_cv <- lme4::ranef(fit, condVar = TRUE)
head(as.data.frame(re_cv), 4)
```

The long format has `condsd` filled in when `condVar = TRUE`, which is what
you want for a caterpillar plot.

## `coef()`

```{r coef}
head(coef(fit)$Subject)
```

For `merMod` objects `coef()` adds `fixef` to `ranef`. Here `ranef()`
already returns full coefficients, so `coef()` returns the same values —
kept for interface compatibility rather than because it computes anything
new.

## `fixef()`

```{r fixef}
lme4::fixef(fit)                        # type = "mer": the reference lmer fit
lme4::fixef(fit, type = "hyper")        # Block 2 posterior mode
lme4::fixef(fit, type = "hyper.mean")   # Block 2 posterior mean
```

The default is the `lmer` reference so that `fixef()` on this object
behaves the way it does on a `merMod`. The Bayesian quantities are behind
`type`.

Names differ between the two: the `lmer` version uses model-matrix names
(`Days_c`), while the hyper versions use `component.parameter`
(`Days_c.(Intercept)`), reflecting that each population parameter belongs
to a specific random coefficient's level-2 row.

# The raw draws

Accessors summarize. When you want the draws themselves:

```{r popef-draws}
str(fit$popef)
```

`fit$popef` is a list with one matrix per random coefficient: rows are
draws, columns are that coefficient's level-2 parameters. Because the draws
are independent, ordinary Monte Carlo standard errors apply directly:

```{r mcse}
draws_int <- fit$popef[["(Intercept)"]][, 1]
c(mean = mean(draws_int),
  sd   = sd(draws_int),
  mcse = sd(draws_int) / sqrt(length(draws_int)))
```

No effective-sample-size correction, no thinning. That is the payoff of the
replicate-chain design described in
`vignette("Chapter-B01", package = "lmebayesCore")`.

```{r groupef-draws}
head(fit$groupef)
nrow(fit$groupef)
```

`fit$groupef` is long format: `n` draws times `J` groups. Filter it
directly, or use the printer:

```{r print-groupef}
lmebayesCore::print_groupef(fit, draws = 1:3, groups = 1:3)
```

`print_groupef()` lives in **lmebayesCore** and is not re-exported here, so
it needs the namespace prefix. It takes the fit, not the data frame.

# Credible intervals

Nothing special is required — these are ordinary independent draws.

```{r ci}
t(apply(fit$popef[["(Intercept)"]], 2, quantile,
        probs = c(0.025, 0.5, 0.975)))
t(apply(fit$popef[["Days_c"]], 2, quantile,
        probs = c(0.025, 0.5, 0.975)))
```

Per-group intervals come from the long-format container:

```{r ci-groups}
by_subject <- split(fit$groupef$Days_c, fit$groupef$Subject)
head(t(sapply(by_subject, quantile, probs = c(0.025, 0.5, 0.975))), 5)
```

# Dispersion and convergence metadata

```{r dispersion}
fit$group.dispersion       # sigma^2 (fixed here)
fit$m_convergence          # inner Gibbs sweeps per stored draw
fit$sim_method_used        # which route ran
```

`summary_sigma2()` gives a full report, but only when the observation
dispersion is per-group — with a single fixed `sigma^2` there is nothing to
report and it stops with a message. Chapter 08 covers that case.

# Quick reference

| You want | Use |
|---|---|
| Population posterior mean | `fit$popef.means` or `fixef(fit, "hyper.mean")` |
| Population draws | `fit$popef[[k]]` |
| Per-group coefficients | `lme4::ranef(fit)` or `fit$groupef.mode` |
| Per-group draws | `fit$groupef` (long format) |
| Observation variance | `fit$group.dispersion` |
| Variance components | `summary(fit)`, Block 2 dispersion section |
| Sweeps per draw | `fit$m_convergence` |
| Reference `lmer` fit | `fit$lmer` |
| The parsed design | `fit$model_setup` |

# See also

- **Chapter 06** — `m_convergence` and the sweep history.
- **Chapter 08** — `summary_sigma2()` and per-group dispersion.
- `?lmerb`, `?ranef.lmerb`, `?fixef.lmerb`
