---
title: "Chapter 11: Point estimates without simulation"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 11: Point estimates without simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

# Chapter 11: Point estimates without simulation

Both `lmerb()` and `glmerb()` accept `simulate = FALSE`. The fit then stops
after computing a point estimate and returns without drawing anything.

This is not a degraded mode. For Gaussian models with fixed variance
components, the point estimate is the **exact posterior mean**, computed in
closed form. It is also the starting point every replicate chain uses when
you do simulate, which is why it is worth understanding on its own.

```{r data}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)
form <- Reaction ~ Days_c + (1 + Days_c || Subject)
ps <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01)
```

# Iterated conditional modes

The algorithm is simple. The two blocks each have a closed-form
conditional mode, so alternate:

```
repeat:
  beta_j  <- mode of  p(beta_j | y_j, gamma, Psi)   for every group j
  gamma   <- mode of  p(gamma | beta_1..J, priors)
until nothing moves
```

That is **iterated conditional modes** (ICM) — coordinate ascent on the
joint posterior, one block at a time. For a Gaussian model with fixed
\(\tau^2_k\) and \(\sigma^2\), the joint posterior is multivariate normal,
so its mode is also its mean, and each conditional mode is a linear solve.
The iteration is then exactly block Gauss–Seidel on the joint precision
system: it converges to the unique stationary point, and it does so
quickly.

```{r fit-icm, results = "hide"}
pt <- lmerb(form, data = dat,
            pfamily_list     = pfamily_list(ps),
            dispersion_ranef = ps$group.dispersion,
            simulate         = FALSE)
```

```{r fit-icm-print}
pt
```

```{r icm-info}
pt$convergence$icm_info
```

One iteration, delta zero. With no group-level covariates the two blocks
decouple after a single pass, so the answer is reached immediately. Models
with level-2 predictors take a few more.

# It really is the posterior mean

```{r verify, results = "hide"}
set.seed(1)
fit <- lmerb(form, data = dat,
             pfamily_list     = pfamily_list(ps),
             dispersion_ranef = ps$group.dispersion,
             n = 2000L, progbar = FALSE)
```

```{r verify-table}
n <- nrow(fit$popef[["Days_c"]])
tab <- data.frame(
  icm       = unlist(pt$popef.mode),
  draws     = unlist(fit$popef.means),
  mcse      = vapply(fit$popef, function(m) sd(m[, 1]) / sqrt(n), numeric(1))
)
tab$z <- (tab$draws - tab$icm) / tab$mcse
round(tab, 4)
```

The draws' average agrees with the ICM value to within Monte Carlo error.
This is the package's own regression test, and it is worth running whenever
you change sampler settings: if the two disagree by more than a few
standard errors, the sampler is not targeting the posterior you think it
is.

# What you get and what you do not

```{r contents}
setdiff(names(fit), names(pt))
```

```{r contents2}
c(has_groupef_mode = !is.null(pt$groupef.mode),
  has_groupef_draws = !is.null(pt$groupef))
```

Point estimates for both blocks are present; the draw containers are not.
So `ranef(pt)` works — falling back to the mode with a warning — but
anything requiring a distribution does not:

```{r accessors}
head(lme4::ranef(pt, type = "mode")$Subject, 3)
```

No credible intervals, no posterior standard deviations, no tail
probabilities. If you need uncertainty, you need draws.

# The non-Gaussian case

For Poisson, binomial, and Gamma responses, ICM still converges — the
log-likelihoods are log-concave, so each conditional has a unique mode —
but it converges to the **posterior mode**, and the mode is not the mean.
The printed column header changes from `fixef.mode` to `ICM mode` to say so.

That gap is why non-Gaussian fits always run a pilot stage (Chapter 09):
a chain started at the mode begins at an unknown positive distance from the
target mean, so the mean-error term of the convergence bound is no longer
zero and has to be bounded empirically instead.

For a skewed posterior the mode can be a poor summary. Treat
`simulate = FALSE` on a GLMM as a fast diagnostic, not a substitute for
inference.

# When to use it

**Model development.** Checking that a formula parses, that groups are
retained, and that estimates are sane costs a fraction of a second instead
of thirty.

```{r dev-loop, results = "hide"}
forms <- list(
  intercept_only = Reaction ~ 1 + (1 | Subject),
  random_slope   = Reaction ~ Days_c + (1 + Days_c || Subject)
)
fits <- lapply(forms, function(f) {
  p <- Prior_Setup_GLMM(f, data = dat, pop.pwt = 0.01)
  lmerb(f, data = dat, pfamily_list = pfamily_list(p),
        dispersion_ranef = p$group.dispersion, simulate = FALSE)
})
```

```{r dev-loop-out}
lapply(fits, function(f) round(unlist(f$popef.mode), 3))
```

**Prior sensitivity sweeps.** When you only need to see how the point
estimate moves across prior weights, there is no reason to sample.

**Large problems.** Many groups, a non-Gaussian family, and a tight
tolerance can make a full fit expensive. Getting the mode first tells you
whether the model is worth sampling.

**Not for reporting.** A point estimate without a posterior is a maximum
a posteriori estimate with extra steps. The reason to use this package at
all is the certified draws.

# See also

- **Chapter 06** — why the ICM value doubles as the chain start point.
- **Chapter 09** — the pilot stage, which exists because mode is not mean.
- `?lmerb`, `?glmerb`
