---
title: "Chapter 02: From lm and glm to mixed models"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 02: From lm and glm to mixed models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

# Chapter 02: From `lm`/`glm` to mixed models

Chapter 01 fit a model without saying much about what it means. This
chapter fills that gap: what partial pooling buys you, how **lmebayes**
parameterizes a mixed model, and why the two model restrictions
(one grouping factor, uncorrelated terms) exist.

There is no sampling in this chapter.

# Three ways to handle groups

Take reaction time measured repeatedly on each of \(J\) subjects. You have
three options.

**Complete pooling.** `lm(Reaction ~ Days_c)`. One intercept, one slope,
everyone the same. Simple, and wrong whenever subjects genuinely differ:
the residuals are correlated within subject, so standard errors are too
small.

**No pooling.** `lm(Reaction ~ Days_c * Subject)`, or 18 separate
regressions. Each subject gets their own coefficients, estimated from their
own 10 observations and nothing else. Honest, but noisy — a subject with an
unlucky run of measurements gets an extreme slope, and nothing pulls it
back.

**Partial pooling.** The mixed model. Each subject still gets their own
coefficients, but those coefficients are themselves modeled as draws from a
population distribution. A subject with little or noisy data is shrunk
toward the population mean; a subject with lots of clean data is barely
moved.

The amount of shrinkage is not a tuning choice. It falls out of the ratio
of within-group information to between-group variance:

\[
\text{shrinkage for group } j \;\approx\;
\frac{1/\tau^2}{1/\tau^2 + n_j/\sigma^2}.
\]

Small groups (small \(n_j\)) and tight population distributions (small
\(\tau^2\)) shrink more. That expression is not an analogy: it is
*literally* the convergence rate \(\lambda^*\) of the two-block Gibbs
sampler for the balanced random-intercept model, derived in
`vignette("Chapter-B02", package = "lmebayesCore")`. Models that shrink
hard are exactly the models whose sampler mixes slowly, because in both
cases the group coefficients and the population mean are tightly coupled.

# The parameterization

Here is where **lmebayes** differs from `lme4` in a way you will notice in
the output.

**`lme4` writes** the model as a fixed part plus a mean-zero deviation:

\[
y_j = X_j\beta + Z_j b_j + \varepsilon_j,
\qquad b_j \sim N(0, G).
\]

`fixef()` gives \(\beta\), `ranef()` gives \(b_j\) (centered at zero), and
`coef()` gives \(\beta + b_j\).

**lmebayes writes** the same model in centered, two-stage form
[@LindleySmith1972]:

\[
\begin{aligned}
\text{Stage 1:}\quad & y_j \mid \beta_j \sim N(D_j\beta_j,\; \sigma^2 I) \\
\text{Stage 2:}\quad & \beta_j \mid \gamma \sim N(\mathcal{W}_j\gamma,\; \Psi)
\end{aligned}
\]

The group-level coefficient vector \(\beta_j\) is *not* mean zero. It is
the full coefficient vector for group \(j\), and it enters the likelihood
directly. The population coefficients \(\gamma\) appear only in stage 2, as
the mean of the \(\beta_j\).

Substituting stage 2 into stage 1 recovers `lme4`'s form exactly:

\[
y_j = D_j(\mathcal{W}_j\gamma + u_j) + \varepsilon_j
    = \underbrace{(D_j\mathcal{W}_j)}_{X_j}\gamma
    + \underbrace{D_j}_{Z_j} u_j + \varepsilon_j,
\qquad u_j := \beta_j - \mathcal{W}_j\gamma .
\]

So the two are the same model, written differently.

| lmebayes | `lme4` | Note |
|---|---|---|
| \(\gamma\) | \(\beta\) | population / fixed effects |
| \(\beta_j\) | \(\hat\beta + \hat b_j\) | what `lme4::coef()` returns |
| \(u_j = \beta_j - \mathcal{W}_j\gamma\) | \(b_j\) | mean-zero deviation |
| \(D_j\) | \(Z_j\) | random-effects design |
| \(D_j\mathcal{W}_j\) | \(X_j\) | fixed-effects design |
| \(\Psi\) | per-group block of \(G\) | \(G = I_J \otimes \Psi\) |

**Why it matters in practice.** `ranef()` on an `lmerb` fit returns
\(\beta_j\), not \(u_j\). In Chapter 01, subject 308's intercept was about
339 ms — an actual expected reaction time, not a `+40` offset from the
population mean. If you are used to reading `ranef()` output as deviations
that should average to zero, this will look wrong until you recognize the
scale.

**Why the package does it this way.** The centered form is what makes the
two-block split natural. Given \(\gamma\), the \(\beta_j\) are
conditionally independent across groups, so Block 1 is \(J\) small
independent draws. Given all \(\beta_j\), Block 2 is a plain regression of
the \(\beta\)'s on the group-level designs. Neither block ever forms a
matrix bigger than it has to, and the convergence theory of @Nygren2020
applies directly to this two-block structure.

# The level-2 design

\(\mathcal{W}_j\) is the piece with no `lme4` counterpart. It maps the
population vector \(\gamma\) to group \(j\)'s prior mean.

In the simplest case — no group-level covariates — \(\mathcal{W}_j\) is an
identity-like map and every group has the same prior mean. That is the
`sleepstudy` model of Chapter 01, where `model_setup()` reported both
level-2 rows as `~ 1`.

Add a group-level covariate and it becomes interesting. If schools have a
`title1` indicator, the level-2 row for the intercept reads
`~ 1 + title1`, and each school's *prior mean intercept* depends on whether
it is a Title I school. Shrinkage is then toward a fitted line rather than
a single grand mean. Chapter 04 does exactly this.

Crucially, each random coefficient may carry its own level-2 predictors:

```
(Intercept)     ~ 1 + private_school + title1
distracted_ppvt ~ 1
distracted_a1   ~ 1 + free_reduced_lunch
```

An interaction between a group-level variable and an individual-level
random slope — `free_reduced_lunch:distracted_a1` — is *cross-level
moderation*, and it lands in the level-2 row of `distracted_a1`.
`model_setup()` prints this mapping, and it is worth reading before every
fit: it is the clearest statement of what model you actually specified.

# The two restrictions

## One grouping factor

`(1 | school) + (1 | teacher)` is not supported. Neither is nesting.

With a single grouping factor the parameter vector splits cleanly into
"the group coefficients" and "everything else", and the group coefficients
are conditionally independent given the population level. Crossed factors
destroy that: a given observation belongs to two groups at once, so the
Block 1 conditional no longer factorizes and the rate matrix of
`Chapter-B02` is no longer block-diagonal by group.

## Uncorrelated random effects: `||` not `|`

`(1 + Days_c | Subject)` estimates a correlation between each subject's
intercept and slope. `lmebayes` requires `(1 + Days_c || Subject)`, which
forces that correlation to zero, so \(\Psi\) is diagonal.

Two things follow from a diagonal \(\Psi\). Block 2 factorizes across
coefficients: each \(\gamma_k\) is drawn from its own small regression, and
each variance component \(\tau^2_k\) is handled on its own. That is why
priors are specified as a `pfamily_list` — one object per coefficient —
rather than one joint prior over all of \(\gamma\).

If you pass a `|` formula, `model_setup()` stops with an explicit message
rather than silently ignoring the correlation:

```{r bad-formula, error = TRUE}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)
model_setup(Reaction ~ Days_c + (1 + Days_c | Subject), data = dat)
```

**Does the restriction hurt?** Sometimes. If intercept and slope really are
correlated, forcing independence changes the shrinkage pattern: a subject
who is slow at baseline will not be pulled toward also having a steep
slope. Centering the covariate — as Chapter 01 did with `Days_c` — usually
removes most of the correlation, because the intercept-slope correlation in
an uncentered model is largely an artifact of where zero sits. Fit both in
`lme4`, compare the estimated correlation, and if it is substantial after
centering, this package is not the right tool. Chapter 13 works through
such a comparison.

## Every fixed effect must have a home

There is a third restriction, and it is the one that catches people first.
A fixed effect is only allowed if it can be placed in some group's level-2
row. Concretely, each fixed-effect term must be one of:

- **constant within group** — a group-level covariate, which goes on the
  intercept's level-2 row;
- **the population mean of a random slope** — its name must match a term in
  the random-effects part;
- **a cross-level moderator** — `moderator:random_slope`, where the
  moderator is constant within group.

So `Reaction ~ Days_c + (1 + Days_c || Subject)` is fine, because `Days_c`
has a matching random slope. But `Reaction ~ Days_c + (1 | Subject)` is
not:

```{r fixed-effect-rule, error = TRUE}
model_setup(Reaction ~ Days_c + (1 | Subject), data = dat)
```

This is not arbitrary. In the centered parameterization there is no place
to put a coefficient on an individual-level covariate that does not vary by
group: stage 1 only contains \(D_j\beta_j\), and stage 2 only contains
\(\mathcal{W}_j\gamma\). An individual-level predictor with a *common*
slope across groups has no slot in either. `lme4` can express it because it
writes \(X_j\beta\) as a free-standing term; this package cannot, and stops
rather than silently reinterpreting your formula.

The practical consequence: **if you want a fixed effect for an
individual-level covariate, give it a random slope**. If its variance
component is genuinely near zero, the fit will say so.

## Log-concave likelihoods

Gaussian, Poisson, binomial, and Gamma with the standard links. The
restriction comes from the envelope accept-reject sampler
[@Nygren2006]: it builds a tangent-plane envelope over the log-likelihood,
which requires log-concavity to be a valid bound. Chapters 09 and 10 cover
the non-Gaussian cases.

# See also

- **Chapter 03** — the prior workflow.
- **Chapter 04** — group-level covariates and cross-level moderation.
- `vignette("Chapter-B00", package = "lmebayesCore")` — the same notation,
  stated formally.
- `vignette("Chapter-B02", package = "lmebayesCore")` — where the shrinkage
  ratio reappears as a convergence rate.
