---
title: "Chapter 07: Estimated variance components and the pilot stage"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 07: Estimated variance components and the pilot stage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

# Chapter 07: Estimated variance components and the pilot stage

Every fit so far has held the variance components \(\tau^2_k\) fixed at the
`lmer` estimates. That is a real assumption: it says you know exactly how
much subjects vary, and it makes the posterior for everything else
narrower than it should be. This chapter relaxes it.

```{r data}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)
form <- Reaction ~ Days_c + (1 + Days_c || Subject)
```

# Fixed versus sampled

The choice is made per random coefficient, through the prior family.

| Family | \(\tau^2_k\) | Consequence |
|---|---|---|
| `dNormal` | Fixed at the reference estimate | Joint posterior stays Gaussian; exact iid draws |
| `dIndependent_Normal_Gamma` | Sampled, Gamma prior on \(1/\tau^2_k\) | Target no longer Gaussian; Gibbs sweeps and a pilot stage |

Calibrate as usual, adding a prior weight for the dispersion:

```{r prior}
ps <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01,
                       pop.dispersion.pwt = 0.2)
ps$pop.ing_prior
```

`pop.dispersion.pwt = 0.2` gives the dispersion prior the weight of about
20 percent of the groups. With only 18 subjects there is genuinely little
information about a variance component, so this is not a small choice —
compare it against `0.05` and `0.5` before trusting the width of your
intervals.

# The truncation window

Each ING component carries `disp_lower` and `disp_upper`, and the sampled
\(\tau^2_k\) is confined to that interval.

The bounds are quantiles of the *limiting posterior* of the variance
component, not of the prior. That distinction matters: a Gamma prior on a
precision has a fat tail toward zero variance, and a \(\tau^2\) driven to
zero collapses the hierarchy — every group's coefficient is pinned to the
population mean, and the sampler's coupling becomes total. Truncating away
that region is what keeps the chain geometrically ergodic.

The lower bound does double duty. The sampler evaluates its convergence
rate at `disp_lower`, because smaller \(\tau^2\) means tighter coupling and
a slower chain, so a bound computed there is valid across the entire
window. That is conservative rather than correct-on-average, which is the
right direction for a certificate. See
`vignette("Chapter-B05", package = "lmebayesCore")` and
`inst/ING_TRUNCATION_WINDOW.md` in the package sources.

# Fitting

```{r pfamily}
pf_ing <- pfamily_list(ps, ptypes = "dIndependent_Normal_Gamma")
class(pf_ing[["Days_c"]])
```

`ptypes` also accepts a named vector when you want to sample some
components and fix others — for example
`ptypes = c("(Intercept)" = "dNormal", "Days_c" = "dIndependent_Normal_Gamma")`.

```{r fit, results = "hide"}
set.seed(1)
fit <- lmerb(form, data = dat,
             pfamily_list     = pf_ing,
             dispersion_ranef = ps$group.dispersion,
             n = 300L, progbar = FALSE)
```

```{r fit-print}
fit
```

Two things changed in the header. The engine is now `two-block Gibbs`
rather than `exact iid`, and the random-effects line says the variance
components are sampled rather than fixed.

# What the variance components look like

```{r tau2-draws}
head(fit$popef.dispersion)
round(fit$popef.dispersion.mean, 3)
```

`fit$popef.dispersion` holds the \(\tau^2_k\) draws — one column per random
coefficient. Now they have a distribution:

```{r tau2-quantiles}
round(t(apply(fit$popef.dispersion, 2, quantile,
              probs = c(0.025, 0.5, 0.975))), 2)
```

On the standard-deviation scale, which is easier to read against `lmer`:

```{r tau2-sd}
round(t(apply(sqrt(fit$popef.dispersion), 2, quantile,
              probs = c(0.025, 0.5, 0.975))), 3)
```

Compare against the `lmer` point estimates:

```{r lmer-sd}
print(lme4::VarCorr(fit$lmer), comp = "Std.Dev.")
```

The `lmer` numbers should sit inside the credible intervals. They are point
estimates with no uncertainty attached; the intervals here are wide, and
honestly so — 18 subjects is not much information about how subjects vary.

# The cost of honesty

```{r cost, results = "hide"}
set.seed(1)
fit_fixed <- lmerb(form, data = dat,
                   pfamily_list     = pfamily_list(ps),
                   dispersion_ranef = ps$group.dispersion,
                   n = 300L, progbar = FALSE)
```

```{r cost-table}
sd_fixed <- vapply(fit_fixed$popef, function(m) sd(m[, 1]), numeric(1))
sd_ing   <- vapply(fit$popef,       function(m) sd(m[, 1]), numeric(1))
round(data.frame(
  fixed_tau2   = sd_fixed,
  sampled_tau2 = sd_ing,
  ratio        = sd_ing / sd_fixed
), 4)
```

Posterior standard deviations on the population coefficients are wider once
the variance components are uncertain. That widening is the point: fixing
\(\tau^2\) at a point estimate and reporting the resulting intervals
understates what you know.

# The pilot stage

The convergence theory of Chapter 06 assumed chains start at the exact
posterior mean, making the mean-error term of the bound exactly zero. With
sampled variance components the joint target is no longer Gaussian, so
there is no exact mean to start from.

The engine's answer is a **pilot stage**: a batch of short chains whose
cross-chain mean becomes the start point for the main chains. The pilot
draws are discarded. Their job is not to estimate anything but to produce a
starting point whose distance from the truth can be *bounded*.

```{r pilot}
fit$pilot
```

If the pilot mean averages `n_pilot` independent chains, its sampling
distribution has covariance \(\Sigma/n_{\text{pilot}}\), so the squared
standardized start distance obeys
\(D_0 = \chi^2_{p,\alpha} / n_{\text{pilot}}\) at confidence level
\(\alpha\). Bigger pilots give smaller \(D_0\) and therefore fewer
main-stage sweeps — but the pilot itself costs sweeps. The engine minimizes
total cost over `n_pilot`; the optimization is worked through in
`vignette("Chapter-B03", package = "lmebayesCore")`.

Both stages leave a sweep history:

```{r sweep-history}
names(fit$sweep_history)
fit$sweep_history$pilot
```

```{r sweep-history-main}
fit$sweep_history$main
```

The pilot history starts at the ICM mode and moves; the main history starts
at the pilot mean and barely moves. That contrast is the pilot doing its
job.

# Reading the summary

```{r summary}
summary(fit)
```

The Block 2 dispersion section is now informative. `Post.Mode`,
`Post.Mean`, `Post.Sd`, and the percentile rows are real distributions
rather than repeated constants, `disp_lower` / `disp_upper` show the
truncation window, and `Cand/draw` reports how many accept-reject
candidates the envelope sampler needed per draw — a value near 1 means the
envelope is tight, larger values mean it is working harder.

# When to sample the variance components

**Sample them** when the number of groups is small (a variance component
from 18 subjects is barely identified), when the variance component is
itself of interest, or when you are reporting intervals you want to be
honest.

**Fix them** when there are many groups and \(\tau^2\) is well determined,
when you want exact iid draws with no sweeps, or when speed matters and the
sensitivity check shows it makes no difference.

The cheapest way to decide is to fit both, as above, and look at the ratio
of posterior standard deviations.

# See also

- **Chapter 08** — the same question for the *observation* dispersion.
- `vignette("Chapter-B05", package = "lmebayesCore")` — the ING Block 2
  derivation and the truncation window.
- `vignette("Chapter-B06", package = "lmebayesCore")` — the local rate
  diagnostic used when \(\tau^2\) is sampled.
- `?pfamily_list`, `?dIndependent_Normal_Gamma`
