---
title: "Chapter 04: Linear mixed models with lmerb()"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 04: Linear mixed models with lmerb()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

```{r load-stored}
ch04 <- readRDS(system.file(
  "extdata", "Chapter04_big_word_club.rds", package = "lmebayes"
))
```

# Chapter 04: Linear mixed models with `lmerb()`

Chapter 01 fit a model with no group-level structure: every subject shrank
toward the same grand mean. This chapter uses a dataset where the groups
themselves differ in known ways, which is where the level-2 design earns
its keep.

The data is `big_word_club` from the **bayesrules** package
[@JohnsonLauEtAl2022] — a vocabulary intervention study in
`r ch04$n_schools` schools. Because **bayesrules** is only suggested and a
1000-draw fit takes a few seconds, the model is fitted ahead of time by
`data-raw/make_vignette_outputs.R` and the results are loaded from
`inst/extdata/`. Every code chunk that fits is shown but not evaluated.

# The data and the question

```{r data-code, eval = FALSE}
data(big_word_club, package = "bayesrules")
dat <- big_word_club
dat$school_id <- factor(dat$school_id)
keep <- c("score_ppvt", "distracted_a1", "distracted_ppvt",
          "private_school", "title1", "free_reduced_lunch", "school_id")
dat <- subset(
  dat,
  !is.na(score_ppvt) & !is.na(invalid_ppvt) & invalid_ppvt == 0L &
    complete.cases(dat[, keep])
)
dat$school_id <- droplevels(dat$school_id)
```

That leaves `r ch04$n_obs` students in `r ch04$n_schools` schools. The
outcome `score_ppvt` is a vocabulary score. Two student-level covariates
record whether the child was distracted — during the vocabulary assessment
(`distracted_ppvt`) and during an earlier assessment (`distracted_a1`).
Three school-level covariates describe the school: `private_school`,
`title1`, and `free_reduced_lunch` (the percentage of students on subsidized
meals).

The distinction between the two kinds of covariate is the whole chapter.
Student-level variables can have random slopes; school-level variables
cannot, because they do not vary within a school.

# A first model: random slopes only

```{r form-simple, eval = FALSE}
form_simple <- score_ppvt ~ private_school + title1 +
  distracted_ppvt + distracted_a1 +
  (1 + distracted_ppvt + distracted_a1 || school_id)

model_setup(form_simple, data = dat)
```

```{r design-simple, echo = FALSE}
cat(ch04$design_simple, sep = "\n")
```

Look at the Level 2 block. `model_setup()` has taken the two school-level
fixed effects and attached them to the **intercept**'s level-2 row:

```
(Intercept)     ~ 1 + private_school + title1
distracted_ppvt ~ 1
distracted_a1   ~ 1
```

That is the right thing to do, and it is worth being explicit about why.
`private_school` does not vary within a school, so a "fixed effect of
`private_school`" is a statement about how school *intercepts* differ. In
the centered parameterization of Chapter 02, it belongs in
\(\mathcal{W}_j\), shifting each school's prior mean intercept. The two
distraction slopes keep bare `~ 1` rows: their prior means are common
across schools.

Also note the rank report. Only 34 of the 47 schools have a full-rank
within-group design — the rest lack the variation in `distracted_ppvt` or
`distracted_a1` needed to identify their own slopes.

They are **not** removed. All 47 schools remain in the fit and all 47 get
coefficients; a rank-deficient school simply contributes no likelihood
information in the deficient direction, so there it is drawn from its
prior — shrunk all the way to \(\mathcal{W}_j\gamma\). That is the
statistically sensible outcome, and it is why the model tolerates schools
that a no-pooling analysis could not fit at all.

What the 34 are used for is the *level-2* rank check, which runs on the
estimable subset. That is the correct order: a school-level covariate that
happens to be constant across the estimable schools would leave \(\gamma\)
unidentified even though it varied in the raw data, and no amount of
information from the other 13 would fix it.

# Adding cross-level moderation

Now the interesting specification. Does the effect of being distracted
during the earlier assessment depend on the school's poverty level?

```{r form-full, eval = FALSE}
form_full <- score_ppvt ~ private_school + title1 + free_reduced_lunch +
  distracted_ppvt + distracted_a1 +
  free_reduced_lunch:distracted_a1 +
  (1 + distracted_ppvt + distracted_a1 || school_id)

design <- model_setup(form_full, data = dat)
design
```

```{r design-full, echo = FALSE}
cat(ch04$design_full, sep = "\n")
```

The interaction has landed in the level-2 row of `distracted_a1`:

```
(Intercept)     ~ 1 + private_school + title1 + free_reduced_lunch
distracted_ppvt ~ 1
distracted_a1   ~ 1 + free_reduced_lunch
```

This is **cross-level moderation**. A school-level variable is modifying an
individual-level random slope. In `lme4` you would write the same
interaction and get a single fixed-effect coefficient; here you can see
structurally what it does — each school's *prior mean* `distracted_a1`
slope now sits on a line in `free_reduced_lunch` instead of at a common
value. Schools still shrink toward that line, just not toward the same
point.

Reading the Level 2 block before fitting is the fastest way to confirm you
specified the model you intended. If an interaction shows up on the
intercept row when you meant it on a slope row, the formula is wrong.

# Priors

```{r prior-code, eval = FALSE}
ps <- Prior_Setup_GLMM(form_full, data = dat, pop.pwt = 0.01)
ps
```

```{r prior-out, echo = FALSE}
cat(ch04$prior_setup, sep = "\n")
```

Three prior blocks, one per random coefficient, and their dimensions follow
the level-2 rows exactly: the intercept block has a 4-by-4 `Sigma`
(intercept plus three school covariates), `distracted_ppvt` has a scalar,
and `distracted_a1` has 2-by-2. Every non-intercept `mu` is zero, per the
default `pop.effects_source = "null_effects"` described in Chapter 03.

Each block also reports `dispersion`, which is \(\tau^2_k\) — how much
individual school coefficients scatter around their level-2 fitted value —
alongside `pop.dispersion.ref`, the classical `VarCorr` estimate. The gap
between them (168.7 versus 199.3 for the intercept) is the marginal
correction described in `vignette("Chapter-A12", package = "glmbayes")`.

# Fitting

```{r fit-code, eval = FALSE}
set.seed(2024)
fit <- lmerb(
  form_full,
  data             = dat,
  pfamily_list     = pfamily_list(ps),
  dispersion_ranef = ps$group.dispersion,
  n                = 1000L,
  progbar          = FALSE
)
fit
```

```{r fit-out, echo = FALSE}
cat(ch04$fit_print, sep = "\n")
```

`exact iid` again: the observation variance is fixed and all three
\(\tau^2_k\) are `dNormal`, so the joint posterior is multivariate normal
and the 1000 draws are independent. `m_convergence` is
`r ch04$m_convergence` — no Gibbs sweeps at all.

# Reading the result

The population table has seven rows, one per level-2 parameter, and they
group by which random coefficient they modify.

**Intercept block.** `title1` is large and positive (about 17.8 points)
with a posterior SD near 7.8, so Title I schools score higher after
adjustment. `private_school` is around -3.5 with an SD of 7.8 — the data
does not distinguish it from zero. `free_reduced_lunch` is small and
negative, but its scale is per percentage point, so a 40-point difference
in subsidized-meal rate moves the intercept by about 14 points.

**Slope blocks.** `distracted_ppvt` has a population mean near -1.3:
children distracted during the vocabulary assessment score slightly lower,
though the posterior SD of 1.4 makes that weak evidence.

**The moderation.** The `distracted_a1 / free_reduced_lunch` row is about
-0.09 with an SD of 0.05. The main `distracted_a1` effect is near zero
(0.34, SD 3.7), so the story is not "distraction hurts" but "distraction
hurts more in higher-poverty schools" — and even that is modest evidence.
This is exactly the kind of claim that a single point estimate plus a
p-value tends to overstate and a posterior distribution does not.

# The full summary

```{r summary-code, eval = FALSE}
summary(fit)
```

```{r summary-out, echo = FALSE}
cat(ch04$fit_summary, sep = "\n")
```

Three sections, and each answers a different question.

**Block 2 dispersion** compares each \(\tau^2_k\) against the `lmer`
reference. All three read `dNormal` in the Prior column and show identical
percentiles, because they are fixed rather than sampled. Chapter 07 is
what makes those rows interesting.

**Block 2 level-2 fixed effects** is the population table again, with prior
means and `lmer` reference values alongside, plus `Pr(Prior_tail)`. That
last column is a directional tail probability: how far the posterior has
moved into the prior's tail. Values near 0.5 mean the data barely moved the
coefficient; small values mean it moved decisively.

**Block 1 random effects** summarizes the per-school coefficients across
schools rather than listing all 34. Use `summary(fit, groups = ...)` for
specific schools, or go to the draws directly.

# Per-school coefficients

```{r ranef-code, eval = FALSE}
head(as.data.frame(lme4::ranef(fit)), 12)
```

```{r ranef-out, echo = FALSE}
cat(ch04$ranef_head, sep = "\n")
```

Remember from Chapter 02 that these are full coefficients, not deviations:
school 1's intercept of about 97 is its expected score, not an offset. The
long format has one row per school and term, which makes it convenient for
plotting or joining back to school-level metadata.

# What changed relative to Chapter 01

| | Chapter 01 (`sleepstudy`) | Chapter 04 (`big_word_club`) |
|---|---|---|
| Level-2 rows | all `~ 1` | intercept and one slope carry covariates |
| Shrinkage target | one grand mean per coefficient | a fitted line per coefficient |
| Groups retained | 18 of 18 | 34 of 47 |
| Cross-level terms | none | `free_reduced_lunch:distracted_a1` |

The sampler did not change at all. Both fits are exact iid draws from a
multivariate normal posterior. What changed is the model, and
`model_setup()` is where you see it.

# See also

- **Chapter 05** — every accessor on a fitted object.
- **Chapter 07** — letting the variance components be sampled.
- **Chapter 13** — comparing this fit against `lme4` and `glmmTMB`.
- `?lmerb`, `?model_setup`
