---
title: "Chapter A12: Details for the Priors returned by the Prior_Setup Function"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chapter A12: Details for the Priors returned by the Prior_Setup Function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: REFERENCES.bib
reference-section-title: References
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(glmbayes)
```

# Chapter A12: Details for the Priors returned by the Prior_Setup Function

## 1. Introductory Discussion

This appendix records **precise formulas and derivations** for the prior objects
returned by `Prior_Setup()` and for the conjugate Normal--Gamma Gaussian model
used by `dNormal_Gamma()`. The goal is to connect implementation quantities
(`mu`, `Sigma`, `Sigma_0`, `dispersion`, `shape`, `rate`, and related settings)
to the **weighted likelihood notation** and **$S_{\mathrm{marg}}$** machinery in
Chapter 11 (especially Section 3.2 and Appendix A3), with steps spelled out
rather than only stated.

This chapter is a companion to the main vignettes: it emphasizes **theory**,
mapping to `pfamily` constructors, and how defaults encode prior strength.

**Roadmap.** Chapter 11 fixes notation for weighted Gaussian regression
($n_w$, $G = X^{\mathsf T} W X$, precision $\tau = 1/\phi$, and the conjugate
Normal--Gamma structure). Appendix A3 there gives closed-form posterior moments
for $\beta$ under the Zellner-type prior implied by scalar `pwt`. Chapter A02
documents how `pfamily` objects map to lower-level simulation functions. Here
we tie those ideas to **what `Prior_Setup()` actually returns** and how to pass
those fields into `dNormal()`, `dNormal_Gamma()`, and
`dIndependent_Normal_Gamma()` without mixing coefficient-scale `Sigma`,
dispersion-free `Sigma_0`, and optional fixed `dispersion` (see `?Prior_Setup`,
`?compute_gaussian_prior`).

## 2. Default Priors for Coefficient Means and Covariance Matrices

This section concerns families such as **binomial** and **Poisson** where the
usual exponential-family dispersion is **$\phi=1$** (Chapters 5, 7, and 8).
**Gaussian** models and `dNormal_Gamma` are in Section 3.

Let $n_w = \sum_i w_i$ for nonnegative **observation weights** $w_i$ in the
weighted likelihood (the same totals appear as `PriorSettings$n_effective`).
These $w_i$ are **fixed** by design and do not depend on $\beta$.

### 2.1 How prior means are determined

The Prior_Setup function provides three options for setting the prior mean vector mu. By default, it is set to correspond to the NULL (intercept only) model (`intercept_source = "null_model"`,`effects_source = "null_effects"`) . Alternatively, the user can change this to correspond to the OLS estimates for the intercept (`intercept_source =full_model"`) , the predictors (`effects_source ="full_model"`), or both. Finally, the user can also optionally provide their own custom prior mean vector `mu` directly to the Prior_Setup function.


### 2.2 Data precision $P(\beta)$.

Let $\ell(\beta)$ be the weighted log-likelihood as in Chapters 7--8, with
$\eta_i = x_i^{\mathsf T}\beta$. Define the **data precision matrix**
\[
P(\beta) := \nabla^2_\beta\bigl(-\ell(\beta)\bigr),
\]
the Hessian of the **negative** log-likelihood. With
$\ell(\beta)=\sum_i \ell_i(\eta_i)$,
\[
P(\beta) = X^{\mathsf T} W(\beta)\, X,
\qquad
W_i(\beta) := -\frac{d^2 \ell_i}{d\eta_i^2}\Big|_{\eta_i=x_i^{\mathsf T}\beta}
\ge 0
\]
(log-concavity in $\eta$; Chapter 5), and $W(\beta)$ diagonal.

Write $W_i(\beta) = w_i\,\omega_i(\beta)$ with fixed $w_i$ and mean-dependent
$\omega_i(\beta)$. Let $W_{\mathrm{obs}}=\mathrm{diag}(w_i)$ and
$\Omega(\beta)=\mathrm{diag}(\omega_i(\beta))$. Then
$W(\beta) = W_{\mathrm{obs}}\,\Omega(\beta)$ (indexwise) and
\[
P(\beta) = X^{\mathsf T} W_{\mathrm{obs}}\,\Omega(\beta)\, X.
\]

**Examples** (Chapters 7--8):

- **Poisson, log link:** $P(\beta) = X^{\mathsf T}\,\mathrm{diag}\bigl(w_i\,\mu_i(\beta)\bigr)\, X$.
- **Binomial, logit:** $P(\beta) = X^{\mathsf T}\,\mathrm{diag}\bigl(w_i\,\mu_i(\beta)(1-\mu_i(\beta))\bigr)\, X$.

### 2.3 Zellner-type prior using $P(\beta^{\ast})$

#### 2.3.1 Precision mapping and default covariance scaling

For these families, `Prior_Setup()` sets `dispersion`, `shape`, `rate`, and
`Sigma_0` to `NULL`. Let $V_0$ denote the **sampling covariance matrix** of the
fitted coefficients $\beta^{\ast}$ under the stated model ($\phi=1$). Then
\[
V_0^{-1} = P(\beta^{\ast}).
\]

**Weighted Gaussian, fixed dispersion $d$.** (See Section 3 for prior outputs.)
Then $P(\beta)=\frac{1}{d} X^{\mathsf T} W_{\mathrm{obs}} X$ for all $\beta$, and
the same identification gives $V_0^{-1}=\frac{1}{d} X^{\mathsf T} W_{\mathrm{obs}} X$
when $V_0$ is the covariance matrix at dispersion $d$. User-provided $d$ in
`compute_gaussian_prior()` sets returned `dispersion` to $d$ and rescales
`Sigma` so this scale is explicit in the returned list.

**Prior covariance:** scalar `pwt`,
$\Sigma = \frac{1-\mathrm{pwt}}{\mathrm{pwt}}\, V_0$ (equivalently
$\Sigma^{-1} = \frac{\mathrm{pwt}}{1-\mathrm{pwt}}\, P(\beta^{\ast})$)


This `Sigma` is what `Prior_Setup()` returns by default on the coefficient
scale. For Gaussian fits, the returned dispersion-free matrix is

\[
\Sigma_0 = \Sigma / d,
\]
so
\[
\Sigma_0^{-1}
=
d\,\Sigma^{-1}
=
d\,\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^{\ast}).
\]
Using $P(\beta^{\ast})=\frac{1}{d}X^{\mathsf T}WX$ in weighted Gaussian
regression gives
\[
\Sigma_0^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,X^{\mathsf T}WX,
\]
which is independent of $d$; this is the default `Sigma_0` returned by `Prior_Setup()`.

#### 2.3.2 Posterior mean and Variance under `dNormal()`

When default settings are used, the Gaussian posterior means reduce to simple
weighted averages of the fitted coefficient vector $\beta^\ast$ and prior mean
$\mu$.

**`dNormal()` (Gaussian, coefficient-scale covariance `Sigma`).**
For Gaussian likelihood precision $P(\beta^\ast)$ and prior precision
$\Sigma^{-1}$,
\[
E(\beta\mid y)
=
\bigl(P(\beta^\ast)+\Sigma^{-1}\bigr)^{-1}
\Bigl(P(\beta^\ast)\beta^\ast+\Sigma^{-1}\mu\Bigr).
\]
With the default scalar `pwt`,
\[
\Sigma^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^\ast),
\]
so
\[
\begin{aligned}
E(\beta\mid y)
&=
\left(P(\beta^\ast)+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast) \right)^{-1}
\left(P(\beta^\ast)\beta^\ast+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast)\mu\right) \\
&=
\left(\frac{1-\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast)+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast) \right)^{-1}
\left(\frac{1-\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast)\beta^\ast+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast)\mu\right) \\
&=
\left(\frac{1}{1-\mathrm{pwt}}P(\beta^\ast)\right)^{-1}
\left(\frac{1}{1-\mathrm{pwt}}P(\beta^\ast)\bigl((1-\mathrm{pwt})\beta^\ast+\mathrm{pwt}\mu\bigr)\right) \\
&=
(1-\mathrm{pwt})\beta^\ast+\mathrm{pwt}\,\mu.
\end{aligned}
\]
Thus the posterior mean is a convex combination of the likelihood estimate
$\beta^\ast$ and prior mean $\mu$: larger `pwt` gives more pull toward $\mu$.
In the limit as $\mathrm{pwt}\to 0$, it approaches $\beta^\ast$.



The posterior covariance is
\[
\mathrm{Var}(\beta\mid y)
=
\bigl(P(\beta^\ast)+\Sigma^{-1}\bigr)^{-1}.
\]
With the default scalar `pwt`,
\[
\Sigma^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^\ast),
\]
so
\[
\begin{aligned}
\mathrm{Var}(\beta\mid y)
&=
\left(P(\beta^\ast)+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^\ast)\right)^{-1} \\
&=
\left(\frac{1-\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^\ast)+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^\ast)\right)^{-1} \\
&=
\left(\frac{1}{1-\mathrm{pwt}}\,P(\beta^\ast)\right)^{-1} \\
&=
(1-\mathrm{pwt})\,P(\beta^\ast)^{-1}.
\end{aligned}
\]

Thus the posterior covariance is the likelihood-based covariance
$P(\beta^\ast)^{-1}$ shrunk by the factor $1-\mathrm{pwt}$: larger `pwt`
(stronger prior pull) gives tighter posterior uncertainty. In the limit as
$\mathrm{pwt}\to 0$, it approaches the likelihood-based covariance.

#### 2.3.3 Marginal posterior mean under `dNormal_Gamma()`

**`dNormal_Gamma()` (Gaussian conjugate Normal--Gamma, using `Sigma_0`).**
The **marginal** posterior mean is
\[
E(\beta\mid y)
=
E_{\tau\mid y}\!\left[E(\beta\mid \tau,y)\right].
\]
For fixed $\tau$,
\[
E(\beta\mid \tau,y)
=
\bigl(\tau X^{\mathsf T}W_{\mathrm{obs}}X+\tau\Sigma_0^{-1}\bigr)^{-1}
\Bigl(\tau X^{\mathsf T}W_{\mathrm{obs}}X\,\beta^\ast+\tau\Sigma_0^{-1}\mu\Bigr).
\]
Under the default scalar `pwt` calibration for `Sigma_0`,
\[
\Sigma_0^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,X^{\mathsf T}W_{\mathrm{obs}}X,
\]
so
\[
\begin{aligned}
E(\beta\mid \tau,y)
&=
\left(\tau X^{\mathsf T}W_{\mathrm{obs}}X+\tau\frac{\mathrm{pwt}}{1-\mathrm{pwt}}X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
\left(\tau X^{\mathsf T}W_{\mathrm{obs}}X\,\beta^\ast+\tau\frac{\mathrm{pwt}}{1-\mathrm{pwt}}X^{\mathsf T}W_{\mathrm{obs}}X\,\mu\right) \\
&=
\left(\frac{\tau}{1-\mathrm{pwt}}X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
\left(\frac{\tau}{1-\mathrm{pwt}}X^{\mathsf T}W_{\mathrm{obs}}X\bigl((1-\mathrm{pwt})\beta^\ast+\mathrm{pwt}\mu\bigr)\right) \\
&=
(1-\mathrm{pwt})\beta^\ast+\mathrm{pwt}\,\mu.
\end{aligned}
\]
Because this expression is free of $\tau$, averaging over $\tau\mid y$ gives
\[
E(\beta\mid y)=(1-\mathrm{pwt})\beta^\ast+\mathrm{pwt}\,\mu.
\]
Thus the marginal posterior mean has the same weighted-average interpretation:
larger `pwt` gives more pull toward $\mu$, and as $\mathrm{pwt}\to 0$ it
approaches $\beta^\ast$.
For general non-Gaussian GLMs these equalities are not exact in finite samples,
because the likelihood is not exactly quadratic in $\beta$; however, the same
weighted-average form is often a good approximation when the likelihood is
close to multivariate normal, as typically occurs in large samples.

### 2.4 Vector `pwt` and optional `sd`

**Vector `pwt`:** same Hadamard construction as above; correlations in $V_0$ are
preserved, variances scaled per coordinate.

**`sd`:** $\mathrm{pwt}_j = (V_0)_{jj}/\bigl((V_0)_{jj}+\mathrm{sd}_j^2\bigr)$;
vector `pwt` is not overwritten from scalar `n_prior`. **Gaussian** fits may
require scalar `n_prior` in addition (Section 3).

## 3. Default Priors for Dispersion, Shape, and Rate Parameters






### 3.1 Posterior pieces: what comes from likelihood + Normal block

To mirror Section 2, start from the conjugate decomposition for Gaussian models.
With
$\beta\mid\tau \sim N(\mu,(\tau\Sigma_0)^{-1})$ and a Gamma prior on residual
precision $\tau$, the Normal part and Gaussian likelihood contribute to the
posterior through the coefficient precision update (via
$X^{\mathsf T}W_{\mathrm{obs}}X$ and $\Sigma_0^{-1}$) and the scalar quadratic
term
\[
S_{\mathrm{marg}}
=
\mathrm{RSS}_w
+
(\hat\beta - \mu)^{\mathsf T}
\left(\Sigma_0 + \left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}\right)^{-1}
(\hat\beta - \mu),
\]

Under the conjugate Normal--Gamma algebra (Chapter 11), integrating out $\beta$
contributes $\frac{n_w}{2}$ to the posterior Gamma shape for $\tau$ (not an
additive $\frac{p}{2}$ in this parameterization). With prior Gamma shape
$a_0=\texttt{shape}$, the posterior satisfies
\[
a_n = a_0 + \frac{n_w}{2},
\qquad
b_n = b_0 + \frac{1}{2}S_{\mathrm{marg}},
\]
with $n_w =$ `n_effective` and $p=\texttt{ncol}(X)$ (number of coefficients).

### 3.2 Prior-strength parameterization from `pwt`

For scalar `pwt`, the same strength mapping used in Section 2 is
\[
n_{\mathrm{prior}}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}.
\]
Interpretation: larger `pwt` means more prior pseudo-information; as
$\mathrm{pwt}\to 0$, the setup approaches likelihood-dominated behavior.

The dispersion-free covariance used in `dNormal_Gamma()` remains
\[
\Sigma_0 = \frac{\Sigma_{\mathrm{pre}}}{d_{\mathrm{OLS}}},
\]
and for scalar `pwt` (Section 2.3.1) this is exactly
\[
\Sigma_0^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,X^{\mathsf T}W_{\mathrm{obs}}X,
\]
equivalently
\[
\Sigma_0
=
\frac{1-\mathrm{pwt}}{\mathrm{pwt}}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
the same Zellner $g$-type scaling relative to
$\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}$ used in Chapter 11
Appendix A3.

Substituting this $\Sigma_0$ into $S_{\mathrm{marg}}$ from Section 3.1 gives
\[
\begin{aligned}
S_{\mathrm{marg}}
&=
\mathrm{RSS}_w
+
(\hat\beta - \mu)^{\mathsf T}
\left(\Sigma_0 + \left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}\right)^{-1}
(\hat\beta - \mu) \\
&=
\mathrm{RSS}_w
+
(\hat\beta - \mu)^{\mathsf T}
\left(
\frac{1-\mathrm{pwt}}{\mathrm{pwt}}\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
+
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
\right)^{-1}
(\hat\beta - \mu) \\
&=
\mathrm{RSS}_w
+
(\hat\beta - \mu)^{\mathsf T}
\left(
\frac{1}{\mathrm{pwt}}\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
\right)^{-1}
(\hat\beta - \mu) \\
&=
\mathrm{RSS}_w
+
\mathrm{pwt}\,
(\hat\beta - \mu)^{\mathsf T}
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)
(\hat\beta - \mu).
\end{aligned}
\]
So under scalar `pwt`, the prior-mean penalty term in $S_{\mathrm{marg}}$ is
scaled directly by `pwt`.

### 3.3 `dNormal_Gamma()`: default `shape`, `rate`, and implied posterior Gamma parameters

This subsection develops the **Normal--Gamma** prior returned by `Prior_Setup()` for
Gaussian models when the sampling model is taken to be **`dNormal_Gamma()`** with the
default **`shape_df = "n_prior"`**. (Other `shape_df` settings swap in
$n_{\mathrm{prior}}\pm p$ in the implementation’s shape numerator; see `?Prior_Setup`.
The algebra below is the same after renaming that numerator.)

#### 3.3.1 Default calibration and posterior Gamma shape/rate

With scalar `pwt`, define
\[
n_{\mathrm{prior}}=\frac{\mathrm{pwt}}{1-\mathrm{pwt}}.
\]
`compute_gaussian_prior()` then sets the prior Gamma on residual precision $\tau$ with
\[
\mathrm{shape}=\frac{n_{\mathrm{prior}}+1}{2}.
\]
Gaussian calibration requires **$n_w>p$** (strictly, so $n_w-p$ is a positive
residual degrees-of-freedom denominator).

Calibrated **dispersion** (posterior mean $\sigma^2$ under this Gamma update) is
\[
\mathrm{dispersion}
=
\frac{S_{\mathrm{marg}}}{n_w-p}.
\]
The prior **rate** is
\[
\mathrm{rate}
=
\frac{1}{2}\,\frac{n_{\mathrm{prior}}+p-1}{n_w-p}\,S_{\mathrm{marg}}.
\]
The numerator $n_{\mathrm{prior}}+p-1$ pairs with $\mathrm{shape}=(n_{\mathrm{prior}}+1)/2$
so that the implied posterior for $\sigma^2$ matches $\mathrm{dispersion}$ (derived below).
\[
\Sigma_{\mathrm{calibrated}}
=
\frac{n_w}{n_{\mathrm{prior}}}\,\mathrm{dispersion}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]
The prior rate requires $n_{\mathrm{prior}}+p>1$ so $\mathrm{rate}>0$ (see `?compute_gaussian_prior`).

Plugging these defaults into the posterior Gamma update,
\[
a_n = \mathrm{shape} + \frac{n_w}{2},
\qquad
b_n = \mathrm{rate} + \frac{1}{2}S_{\mathrm{marg}}.
\]
Since $\mathrm{shape}=(n_{\mathrm{prior}}+1)/2$,
\[
a_n=\frac{n_{\mathrm{prior}}+n_w+1}{2},
\qquad
a_n-1=\frac{n_{\mathrm{prior}}+n_w-1}{2}.
\]
For $b_n$, substitute $\mathrm{rate}=\frac{1}{2}S_{\mathrm{marg}}(n_{\mathrm{prior}}+p-1)/(n_w-p)$:
\[
\begin{aligned}
b_n
&=
\frac{1}{2}\,\frac{n_{\mathrm{prior}}+p-1}{n_w-p}\,S_{\mathrm{marg}}
+\frac{1}{2}S_{\mathrm{marg}} \\
&=
\frac{S_{\mathrm{marg}}}{2}\left(\frac{n_{\mathrm{prior}}+p-1}{n_w-p}+1\right) \\
&=
\frac{S_{\mathrm{marg}}}{2}\cdot\frac{n_{\mathrm{prior}}+p-1+n_w-p}{n_w-p} \\
&=
\frac{S_{\mathrm{marg}}}{2}\cdot\frac{n_{\mathrm{prior}}+n_w-1}{n_w-p}.
\end{aligned}
\]
**Posterior mean dispersion (step-by-step).** For $\tau\mid y\sim\mathrm{Gamma}(a_n,b_n)$ in the
shape--rate parameterization used here, $E(\tau^{-1}\mid y)=b_n/(a_n-1)$ when $a_n>1$. Therefore
\[
\begin{aligned}
\frac{b_n}{a_n-1}
&=
\frac{\displaystyle\frac{S_{\mathrm{marg}}}{2}\cdot\frac{n_{\mathrm{prior}}+n_w-1}{n_w-p}}
{\displaystyle\frac{n_{\mathrm{prior}}+n_w-1}{2}} \\
&=
\frac{S_{\mathrm{marg}}}{n_w-p}.
\end{aligned}
\]
The factor $\frac{1}{2}(n_{\mathrm{prior}}+n_w-1)$ cancels between numerator and denominator.
By construction $\mathrm{dispersion}=S_{\mathrm{marg}}/(n_w-p)$, so
\[
\frac{b_n}{a_n-1}
=
\frac{S_{\mathrm{marg}}}{n_w-p}
=
\mathrm{dispersion}.
\]
The denominator $n_w-p$ aligns the weak-prior limit of $\mathrm{dispersion}$ with the usual
weighted **$n_w-p$** residual variance rather than $n_w-2$.

#### 3.3.2 Posterior dispersion and covariance under `dNormal_Gamma()`

Continuing the **`dNormal_Gamma()`** default prior from Section 3.3.1:

**Posterior expected dispersion.** From the posterior Gamma law for $\tau\mid y$,
\[
E(\sigma^2\mid y)=E(\tau^{-1}\mid y)=\frac{b_n}{a_n-1},
\]
and Section 3.3.1 implies directly
\[
E(\sigma^2\mid y)
=
\frac{S_{\mathrm{marg}}}{n_w-p}
=
\mathrm{dispersion}.
\]
Substitute
\[
S_{\mathrm{marg}}
=
\mathrm{RSS}_w
+
\mathrm{pwt}\,(\hat\beta-\mu)^{\mathsf T}(X^{\mathsf T}W_{\mathrm{obs}}X)(\hat\beta-\mu),
\]
to obtain
\[
E(\sigma^2\mid y)=
\frac{\mathrm{RSS}_w
+
\mathrm{pwt}\,(\hat\beta-\mu)^{\mathsf T}(X^{\mathsf T}W_{\mathrm{obs}}X)(\hat\beta-\mu)}
{n_w-p}.
\]
As $\mathrm{pwt}\to 0$ (equivalently $n_{\mathrm{prior}}\to 0$), this converges to
\[
E(\sigma^2\mid y)\to \frac{\mathrm{RSS}_w}{n_w-p},
\]
the usual weighted Gaussian residual-variance estimate (denominator $n_w-p$).

**Posterior covariance of coefficients.** Using total variance,
\[
\mathrm{Var}(\beta\mid y)
=
E_{\tau\mid y}\!\left[\mathrm{Var}(\beta\mid \tau,y)\right]
+
\mathrm{Var}_{\tau\mid y}\!\left[E(\beta\mid \tau,y)\right].
\]
From Section 2.3.3, $E(\beta\mid\tau,y)$ is free of $\tau$, so the second term
is $0$. Also,
\[
\mathrm{Var}(\beta\mid \tau,y)=\tau^{-1}V_n,
\qquad
V_n=\left(X^{\mathsf T}W_{\mathrm{obs}}X+\Sigma_0^{-1}\right)^{-1},
\]
hence
\[
\mathrm{Var}(\beta\mid y)=\frac{b_n}{a_n-1}\,V_n.
\]
With scalar `pwt`,
\[
V_n
=
\left(\frac{1}{1-\mathrm{pwt}}X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
=
\frac{n_w}{n_w+n_{\mathrm{prior}}}\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
so
\[
\mathrm{Var}(\beta\mid y)
=
\frac{S_{\mathrm{marg}}}{n_w-p}\,
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]
There is no cancellation to a single $(n_w+n_{\mathrm{prior}}-2)$ denominator
here; both residual df $(n_w-p)$ and the Zellner factor
$n_w/(n_w+n_{\mathrm{prior}})$ appear. Substituting $S_{\mathrm{marg}}$ gives
\[
\mathrm{Var}(\beta\mid y)
=
\frac{\mathrm{RSS}_w
+
\mathrm{pwt}\,(\hat\beta-\mu)^{\mathsf T}(X^{\mathsf T}W_{\mathrm{obs}}X)(\hat\beta-\mu)}
{n_w-p}
\cdot
\frac{n_w}{n_w+n_{\mathrm{prior}}}
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
and therefore, as $\mathrm{pwt}\to 0$ (equivalently $n_{\mathrm{prior}}\to 0$),
\[
\mathrm{Var}(\beta\mid y)\to
\frac{\mathrm{RSS}_w}{n_w-p}\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
the classical weighted least-squares covariance.

#### 3.3.3 Posterior covariance under `dNormal()` with default dispersion

For comparison, **`dNormal()`** fixes dispersion at the calibrated value rather than
integrating over $\tau$ as in Sections 3.3.1--3.3.2. From Section 2.3.2, under scalar `pwt`,
\[
\mathrm{Var}(\beta\mid y,\sigma^2)
=
(1-\mathrm{pwt})\,P(\beta^\ast)^{-1}.
\]
For Gaussian weighted regression,
$P(\beta^\ast)=\sigma^{-2}X^{\mathsf T}W_{\mathrm{obs}}X$, so
\[
\mathrm{Var}(\beta\mid y,\sigma^2)
=
(1-\mathrm{pwt})\,\sigma^2\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]
Using
$n_{\mathrm{prior}}=\mathrm{pwt}/(1-\mathrm{pwt})$, equivalently
$1-\mathrm{pwt}=n_w/(n_w+n_{\mathrm{prior}})$, this is
\[
\mathrm{Var}(\beta\mid y,\sigma^2)
=
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,\sigma^2
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]

If one plugs in the default dispersion from Section 3.3.2,
\[
\sigma^2_{\mathrm{default}}
=
E(\sigma^2\mid y)
=
\frac{S_{\mathrm{marg}}}{n_w-p},
\]
then
\[
\begin{aligned}
\mathrm{Var}(\beta\mid y,\sigma^2_{\mathrm{default}})
&=
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,
\frac{S_{\mathrm{marg}}}{n_w-p}
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\end{aligned}
\]
This matches the `dNormal_Gamma()` expression in Section 3.3.2, so `dNormal()`
with default dispersion has the same convergence property:
\[
\mathrm{Var}(\beta\mid y)\to
\frac{\mathrm{RSS}_w}{n_w-p}\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
\quad\text{as}\quad \mathrm{pwt}\to 0.
\]

This keeps `shape`, `rate`, `dispersion`, and returned covariance objects
internally consistent under the package calibration.

### 3.4 Returned objects and model mapping

For Gaussian workflows, `Prior_Setup()` returns a matched set:
- `Sigma_0` (dispersion-free, for `dNormal_Gamma()`)
- `shape`, `rate` (Gamma prior on residual precision)
- `dispersion` and `Sigma` (coefficient-scale companion quantities)

Always pass **`Sigma_0 = ps$Sigma_0`** into `dNormal_Gamma()` (not `ps$Sigma`):

```r
ps <- Prior_Setup(formula, data, family = gaussian(), n_prior = n0)
lmb(
  formula,
  data,
  pfamily = dNormal_Gamma(
    ps$mu,
    Sigma_0 = ps$Sigma_0,
    shape     = ps$shape,
    rate      = ps$rate
  )
)
```

For matrix-input workflows, Chapter A02: `prior_list` for `rNormalGamma_reg()`
uses `Sigma = ps$Sigma_0` (dispersion-free slot) together with `shape` and
`rate`.

### 3.5 Variants (`pwt` vector, `sd`, user `Sigma`)

With vector `pwt`, $\Sigma_{\mathrm{pre}}$ uses the Hadamard scaling from
Section 2.4; coefficient blending remains elementwise
$(1-\mathrm{pwt}_j)\hat\beta_j + \mathrm{pwt}_j\mu_j$.

With **`sd`**, `pwt` is derived from $\operatorname{diag}(V_0)$ (Section 2.4),
and coefficient-scale $\Sigma_{\mathrm{pre}}$ is passed as `Sigma` into
`compute_gaussian_prior()`. The helper then reconciles returned
`Sigma`, `Sigma_0`, and `dispersion` consistently.

With user-supplied `Sigma`, the helper enforces `Sigma_0 = Sigma / dispersion`
(using calibrated or user-provided `dispersion`), so `dNormal` and
`dNormal_Gamma` parameterizations stay aligned. When only `dispersion` is
provided, returned `Sigma` is `Sigma_0 * dispersion`.

## 4. Other Models With Unknown Dispersion

This appendix focused on the `Prior_Setup()` $\to$ `dNormal_Gamma()` Gaussian
path. Other compound specifications reuse parts of the same output but expect
**different** slots:

- **`dIndependent_Normal_Gamma()`** takes **undivided** `ps$Sigma` (coefficient
  scale) plus the same `shape` and `rate`; do **not** substitute `Sigma_0` for
  `Sigma` there.
- **`dGamma()`** with fixed coefficients uses `shape`, `rate`, and typically
  `coefficients` from the prior list for residual precision updates.

See Chapter A02 for `prior_list` templates (`rNormalGamma_reg()` vs
`rindepNormalGamma_reg()`), Chapter 11 Section 3.3 for the independent
Normal--Gamma model, and Chapter A07 for truncated-Gamma / envelope extensions.
