Yaml

---
title: "Chapter 2: Estimating Bayesian Linear Models using the lmb function"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chapter 2: Estimating Bayesian Linear Models using the lmb function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(glmbayes)
```

# 1. Introductory discussion

The **lmb** function has been designed as a Bayesian version of the classical **lm** function and much of its outputs and methods mirrors those of the **lm** function. In this section we illustrate its use next to the **lm** function and focus in on some of the core outputs and methods. For purpose of this section, we may use of a default prior specification that comes with the **Prior_Setup** function for the package. In the next chapter, we dive into a deeper discussion of Prior specifications for Bayesian Linear models.

# 2. Sample Data - Dobson's Plant weight data


```{r Plant_Data}
## Annette Dobson (1990) "An Introduction to Generalized Linear Models".
## Page 9: Plant Weight Data.
ctl <- c(4.17,5.58,5.18,6.11,4.50,4.61,5.17,4.53,5.33,5.14)
trt <- c(4.81,4.17,4.41,3.59,5.87,3.83,6.03,4.89,4.32,4.69)
group <- gl(2, 10, 20, labels = c("Ctl","Trt"))
weight <- c(ctl, trt)
```

```{r Plant_w_wo_intercept}
lm.D9 <- lm(weight ~ group)
lm.D90 <- lm(weight ~ group - 1) # omitting intercept
```



```{r Plant_summary}
summary(lm.D90)
```


```{r Plant_Prior1}
ps=Prior_Setup(weight ~ group-1,family=gaussian())
mu=ps$mu
V=ps$Sigma
disp_ML=ps$dispersion
mu
V
disp_ML
n_prior=2
shape=n_prior/2
rate= disp_ML*shape
lmb.D90=lmb(weight ~ group-1,dNormal(mu,V,dispersion=disp_ML))
summary(lmb.D90)
```


```{r Plant_Prior}
ps=Prior_Setup(weight ~ group,family=gaussian())
mu=ps$mu
V=ps$Sigma
mu[1,1]=mean(weight)

Prior_Check(weight ~ group,family =gaussian(),
            pfamily=dNormal(mu=mu,Sigma=V))

## May move this step inside the Prior_Check function
lm.D9 <- lm(weight ~ group,x=TRUE,y=TRUE)
disp_ML=sigma(lm.D9)^2
n_prior=2
shape=n_prior/2
rate= disp_ML*shape
```

```{r Models}
lm.D9=lm(weight~group)
lmb.D9=lmb(weight ~ group,dNormal_Gamma(mu,V/disp_ML,shape=shape,rate=rate))
lmb.D9_v2=lmb(weight ~ group,dNormal(mu,V,dispersion=disp_ML))

glm.D9=glm(weight ~ group,family=gaussian())
glmb.D9=glmb(weight ~ group,family=gaussian(),dNormal_Gamma(mu,V/disp_ML,shape=shape,rate=rate))
glmb.D9_v2=glmb(weight ~ group,family=gaussian(),dNormal(mu,V,dispersion=disp_ML))


```


```{r lm_print}
print(lm.D9)
print(lmb.D9)
print(lmb.D9_v2)

```

```{r lm_summary}
summary(lm.D9)
summary(lmb.D9)
summary(lmb.D9_v2)

```


```{r lm_residuals}
residuals(lm.D9)
colMeans(residuals(lmb.D9))
colMeans(residuals(lmb.D9_v2))

residuals(glm.D9)

```


```{r lm_fitted}
fitted(lm.D9)
colMeans(fitted(lmb.D9))
colMeans(fitted(lmb.D9_v2))

```


```{r lm_predicted}
## As this is linear scale, these wll match the above
predict(lm.D9)
colMeans(predict(lmb.D9))
colMeans(predict(lmb.D9_v2))

```

```{r methods_lm}
## As this is linear scale, these wll match the above
methods(class="lm")
```


```{r lm_logLik}
## Now produces correct results for both versions
logLik(lm.D9)
colMeans(logLik(lmb.D9))
colMeans(logLik(lmb.D9_v2))

```

```{r lm_confint}
## Now produces correct results for both versions
confint(lm.D9)
confint(lmb.D9)
confint(lmb.D9_v2)

```
```{r lm_deviance}
## Note: Bayesian Deviance Estimates quite a bit larger than the Classical
## This is because Classical minimizez the deviance
## while the Bayesian model simulates over a range of possible values

deviance(lm.D9)
mean(deviance(lmb.D9))
mean(deviance(lmb.D9_v2))

deviance(glm.D9)
mean(deviance(glmb.D9))
mean(deviance(glmb.D9_v2))

```

```{r lm_vcov}
## Note: Bayesian Deviance Estimates quite a bit larger than the Classical
## This is because Classical minimizez the deviance
## while the Bayesian model simulates over a range of possible values

vcov(lm.D9)
vcov(lmb.D9)
vcov(lmb.D9_v2)

vcov(glm.D9)
vcov(glmb.D9)
vcov(glmb.D9_v2)

```

```{r extractAIC}
## Produces simular results to the AIC function (for lm) and extractAIC (for glm)
AIC(lm.D9)

extractAIC(lm.D9)
extractAIC(lmb.D9)
extractAIC(lmb.D9_v2)

extractAIC(glm.D9)
extractAIC(glmb.D9)
extractAIC(glmb.D9_v2)

```

```{r dummy.coef }
## Produces simular results to the AIC function (for lm) and extractAIC (for glm)

dummy.coef(lm.D9)
dummy.coef(lmb.D9)
dummy.coef(lmb.D9_v2)

dummy.coef(glm.D9)
dummy.coef(glmb.D9)
dummy.coef(glmb.D9_v2)

```


```{r model.frame}
## Produces simular results to the AIC function (for lm) and extractAIC (for glm)

model.frame(lm.D9)
model.frame(lmb.D9)
model.frame(lmb.D9_v2)

model.frame(glm.D9)
model.frame(glmb.D9)
model.frame(glmb.D9_v2)

```

```{r model.matrix}
## Produces simular results to the AIC function (for lm) and extractAIC (for glm)

model.matrix(lm.D9)
model.matrix(lmb.D9)
model.matrix(lmb.D9_v2)

model.matrix(glm.D9)
model.matrix(glmb.D9)
model.matrix(glmb.D9_v2)

```