---
title: "Chapter 03: The prior workflow"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 03: The prior workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

# Chapter 03: The prior workflow

Every fit in this package goes through the same three steps.

```
formula + data
     |
     v
model_setup()          what model am I actually fitting?
     |
     v
Prior_Setup_GLMM()     calibrate priors against a reference lmer fit
     |
     v
pfamily_list()         one prior object per random-effect coefficient
     |
     v
lmerb() / glmerb()
```

You can skip the first step — `lmerb()` calls it internally — but you
should not, because `model_setup()` is where you find out whether the model
you wrote is the model you meant.

```{r data}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)
form <- Reaction ~ Days_c + (1 + Days_c || Subject)
```

# Step 1: `model_setup()`

```{r model-setup}
design <- model_setup(form, data = dat)
design
```

The printed report has three parts.

**Level 1** lists the within-group design: which coefficients vary by
group, how many observations there are, and how many groups.

**Level 2** lists, for each random coefficient, which group-level
predictors enter its prior mean. A bare `~ 1` means "shrink toward a single
common value". Anything more means "shrink toward a fitted line".

**Rank** reports two separate checks. `Full-rank D_j: 18 of 18 groups` is
the within-group check: does each group have enough distinct covariate
values to identify its own coefficients? Groups that fail stay in the fit
but are set aside for the next check. The Level 2 Rank block is the
across-group check, run on the groups that passed: do they span the level-2
predictor space? `popef.rank_ok` summarizes it.

The asymmetry matters. A rank-deficient group is survivable — it simply
learns nothing from its data in that direction. A failed *level-2* check is
not: the population parameter \(\gamma\) is then unidentified by every
data-bearing group, and the chain is null recurrent in that direction even
with a proper prior. Both conditions are derived in
`vignette("Chapter-B08", package = "lmebayesCore")`.

The object itself carries the matrices:

```{r design-contents}
names(design)
dim(design$D)                 # stacked within-group design
str(design$groupef.names)     # names of the random coefficients
```

# Step 2: `Prior_Setup_GLMM()`

```{r prior-setup}
ps <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01)
ps
```

This is the calibration step. It fits a reference `lmer()` model and uses
its estimates to place priors on a scale that is commensurate with the
data, so you specify *how much* prior information you want rather than
guessing at numeric hyperparameters.

## The prior-weight idea

The central argument is `pop.pwt`, the **population prior weight**. Setting
`pop.pwt = 0.01` means: give the population-level prior about as much
weight as one percent of the effective sample. The prior covariance is
scaled accordingly.

| Setting | Effect |
|---|---|
| `pop.pwt = 0.01` (default-ish) | Weak prior; posterior close to `lmer()` |
| `pop.pwt = 0.05` | Noticeably regularizing |
| `pop.pwt = 0.25` | Strong shrinkage toward the prior mean |

The equivalent alternative is `pop.nprior`, which states the prior weight
as a number of observations rather than a fraction, or `pop.sd`, which
sets prior standard deviations directly.

## What gets centered where

Two arguments control the prior *mean*.

- `pop.intercept_source` — `"null_model"` (default) centers the intercept
  at the intercept-only fit, so the prior says "I know roughly the overall
  level" without pre-judging any effect.
- `pop.effects_source` — `"null_effects"` (default) centers every
  non-intercept coefficient at **zero**. This is the conservative choice:
  the prior expresses no opinion about direction, and any signal in the
  posterior comes from the data.

Setting either to `"full_model"` centers the prior on the reference `lmer`
estimates. That is legitimate for sensitivity analysis but is not a prior
in the usual sense — it uses the data twice. Use `pop.mu` to supply your
own means when you have real external information.

## The fields

```{r prior-fields}
names(ps)
```

The names are prefixed by which part of the model they configure.

| Prefix | Configures |
|---|---|
| `pop.*` | Block 2: the population coefficients \(\gamma\) and the variance components \(\tau^2_k\) |
| `group.*` | The observation-level dispersion \(\sigma^2\) |

The two you pass on to the sampler are:

```{r prior-key-fields}
ps$group.dispersion     # sigma^2: observation variance
str(ps$pop.prior_list, max.level = 2)
```

`ps$group.dispersion` is the fixed observation variance, taken from the
reference fit. `ps$pop.prior_list` holds, for each random coefficient, the
prior mean vector `mu`, prior covariance `Sigma`, and the dispersion
`tau^2_k` that governs how tightly group coefficients cluster around
\(\mathcal{W}_j\gamma\).

## The ING window

When you ask for sampled variance components (Chapter 07),
`Prior_Setup_GLMM()` also computes a truncation window:

```{r ing-window}
ps_ing <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01,
                           pop.dispersion.pwt = 0.2)
ps_ing$pop.ing_prior
```

`disp_lower` and `disp_upper` bound the support of each \(\tau^2_k\). They
are quantiles of the *limiting posterior* for the variance component, not
of the prior — the reasoning is in `inst/ING_TRUNCATION_WINDOW.md` in the
package sources. The lower bound matters for more than propriety: the
sampler evaluates its convergence rate at `disp_lower`, because smaller
\(\tau^2\) means tighter coupling between blocks and therefore a slower
chain. A bound computed there is valid across the whole window. See
`vignette("Chapter-B03", package = "lmebayesCore")`.

# Step 3: `pfamily_list()`

```{r pfamily}
pf <- pfamily_list(ps)
pf
```

One `pfamily` object per random-effect coefficient. The prior family
determines what happens to that coefficient's variance component.

| Family | \(\tau^2_k\) | Chapter |
|---|---|---|
| `dNormal` | Fixed at the reference estimate | 01, 04 |
| `dIndependent_Normal_Gamma` | Sampled, with a Gamma prior on the precision | 07 |

Ask for the second with `ptypes`:

```{r pfamily-ing}
pf_ing <- pfamily_list(ps_ing, ptypes = "dIndependent_Normal_Gamma")
class(pf_ing[["Days_c"]])
```

`ptypes` accepts one value applied to all coefficients, or a named vector
if you want to sample some variance components and fix others.

# Overriding a prior by hand

Nothing forces you to use the calibrated object. A `pfamily_list` is a
plain named list, so you can replace an element:

```{r manual-prior}
pf_manual <- pf
pf_manual[["Days_c"]] <- dNormal(
  mu         = matrix(5, 1, 1),      # prior mean slope: 5 ms per day
  Sigma      = matrix(4, 1, 1),      # prior SD 2 on that mean
  dispersion = 25                    # tau^2: subject slopes vary with SD 5
)
pf_manual[["Days_c"]]$prior_list[c("mu", "Sigma", "dispersion")]
```

Three numbers, three different jobs, and they are easy to confuse:

- **`mu`** — the prior mean of the *population* slope \(\gamma_k\).
- **`Sigma`** — how uncertain you are about that population slope.
- **`dispersion`** — \(\tau^2_k\), how much *individual subjects* scatter
  around it. This is the variance component, not prior uncertainty.

Setting `Sigma` small says "I know the average slope". Setting `dispersion`
small says "everyone's slope is nearly the same". They are independent
statements.

# Putting it together

```{r fit, results = "hide"}
set.seed(1)
fit <- lmerb(
  form,
  data             = dat,
  pfamily_list     = pf,
  dispersion_ranef = ps$group.dispersion,
  n                = 200L,
  progbar          = FALSE
)
```

```{r fit-print}
fit
```

## Checking prior sensitivity

Because the whole prior is one argument, sensitivity analysis is a loop
over `pop.pwt`:

```{r sensitivity, results = "hide"}
pwts <- c(0.01, 0.10, 0.50)
res <- lapply(pwts, function(w) {
  ps_w <- Prior_Setup_GLMM(form, data = dat, pop.pwt = w)
  f <- lmerb(form, data = dat,
             pfamily_list     = pfamily_list(ps_w),
             dispersion_ranef = ps_w$group.dispersion,
             n = 200L, progbar = FALSE)
  unlist(f$popef.means)
})
```

```{r sensitivity-table}
tab <- do.call(rbind, res)
rownames(tab) <- paste0("pop.pwt = ", pwts)
round(tab, 3)
```

The intercept barely moves — it is centered at the null-model intercept, so
prior and data already agree. The `Days_c` slope is centered at zero, so
heavier prior weight pulls it toward zero. That contrast is the whole point
of `pop.effects_source = "null_effects"`: the prior is deliberately
skeptical about effects, and you can see exactly how much skepticism costs.

# See also

- **Chapter 04** — group-level covariates, where Level 2 stops being `~ 1`.
- **Chapter 07** — sampling the variance components.
- `?Prior_Setup_GLMM`, `?pfamily_list`, `?model_setup`
- `vignette("Chapter-A12", package = "glmbayes")` — the derivations behind
  the Gaussian prior calibration this inherits.
