---
title: "Chapter A12: Technical Derivations for Priors Returned by `Prior_Setup()"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chapter A12: Technical Derivations for Priors Returned by `Prior_Setup()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: REFERENCES.bib
reference-section-title: References
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(glmbayes)
```

# Chapter A12: Technical Derivations for Priors Returned by `Prior_Setup()`

## 1. Introduction

This appendix provides a complete and self‑contained derivation of the prior objects returned by
`Prior_Setup()` and of the Gaussian prior families used throughout **glmbayes**. Its purpose is to make
explicit how the returned quantities—`mu`, `Sigma`, `Sigma_0`, `dispersion`, `shape`, `rate`, and related
fields—arise from the **weighted Gaussian likelihood**, the **Normal–Gamma algebra**, and the
**Zellner‑type calibration** used by the package.

Unlike Chapter 11, which focuses on modeling workflow and examples, this chapter focuses on the
**mathematical structure** underlying the priors:

- the weighted Gaussian likelihood and its precision form,
- the conjugate Normal–Gamma posterior and the derivation of its conditional and marginal components,
- the construction of the Zellner‑type dispersion‑free matrix `Sigma_0`,
- the mapping between `pwt`, `n_prior`, and prior strength,
- the independent Normal–Gamma prior and its log‑concavity,
- and the shared **weak‑prior limit** to which all Gaussian prior families converge.

All formulas needed by the main vignettes are derived here from first principles. No results are imported
from Chapter 11; instead, Chapter 11 now serves as a conceptual overview, while this appendix provides the
full algebraic details.

The goal is to make the calibration used by `Prior_Setup()` transparent, reproducible, and extensible, so
that users can confidently interpret or modify the priors supplied to `dNormal()`, `dNormal_Gamma()`,
and `dIndependent_Normal_Gamma()`.





## 1. Introductory Discussion

This appendix records **precise formulas and derivations** for the prior objects
returned by `Prior_Setup()` and for the conjugate Normal--Gamma Gaussian model
used by `dNormal_Gamma()`. The goal is to connect implementation quantities
(`mu`, `Sigma`, `Sigma_0`, `dispersion`, `shape`, `rate`, and related settings)
to the **weighted likelihood notation** and **$S_{\mathrm{marg}}$** machinery in
Chapter 11 (especially Section 3.2 and Appendix A3), with steps spelled out
rather than only stated.

This chapter is a companion to the main vignettes: it emphasizes **theory**,
mapping to `pfamily` constructors, and how defaults encode prior strength.

**Roadmap.** Chapter 11 fixes notation for weighted Gaussian regression
($n_w$, $G = X^{\mathsf T} W X$, precision $\tau = 1/\phi$, and the conjugate
Normal--Gamma structure). Appendix A3 there gives closed-form posterior moments
for $\beta$ under the Zellner-type prior implied by scalar `pwt`. Chapter A02
documents how `pfamily` objects map to lower-level simulation functions. Here
we tie those ideas to **what `Prior_Setup()` actually returns** and how to pass
those fields into `dNormal()`, `dNormal_Gamma()`, and
`dIndependent_Normal_Gamma()` without mixing coefficient-scale `Sigma`,
dispersion-free `Sigma_0`, and optional fixed `dispersion` (see `?Prior_Setup`,
`?compute_gaussian_prior`).

## 2. Default Priors for Coefficient Means and Covariance Matrices

This section concerns families such as **binomial** and **Poisson** where the
usual exponential-family dispersion is **$\phi=1$** (Chapters 5, 7, and 8).
**Gaussian** models and `dNormal_Gamma` are in Section 3.

Let $n_w = \sum_i w_i$ for nonnegative **observation weights** $w_i$ in the
weighted likelihood (the same totals appear as `PriorSettings$n_effective`).
These $w_i$ are **fixed** by design and do not depend on $\beta$.

### 2.1 How prior means are determined

The Prior_Setup function provides three options for setting the prior mean vector mu. By default, it is set to correspond to the NULL (intercept only) model (`intercept_source = "null_model"`,`effects_source = "null_effects"`) . Alternatively, the user can change this to correspond to the OLS estimates for the intercept (`intercept_source =full_model"`) , the predictors (`effects_source ="full_model"`), or both. Finally, the user can also optionally provide their own custom prior mean vector `mu` directly to the Prior_Setup function.


### 2.2 Data precision $P(\beta)$.

Let $\ell(\beta)$ be the weighted log-likelihood as in Chapters 7--8, with
$\eta_i = x_i^{\mathsf T}\beta$. Define the **data precision matrix**
\[
P(\beta) := \nabla^2_\beta\bigl(-\ell(\beta)\bigr),
\]
the Hessian of the **negative** log-likelihood. With
$\ell(\beta)=\sum_i \ell_i(\eta_i)$,
\[
P(\beta) = X^{\mathsf T} W(\beta)\, X,
\qquad
W_i(\beta) := -\frac{d^2 \ell_i}{d\eta_i^2}\Big|_{\eta_i=x_i^{\mathsf T}\beta}
\ge 0
\]
(log-concavity in $\eta$; Chapter 5), and $W(\beta)$ diagonal.

Write $W_i(\beta) = w_i\,\omega_i(\beta)$ with fixed $w_i$ and mean-dependent
$\omega_i(\beta)$. Let $W_{\mathrm{obs}}=\mathrm{diag}(w_i)$ and
$\Omega(\beta)=\mathrm{diag}(\omega_i(\beta))$. Then
$W(\beta) = W_{\mathrm{obs}}\,\Omega(\beta)$ (indexwise) and
\[
P(\beta) = X^{\mathsf T} W_{\mathrm{obs}}\,\Omega(\beta)\, X.
\]

**Examples** (Chapters 7--8):

- **Poisson, log link:** $P(\beta) = X^{\mathsf T}\,\mathrm{diag}\bigl(w_i\,\mu_i(\beta)\bigr)\, X$.
- **Binomial, logit:** $P(\beta) = X^{\mathsf T}\,\mathrm{diag}\bigl(w_i\,\mu_i(\beta)(1-\mu_i(\beta))\bigr)\, X$.

### 2.3 Zellner-type prior using $P(\beta^{\ast})$

#### 2.3.1 Precision mapping and default covariance scaling

For these families, `Prior_Setup()` sets `dispersion`, `shape`, `rate`, and
`Sigma_0` to `NULL`. Let $V_0$ denote the **sampling covariance matrix** of the
fitted coefficients $\beta^{\ast}$ under the stated model ($\phi=1$). Then
\[
V_0^{-1} = P(\beta^{\ast}).
\]

**Weighted Gaussian, fixed dispersion $d$.** (See Section 3 for prior outputs.)
Then $P(\beta)=\frac{1}{d} X^{\mathsf T} W_{\mathrm{obs}} X$ for all $\beta$, and
the same identification gives $V_0^{-1}=\frac{1}{d} X^{\mathsf T} W_{\mathrm{obs}} X$
when $V_0$ is the covariance matrix at dispersion $d$. User-provided $d$ in
`compute_gaussian_prior()` sets returned `dispersion` to $d$ and rescales
`Sigma` so this scale is explicit in the returned list.

**Prior covariance:** scalar `pwt`,
$\Sigma = \frac{1-\mathrm{pwt}}{\mathrm{pwt}}\, V_0$ (equivalently
$\Sigma^{-1} = \frac{\mathrm{pwt}}{1-\mathrm{pwt}}\, P(\beta^{\ast})$)


This `Sigma` is what `Prior_Setup()` returns by default on the coefficient
scale. For Gaussian fits, the returned dispersion-free matrix is

\[
\Sigma_0 = \Sigma / d,
\]
so
\[
\Sigma_0^{-1}
=
d\,\Sigma^{-1}
=
d\,\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^{\ast}).
\]
Using $P(\beta^{\ast})=\frac{1}{d}X^{\mathsf T}WX$ in weighted Gaussian
regression gives
\[
\Sigma_0^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,X^{\mathsf T}WX,
\]
which is independent of $d$; this is the default `Sigma_0` returned by `Prior_Setup()`.

#### 2.3.2 Posterior mean and Variance under `dNormal()`

When default settings are used, the Gaussian posterior means reduce to simple
weighted averages of the fitted coefficient vector $\beta^\ast$ and prior mean
$\mu$.

**`dNormal()` (Gaussian, coefficient-scale covariance `Sigma`).**
For Gaussian likelihood precision $P(\beta^\ast)$ and prior precision
$\Sigma^{-1}$,
\[
E(\beta\mid y)
=
\bigl(P(\beta^\ast)+\Sigma^{-1}\bigr)^{-1}
\Bigl(P(\beta^\ast)\beta^\ast+\Sigma^{-1}\mu\Bigr).
\]
With the default scalar `pwt`,
\[
\Sigma^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^\ast),
\]
so
\[
\begin{aligned}
E(\beta\mid y)
&=
\left(P(\beta^\ast)+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast) \right)^{-1}
\left(P(\beta^\ast)\beta^\ast+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast)\mu\right) \\
&=
\left(\frac{1-\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast)+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast) \right)^{-1}
\left(\frac{1-\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast)\beta^\ast+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}P(\beta^\ast)\mu\right) \\
&=
\left(\frac{1}{1-\mathrm{pwt}}P(\beta^\ast)\right)^{-1}
\left(\frac{1}{1-\mathrm{pwt}}P(\beta^\ast)\bigl((1-\mathrm{pwt})\beta^\ast+\mathrm{pwt}\mu\bigr)\right) \\
&=
(1-\mathrm{pwt})\beta^\ast+\mathrm{pwt}\,\mu.
\end{aligned}
\]
Thus the posterior mean is a convex combination of the likelihood estimate
$\beta^\ast$ and prior mean $\mu$: larger `pwt` gives more pull toward $\mu$.
In the limit as $\mathrm{pwt}\to 0$, it approaches $\beta^\ast$.



The posterior covariance is
\[
\mathrm{Var}(\beta\mid y)
=
\bigl(P(\beta^\ast)+\Sigma^{-1}\bigr)^{-1}.
\]
With the default scalar `pwt`,
\[
\Sigma^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^\ast),
\]
so
\[
\begin{aligned}
\mathrm{Var}(\beta\mid y)
&=
\left(P(\beta^\ast)+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^\ast)\right)^{-1} \\
&=
\left(\frac{1-\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^\ast)+\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,P(\beta^\ast)\right)^{-1} \\
&=
\left(\frac{1}{1-\mathrm{pwt}}\,P(\beta^\ast)\right)^{-1} \\
&=
(1-\mathrm{pwt})\,P(\beta^\ast)^{-1}.
\end{aligned}
\]

Thus the posterior covariance is the likelihood-based covariance
$P(\beta^\ast)^{-1}$ shrunk by the factor $1-\mathrm{pwt}$: larger `pwt`
(stronger prior pull) gives tighter posterior uncertainty. In the limit as
$\mathrm{pwt}\to 0$, it approaches the likelihood-based covariance.

#### 2.3.3 Marginal posterior mean under `dNormal_Gamma()`

**`dNormal_Gamma()` (Gaussian conjugate Normal--Gamma, using `Sigma_0`).**
The **marginal** posterior mean is
\[
E(\beta\mid y)
=
E_{\tau\mid y}\!\left[E(\beta\mid \tau,y)\right].
\]
For fixed $\tau$,
\[
E(\beta\mid \tau,y)
=
\bigl(\tau X^{\mathsf T}W_{\mathrm{obs}}X+\tau\Sigma_0^{-1}\bigr)^{-1}
\Bigl(\tau X^{\mathsf T}W_{\mathrm{obs}}X\,\beta^\ast+\tau\Sigma_0^{-1}\mu\Bigr).
\]
Under the default scalar `pwt` calibration for `Sigma_0`,
\[
\Sigma_0^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,X^{\mathsf T}W_{\mathrm{obs}}X,
\]
so
\[
\begin{aligned}
E(\beta\mid \tau,y)
&=
\left(\tau X^{\mathsf T}W_{\mathrm{obs}}X+\tau\frac{\mathrm{pwt}}{1-\mathrm{pwt}}X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
\left(\tau X^{\mathsf T}W_{\mathrm{obs}}X\,\beta^\ast+\tau\frac{\mathrm{pwt}}{1-\mathrm{pwt}}X^{\mathsf T}W_{\mathrm{obs}}X\,\mu\right) \\
&=
\left(\frac{\tau}{1-\mathrm{pwt}}X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
\left(\frac{\tau}{1-\mathrm{pwt}}X^{\mathsf T}W_{\mathrm{obs}}X\bigl((1-\mathrm{pwt})\beta^\ast+\mathrm{pwt}\mu\bigr)\right) \\
&=
(1-\mathrm{pwt})\beta^\ast+\mathrm{pwt}\,\mu.
\end{aligned}
\]
Because this expression is free of $\tau$, averaging over $\tau\mid y$ gives
\[
E(\beta\mid y)=(1-\mathrm{pwt})\beta^\ast+\mathrm{pwt}\,\mu.
\]
Thus the marginal posterior mean has the same weighted-average interpretation:
larger `pwt` gives more pull toward $\mu$, and as $\mathrm{pwt}\to 0$ it
approaches $\beta^\ast$.
For general non-Gaussian GLMs these equalities are not exact in finite samples,
because the likelihood is not exactly quadratic in $\beta$; however, the same
weighted-average form is often a good approximation when the likelihood is
close to multivariate normal, as typically occurs in large samples.

### 2.4 Vector `pwt` and optional `sd`

**Vector `pwt`:** same Hadamard construction as above; correlations in $V_0$ are
preserved, variances scaled per coordinate.

**`sd`:** $\mathrm{pwt}_j = (V_0)_{jj}/\bigl((V_0)_{jj}+\mathrm{sd}_j^2\bigr)$;
vector `pwt` is not overwritten from scalar `n_prior`. **Gaussian** fits may
require scalar `n_prior` in addition (Section 3).

## 3. Default Priors for Dispersion, Shape, and Rate Parameters

This section develops the Gaussian prior families used when the dispersion
parameter is unknown. The goal is to show how `Prior_Setup()` constructs the
Gamma prior on the residual precision \(\tau = 1/\phi\), how the Normal block
interacts with the likelihood, and how the resulting posterior hyperparameters
arise.

### 3.1 Posterior pieces: contribution from likelihood + Normal block

We begin with the conjugate Normal–Gamma specification
\[
\beta \mid \tau \sim N\!\left(\mu,\; (\tau \Sigma_0)^{-1}\right),
\qquad
\tau \sim \Gamma(a_0, b_0),
\]
where \(\Sigma_0\) is the **dispersion‑free** prior covariance matrix.

For the weighted Gaussian likelihood,
\[
y \mid \beta,\tau \sim N\!\left(X\beta,\; \tau^{-1} W_{\mathrm{obs}}^{-1}\right),
\]
the Normal block and likelihood combine through:

- the **coefficient precision update**  
  \[
  X^{\mathsf T}W_{\mathrm{obs}}X \quad\text{and}\quad \Sigma_0^{-1},
  \]
- and the **marginal quadratic term**
  \[
  S_{\mathrm{marg}}
  =
  \mathrm{RSS}_w
  +
  (\hat\beta - \mu)^{\mathsf T}
  \left(
    \Sigma_0 + (X^{\mathsf T}W_{\mathrm{obs}}X)^{-1}
  \right)^{-1}
  (\hat\beta - \mu),
  \]
  where \(\hat\beta\) is the weighted least‑squares estimator and
  \(\mathrm{RSS}_w\) is the weighted residual sum of squares.

Integrating out \(\beta\) in the Normal–Gamma algebra adds
\[
\frac{n_w}{2}
\]
to the Gamma **shape** parameter (note: this parameterization does *not* add
\(p/2\)). Thus the posterior hyperparameters are
\[
a_n = a_0 + \frac{n_w}{2},
\qquad
b_n = b_0 + \frac{1}{2} S_{\mathrm{marg}},
\]
with \(n_w\) the effective sample size and \(p = \mathrm{ncol}(X)\).

---

### 3.2 Prior-strength parameterization from `pwt`

The scalar prior‑weight `pwt` is mapped to an **effective prior sample size**
\[
n_{\mathrm{prior}}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\, n_w,
\qquad\text{equivalently}\qquad
\mathrm{pwt}
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}} + n_w}.
\]

Interpretation:

- `pwt` controls how strongly the prior mean \(\mu\) influences the posterior,  
- `n_prior` is the number of “pseudo‑observations” implied by the prior,  
- and as `pwt → 0`, the prior becomes negligible and the posterior becomes
  likelihood‑dominated.

The dispersion‑free covariance used in `dNormal_Gamma()` is
\[
\Sigma_0
=
\frac{1-\mathrm{pwt}}{\mathrm{pwt}}
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
so that
\[
\Sigma_0^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}
\,X^{\mathsf T}W_{\mathrm{obs}}X.
\]

Substituting this into the expression for \(S_{\mathrm{marg}}\) yields
\[
\begin{aligned}
S_{\mathrm{marg}}
&=
\mathrm{RSS}_w
+
(\hat\beta - \mu)^{\mathsf T}
\left(
\frac{1-\mathrm{pwt}}{\mathrm{pwt}}
(X^{\mathsf T}W_{\mathrm{obs}}X)^{-1}
+
(X^{\mathsf T}W_{\mathrm{obs}}X)^{-1}
\right)^{-1}
(\hat\beta - \mu) \\
&=
\mathrm{RSS}_w
+
\mathrm{pwt}\,
(\hat\beta - \mu)^{\mathsf T}
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)
(\hat\beta - \mu).
\end{aligned}
\]

Thus under scalar `pwt`, the prior‑mean penalty in \(S_{\mathrm{marg}}\) is scaled
**directly** by `pwt`. This is the key link between the Normal block and the
Gamma update for \(\tau\).



### 3.3 Gaussian prior-family calibration and parameter mapping

This section explains how the outputs of `Prior_Setup()` map into the Gaussian
prior families and how a single calibration—based on `pwt`,
\(n_{\mathrm{prior}}\), and the Zellner form of \(\Sigma_0\)—governs all of them.

We proceed in four parts:

1. Default calibration of the Gamma prior on \(\tau\) from `n_prior`.  
2. Conjugate Normal–Gamma posterior (Theorem 1).  
3. Weak‑prior limit as \(n_{\mathrm{prior}}\to 0^{+}\) (Theorem 2).  
4. Independent Normal–Gamma analogue (Theorem 3).

A final subsection states the unified weak‑limit theorem.

---

#### 3.3.1 Default calibration and posterior Gamma shape/rate

Let \(n_w=\sum_i w_i\) be the effective sample size (`n_effective`).  
For scalar `pwt`, `Prior_Setup()` defines the **effective prior sample size**
\[
n_{\mathrm{prior}}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,n_w,
\qquad
\mathrm{pwt}
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}.
\]

Under the shape–rate parameterization \(\Gamma(a_0,b_0)\) with density
\(\propto \tau^{a_0-1}e^{-b_0\tau}\), the default prior on the residual precision
\(\tau\) is
\[
a_0=\frac{n_{\mathrm{prior}}+1}{2},
\qquad
b_0
=
\frac{1}{2}\,(n_{\mathrm{prior}}+p-1)\,
\frac{S_{\mathrm{marg}}}{n_w-p},
\]
where \(S_{\mathrm{marg}}\) is the marginal quadratic term from Section 3.1 and
\(n_w>p\) is required for propriety.

The posterior hyperparameters and induced moments follow from this calibration
and are summarized in Theorem 1.



#### 3.3.2 Conjugate Normal–Gamma posterior (dNormal_Gamma())

##### **Theorem 1 (Conjugate posterior under the default `dNormal_Gamma()` calibration)**


Assume:

1. \(X^{\mathsf T}W_{\mathrm{obs}}X\) is positive definite,  
2. \(n_w>p\),  
3. \(\mathrm{RSS}_w>0\).

Let the prior be
\[
\beta\mid\tau\sim N(\mu,\tau^{-1}\Sigma_0),
\qquad
\tau\sim\Gamma(a_0,b_0),
\]
with
\[
\Sigma_0
=
\frac{1-\mathrm{pwt}}{\mathrm{pwt}}
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
=
\frac{n_w}{n_{\mathrm{prior}}}
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]

Then the posterior is again Normal–Gamma with the following hyperparameters.

---

#### **(i) Posterior mean of \(\beta\)**

\[
\mu_{\mathrm{post}}
=
\mathrm{pwt}\,\mu+(1-\mathrm{pwt})\,\hat\beta
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}\,\mu
+
\frac{n_w}{n_{\mathrm{prior}}+n_w}\,\hat\beta.
\]

---

#### **(ii) Posterior dispersion‑free covariance**

\[
\Sigma_{0,\mathrm{post}}
=
(\Sigma_0^{-1}+X^{\mathsf T}W_{\mathrm{obs}}X)^{-1}
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}\,\Sigma_0
=
\frac{n_w}{n_{\mathrm{prior}}+n_w}
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]

For general \(\Sigma_0\), use  
\(\Sigma_{0,\mathrm{post}}=(\Sigma_0^{-1}+X^{\mathsf T}W_{\mathrm{obs}}X)^{-1}\).

---

#### **(iii) Posterior Gamma shape**

\[
a_n
=
a_0+\frac{n_w}{2}
=
\frac{n_{\mathrm{prior}}+n_w+1}{2}.
\]

---

#### **(iv) Posterior Gamma rate**

\[
b_n
=
b_0+\frac{1}{2}S_{\mathrm{marg}}
=
\frac{1}{2}\,(n_{\mathrm{prior}}+n_w-1)\,
\frac{S_{\mathrm{marg}}}{n_w-p}.
\]

---

#### **(v) Marginal posterior mean of \(\beta\)**

\[
\mathbb{E}[\beta\mid y]=\mu_{\mathrm{post}}.
\]

---

#### **(vi) Posterior expectation of \(\sigma^2=1/\tau\)**

For \(a_n>1\),
\[
\mathbb{E}[\sigma^2\mid y]
=
\frac{b_n}{a_n-1}
=
\frac{S_{\mathrm{marg}}}{n_w-p}.
\]

---

#### **(vii) Marginal posterior covariance of \(\beta\)**

Let  
\[
V_n=\Sigma_{0,\mathrm{post}}
=
(\Sigma_0^{-1}+X^{\mathsf T}W_{\mathrm{obs}}X)^{-1}.
\]

Then
\[
\mathrm{Cov}(\beta\mid y)
=
\mathbb{E}[\sigma^2\mid y]\,V_n
=
\frac{S_{\mathrm{marg}}}{n_w-p}\,
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]


*Proof.* See Appendix B.
---

#### **Interpretation**

- `pwt` controls the **pull toward the prior mean** in (i).  
- `pwt` also controls the **shrinkage of the covariance** in (vii) via  
  \(n_w/(n_w+n_{\mathrm{prior}})\).  
- The denominator \(n_w-p\) reflects the **residual degrees of freedom** in the weighted Gaussian model.

Together, these determine how prior strength interacts with sample size and model dimension.


### **Theorem 2 (Weak‑prior limit of the `dNormal_Gamma()` posterior)**

Assume the same identifiability conditions as in Theorem 1:

1. \(X^{\mathsf T}W_{\mathrm{obs}}X\) is positive definite,  
2. \(n_w > p\),  
3. \(\mathrm{RSS}_w > 0\).

Under the default calibration of Theorem 1, let  
\[
n_{\mathrm{prior}} \to 0^{+}
\qquad\text{equivalently}\qquad
\mathrm{pwt} \to 0^{+},
\quad
n_{\mathrm{prior}}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,n_w.
\]

Then \(S_{\mathrm{marg}} \to \mathrm{RSS}_w\), and the conjugate
`dNormal_Gamma()` posterior converges weakly to a **Normal–Gamma** law
\(\Pi_{0}(\cdot\mid y)\) on \((\beta,\tau)\).  
The limiting hyperparameters are the limits of the posterior quantities in
Theorem 1 as \(n_{\mathrm{prior}}\to 0^{+}\).  
(These are *not* the prior hyperparameters \(a_0,b_0\).)

---

#### **(i) Limiting posterior mean of \(\beta\)**

\[
\mu_{\Pi_{0}}
=
\lim_{n_{\mathrm{prior}}\to 0^{+}}
\mu_{\mathrm{post}}
=
\hat\beta.
\]

---

#### **(ii) Limiting dispersion‑free covariance**

\[
\Sigma_{0,\Pi_{0}}
=
\lim_{n_{\mathrm{prior}}\to 0^{+}}
\Sigma_{0,\mathrm{post}}
=
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]

---

#### **(iii) Limiting Gamma shape**

\[
a_{\Pi_{0}}
=
\lim_{n_{\mathrm{prior}}\to 0^{+}}
a_n
=
\frac{n_w+1}{2}.
\]

---

#### **(iv) Limiting Gamma rate**

\[
b_{\Pi_{0}}
=
\lim_{n_{\mathrm{prior}}\to 0^{+}}
b_n
=
\frac{1}{2}\,
\frac{n_w-1}{n_w-p}\,
\mathrm{RSS}_w.
\]

---

#### **(v) Limiting marginal mean of \(\beta\)**

\[
\mathbb{E}_{\Pi_{0}}[\beta\mid y]
=
\mu_{\Pi_{0}}
=
\hat\beta.
\]

---

#### **(vi) Limiting expectation of \(\sigma^2 = 1/\tau\)**

For \(\tau\mid y \sim \Gamma(a_{\Pi_{0}},b_{\Pi_{0}})\),
\[
\mathbb{E}_{\Pi_{0}}[\sigma^2\mid y]
=
\frac{b_{\Pi_{0}}}{a_{\Pi_{0}}-1}
=
\frac{\mathrm{RSS}_w}{n_w-p},
\]
the classical weighted residual‑variance estimator.

---

#### **(vii) Limiting marginal covariance of \(\beta\)**

\[
\mathrm{Cov}_{\Pi_{0}}(\beta\mid y)
=
\mathbb{E}_{\Pi_{0}}[\sigma^2\mid y]\,
\Sigma_{0,\Pi_{0}}
=
\frac{\mathrm{RSS}_w}{n_w-p}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
matching the usual weighted least‑squares covariance.

---

### **Interpretation**

The limit \(\Pi_{0}\) is the **weak‑prior Normal–Gamma law** obtained when the
prior contributes no pseudo‑information.  
It has:

- location \(\hat\beta\),  
- geometry \((X^{\mathsf T}W_{\mathrm{obs}}X)^{-1}\),  
- and Gamma precision determined entirely by the data.

The independent Normal–Gamma posterior (Theorem 3) converges to this **same**
\(\Pi_{0}\) under the same assumptions; only the finite‑\(n_{\mathrm{prior}}\)
joint density differs.


*Proof of Theorem 2.*

By Theorem 1, for each \(n_{\mathrm{prior}}>0\) the dNormal\_Gamma posterior is Normal–Gamma with
hyperparameters
\[
\mu_{\mathrm{post}}(n_{\mathrm{prior}}),\quad
\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}}),\quad
a_n(n_{\mathrm{prior}}),\quad
b_n(n_{\mathrm{prior}}),
\]
given explicitly by
\[
\mu_{\mathrm{post}}(n_{\mathrm{prior}})
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}\,\mu
+
\frac{n_w}{n_{\mathrm{prior}}+n_w}\,\hat\beta,
\]
\[
\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}})
=
\frac{n_{\mathrm{prior}}+n_w}{n_w}\,G^{-1},
\]
\[
a_n(n_{\mathrm{prior}})
=
\frac{n_{\mathrm{prior}}+n_w+1}{2},
\qquad
b_n(n_{\mathrm{prior}})
=
\frac{n_{\mathrm{prior}}+n_w-1}{2}\,
\frac{\mathrm{Smarg}}{n_w-p}.
\]

As \(n_{\mathrm{prior}}\to 0^+\), each of these converges to a finite, valid Normal–Gamma parameter:
\[
\mu_{\mathrm{post}}(n_{\mathrm{prior}})\to\hat\beta,\quad
\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}})\to G^{-1},
\]
\[
a_n(n_{\mathrm{prior}})\to\frac{n_w+1}{2}>0,\quad
b_n(n_{\mathrm{prior}})\to\frac{n_w-1}{2}\,\frac{\mathrm{RSS}_w}{n_w-p}>0,
\]
using \(\mathrm{Smarg}\to\mathrm{RSS}_w\) as \(n_{\mathrm{prior}}\to 0^+\).

The Normal–Gamma family is closed under weak limits when its parameters converge in this way, so the
posteriors \(\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}\) converge weakly to the Normal–Gamma law \(\Pi_0\) with these
limiting hyperparameters. The stated formulas for the limiting mean, covariance, and variance of \(\beta\) and
\(\sigma^2=1/\tau\) follow by plugging the limits into the standard Normal–Gamma moment expressions.
\(\square\)



### 3.3.3 Posterior covariance under `dNormal()` with default dispersion

The `dNormal()` prior fixes the residual variance \(\sigma^2\) at a calibrated
value rather than integrating over \(\tau\) as in the Normal–Gamma model.  
This section shows how the default dispersion is chosen and how the resulting
posterior covariance matches the weak‑prior limit of `dNormal_Gamma()`.

---

#### Covariance under fixed \(\sigma^2\)

From Section 2.3.2, under scalar `pwt`,
\[
\mathrm{Var}(\beta\mid y,\sigma^2)
=
(1-\mathrm{pwt})\,P(\beta^\ast)^{-1}.
\]

For weighted Gaussian regression,
\[
P(\beta^\ast)
=
\sigma^{-2}X^{\mathsf T}W_{\mathrm{obs}}X,
\]
so
\[
\mathrm{Var}(\beta\mid y,\sigma^2)
=
(1-\mathrm{pwt})\,\sigma^2
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]

Using  
\[
n_{\mathrm{prior}}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}},
\qquad
1-\mathrm{pwt}
=
\frac{n_w}{n_w+n_{\mathrm{prior}}},
\]
this becomes
\[
\mathrm{Var}(\beta\mid y,\sigma^2)
=
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,
\sigma^2
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]

---

#### Default dispersion

To choose a default fixed value of \(\sigma^2\), `Prior_Setup()` uses the
posterior mean from the Normal–Gamma model (Theorem 1 (vi)):
\[
\mathrm{dispersion}_{\mathrm{default}}
=
\frac{S_{\mathrm{marg}}}{n_w-p}.
\]

This matches the classical residual degrees‑of‑freedom adjustment.

Substituting this into the covariance expression gives
\[
\mathrm{Var}(\beta\mid y,\mathrm{dispersion}_{\mathrm{default}})
=
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,
\frac{S_{\mathrm{marg}}}{n_w-p}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]

---

#### Calibrated prior covariance returned by `Prior_Setup()`

With the same default dispersion, `Prior_Setup()` returns the coefficient‑scale
prior covariance
\[
\Sigma_{\mathrm{calibrated}}
=
\frac{n_w}{n_{\mathrm{prior}}}\,
\mathrm{dispersion}_{\mathrm{default}}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
which is the matrix used by `dNormal()`.

This matches the Normal–Gamma expression in Section 3.3.2, ensuring that the
fixed‑dispersion and conjugate models share the same calibration.

---

#### Weak‑prior limit

As \(\mathrm{pwt}\to 0\) (equivalently \(n_{\mathrm{prior}}\to 0^{+}\)),
\[
\mathrm{Var}(\beta\mid y)
\;\longrightarrow\;
\frac{\mathrm{RSS}_w}{n_w-p}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
the classical weighted least‑squares covariance.

Thus `dNormal()` with default dispersion has the **same weak‑prior limit** as
`dNormal_Gamma()`, and the returned `shape`, `rate`, `dispersion`, and
coefficient‑scale covariance remain internally consistent under the package
calibration.


### 3.3.4 Independent Normal–Gamma Prior

The independent Normal–Gamma (ING) prior replaces the conjugate covariance structure 
\(\tau^{-1}\Sigma_0\) with a fixed coefficient-scale covariance \(\Sigma\), while using a Gamma prior on 
\(\tau\) whose shape parameter differs from the conjugate Normal–Gamma case by \(p/2\).

The default call is

```text
dIndependent_Normal_Gamma(ps$mu, Sigma = ps$Sigma, shape = ps$shape_ING, rate = ps$rate)
```

Let \(p = \mathrm{ncol}(X)\), and let \(a_0, b_0, S_{\mathrm{marg}}\) be as in Sections 3.3.1–3.3.2.

- Prior mean:
  \[
  \mu = \texttt{ps\$mu}.
  \]

- Coefficient-scale covariance:
  \[
  \Sigma
  =
  \frac{n_w}{n_{\mathrm{prior}}}\,
  \frac{S_{\mathrm{marg}}}{n_w - p}\,
  (X^{\mathsf T}W_{\mathrm{obs}}X)^{-1}.
  \]

- ING Gamma shape:
  \[
  \mathrm{shape}_{\mathrm{ING}}
  =
  a_0 + \frac{p}{2}
  =
  \frac{n_{\mathrm{prior}} + p + 1}{2}.
  \]

- Gamma rate:
  \[
  \texttt{rate}
  =
  b_0
  =
  \frac{1}{2}(n_{\mathrm{prior}} + p - 1)\,
  \frac{S_{\mathrm{marg}}}{n_w - p}.
  \]

---

### Theorem 3 (Weak-prior limit of the Independent Normal–Gamma posterior)

Assume:

1. \(X^{\mathsf T}W_{\mathrm{obs}}X\) is positive definite  
2. \(n_w > p\)  
3. \(\mathrm{RSS}_w > 0\)

For each \(n_{\mathrm{prior}} > 0\), let  
\[
\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\cdot \mid y)
\]
denote the posterior under the ING prior above.  
Let \(\Pi_0(\cdot \mid y)\) be the Normal–Gamma law from Theorem 2 with hyperparameters

\[
\mu_{\Pi_0} = \hat\beta, \qquad
\Sigma_{0,\Pi_0} = (X^{\mathsf T}W_{\mathrm{obs}}X)^{-1},
\]
\[
a_{\Pi_0} = \frac{n_w + 1}{2}, \qquad
b_{\Pi_0} = \frac{1}{2}\frac{n_w - 1}{n_w - p}\,\mathrm{RSS}_w.
\]

Then, as \(n_{\mathrm{prior}} \to 0^{+}\) (equivalently \(\mathrm{pwt} \to 0^{+}\)),

\[
\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\cdot \mid y)
\;\Rightarrow\;
\Pi_0(\cdot \mid y)
\]

in distribution on \(\mathbb{R}^p \times (0,\infty)\).  
Moreover, the posterior moments converge:

1. Coefficient mean:
   \[
   \mathbb{E}_{\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}}[\beta \mid y]
   \longrightarrow
   \hat\beta.
   \]

2. Residual variance:
   \[
   \mathbb{E}_{\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}}[\sigma^2 \mid y]
   \longrightarrow
   \frac{\mathrm{RSS}_w}{n_w - p}.
   \]

3. Coefficient covariance:
   \[
   \mathrm{Cov}_{\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}}(\beta \mid y)
   \longrightarrow
   \frac{\mathrm{RSS}_w}{n_w - p}\,
   (X^{\mathsf T}W_{\mathrm{obs}}X)^{-1}.
   \]

Thus the ING posterior has the same weak-prior limit as the conjugate Normal–Gamma posterior, even though its finite-\(n_{\mathrm{prior}}\) form is not conjugate and its Gamma shape parameter differs by \(p/2\).


*Proof of Theorem 3.*

Fix \(y,X,W_{\mathrm{obs}}\) satisfying Assumptions 1–3. For each \(n_{\mathrm{prior}}>0\), let
\(\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}\) and \(\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}\) denote, respectively, the NG and ING posteriors on
\((\beta,\tau)\).

By Theorem 2, the NG posteriors converge weakly to the limiting Normal–Gamma law \(\Pi_0\):
\[
\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}} \Rightarrow \Pi_0
\quad\text{as }n_{\mathrm{prior}}\to 0^+.
\]

From the posterior ratio identity in A.2 and Lemma B, we have, for each \(n_{\mathrm{prior}}>0\),
\[
\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\mathrm{d}\beta,\mathrm{d}\tau)
=
R_{n_{\mathrm{prior}}}(\beta,\tau)\,
\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\mathrm{d}\beta,\mathrm{d}\tau),
\]
with
\[
R_{n_{\mathrm{prior}}}(\beta,\tau)\to 1
\quad\text{for each fixed }(\beta,\tau),
\]
and a measurable envelope \(M(\beta,\tau)\) such that
\[
\sup_{0<n_{\mathrm{prior}}<\delta}
|R_{n_{\mathrm{prior}}}(\beta,\tau)|
\le M(\beta,\tau),
\qquad
\int M(\beta,\tau)\,\Pi_0(\mathrm{d}\beta,\mathrm{d}\tau)<\infty.
\]

Let \(f\colon\mathbb{R}^p\times(0,\infty)\to\mathbb{R}\) be bounded and continuous. Then
\[
\int f\,\mathrm{d}\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
=
\int f(\beta,\tau)\,R_{n_{\mathrm{prior}}}(\beta,\tau)\,
\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\mathrm{d}\beta,\mathrm{d}\tau).
\]
By Lemma A, the NG path has uniform moment bounds, and together with Claim B.2 this implies that
\(|f\,R_{n_{\mathrm{prior}}}|\) is dominated by an integrable envelope under \(\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}\) for
\(0<n_{\mathrm{prior}}<\delta\). Using the pointwise convergence \(R_{n_{\mathrm{prior}}}\to 1\) and the weak convergence
\(\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}\Rightarrow\Pi_0\), we obtain, by dominated convergence,
\[
\int f\,\mathrm{d}\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
\longrightarrow
\int f\,\mathrm{d}\Pi_0.
\]
Thus \(\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}\Rightarrow\Pi_0\) as \(n_{\mathrm{prior}}\to 0^+\), proving the distributional
convergence.

For the moment statements, take \(f(\beta,\tau)=\beta_j\), \(f(\beta,\tau)=\tau^{-1}\), and
\(f(\beta,\tau)=(\beta-\mathbb{E}_{\Pi_0}[\beta])(\beta-\mathbb{E}_{\Pi_0}[\beta])^\top\) componentwise. Lemma A
provides uniform integrability of the corresponding NG moments, and the same envelope \(M\) from Claim B.2
transfers this to the ING path via the ratio representation. Hence dominated convergence applies to these
(unbounded) test functions as well, yielding
\[
\mathbb{E}_{\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}}[\beta]
\to
\mathbb{E}_{\Pi_0}[\beta]=\hat\beta,
\quad
\mathbb{E}_{\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}}[\tau^{-1}]
\to
\mathbb{E}_{\Pi_0}[\tau^{-1}]
=
\frac{\mathrm{RSS}_w}{n_w-p},
\]
and
\[
\mathrm{Cov}_{\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}}(\beta\mid y)
\to
\mathrm{Cov}_{\Pi_0}(\beta\mid y)
=
\frac{\mathrm{RSS}_w}{n_w-p}\,G^{-1}.
\]

These limits match the expressions stated in Theorem 3, so the ING posterior has the same weak–prior limit
\(\Pi_0\) as the conjugate Normal–Gamma posterior, with convergence of the first two moments. \(\square\)



## Appendix A: Technical Ingredients for the ING Weak‑Prior Limit

This appendix collects the analytical components required to establish Theorem 3.  
Theorems 1 and 2 follow directly from conjugate Normal–Gamma algebra and the Zellner‑type
calibration; only the Independent Normal–Gamma (ING) case requires additional work.  
The purpose of this appendix is therefore to isolate the technical machinery needed to show that the
ING posterior converges to the same weak‑prior limit \(\Pi_0\) as the conjugate Normal–Gamma posterior.

The argument proceeds through five steps:

1. A common Gaussian likelihood representation  
2. A ratio representation comparing ING and NG posteriors  
3. Uniform moment bounds for the NG path (Lemma A)  
4. Ratio convergence and domination (Lemma B)  
5. Weak convergence and moment convergence for the ING posterior

Each subsection states the required intermediate results and provides the structural components of the
proof, while detailed algebraic derivations are deferred to the appropriate claims and lemmas.

### A.1 Common Gaussian Setup

Let

- \(G = X^{\mathsf T}W_{\mathrm{obs}}X\),
- \(\hat\beta\) the weighted least‑squares estimator,
- \(\mathrm{RSS}_w\) the weighted residual sum of squares,
- \(\mathrm{RSS}_w(\beta) = \mathrm{RSS}_w + (\beta - \hat\beta)^{\mathsf T}G(\beta - \hat\beta)\).

The weighted Gaussian likelihood can be written as

\[
L(y \mid \beta,\tau)
\propto
\tau^{n_w/2}
\exp\!\left(
-\frac{\tau}{2}\,\mathrm{RSS}_w(\beta)
\right).
\]

This representation is shared by both the NG and ING posterior paths.

---

### A.2 Posterior Ratio Representation

To compare the ING and NG posterior paths, we first record their correct prior kernels.

#### NG prior (Theorem 1, §3.3.2)

For each \(n_{\mathrm{prior}} > 0\),
\[
\beta \mid \tau \sim N\!\left(\mu,\;\tau^{-1}\Sigma_0\right),
\qquad
\tau \sim \Gamma\!\left(a_0(n_{\mathrm{prior}}),\, b_0(n_{\mathrm{prior}})\right),
\]
where the dispersion–free Zellner matrix is
\[
\Sigma_0
=
\frac{1-pwt}{pwt}\,(X^\top W_{\mathrm{obs}}X)^{-1}.
\]

Thus the NG prior kernel is
\[
\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\beta,\tau)
\propto
\tau^{a_0(n_{\mathrm{prior}})+p/2-1}
\exp\!\left(
-b_0(n_{\mathrm{prior}})\tau
-\frac{\tau}{2}(\beta-\mu)^\top\Sigma_0^{-1}(\beta-\mu)
\right).
\]

#### ING prior (§3.3.4)

The ING prior uses a fixed coefficient–scale covariance and a Gamma shape shifted by \(p/2\):
\[
\beta \mid n_{\mathrm{prior}} \sim N\!\left(\mu,\;\Sigma(n_{\mathrm{prior}})\right),
\qquad
\tau \sim \Gamma\!\left(a_0(n_{\mathrm{prior}})+\tfrac{p}{2},\; b_0(n_{\mathrm{prior}})\right),
\]
with \(\beta\) and \(\tau\) independent, and
\[
\Sigma(n_{\mathrm{prior}})
=
\frac{n_w}{n_{\mathrm{prior}}}\,
\frac{\mathrm{Smarg}}{n_w - p}\,
(X^\top W_{\mathrm{obs}}X)^{-1}.
\]

Thus the ING prior kernel is
\[
\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\beta,\tau)
\propto
\tau^{a_0(n_{\mathrm{prior}})+p/2-1}
\exp\!\left(
-b_0(n_{\mathrm{prior}})\tau
\right)
\exp\!\left(
-\frac{1}{2}(\beta-\mu)^\top\Sigma(n_{\mathrm{prior}})^{-1}(\beta-\mu)
\right).
\]

#### Ratio of prior kernels

Define
\[
R_{n_{\mathrm{prior}}}(\beta,\tau)
=
\frac{
\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\beta,\tau)
}{
\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\beta,\tau)
}.
\]

Because the ING Gamma shape equals the NG Gamma shape plus \(p/2\), the \(\tau\)-powers match and cancel.
The ratio therefore reduces to
\[
R_{n_{\mathrm{prior}}}(\beta,\tau)
=
C_{n_{\mathrm{prior}}}\,
\exp\!\left(
-\frac{1}{2}(\beta-\mu)^\top
\bigl[\Sigma(n_{\mathrm{prior}})^{-1}
-\tau\,\Sigma_0^{-1}\bigr]
(\beta-\mu)
\right),
\]
where \(C_{n_{\mathrm{prior}}}\) absorbs all \(\tau\)-free constants.

#### Posterior ratio identity

The ING posterior is a reweighted NG posterior:
\[
\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\mathrm{d}\beta,\mathrm{d}\tau)
\propto
R_{n_{\mathrm{prior}}}(\beta,\tau)\,
\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\mathrm{d}\beta,\mathrm{d}\tau).
\]

This identity is the starting point for Lemma B and the ING weak‑prior limit.
---

### A.3 Lemma A: Uniform moment bounds for the NG path

#### Lemma A (Uniform moment bounds for the NG posterior)

Assume:

1. \(G := X^{\mathsf T}W_{\mathrm{obs}}X\) is positive definite  
2. \(n_w > p\)  
3. \(\mathrm{RSS}_w > 0\)

Fix \(X, W_{\mathrm{obs}}, y, \mu\), and hence \(\hat\beta\), \(\mathrm{RSS}_w\), \(S_{\mathrm{marg}}\), and \(G\).  
For each \(n_{\mathrm{prior}} > 0\), let \(\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}\) be the conjugate Normal–Gamma posterior
from Section 3.3.2, with hyperparameters

\[
\mu_{\mathrm{post}}(n_{\mathrm{prior}}),
\quad
\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}}),
\quad
a_n(n_{\mathrm{prior}}),
\quad
b_n(n_{\mathrm{prior}}).
\]

Then there exists \(\delta > 0\) and constants \(C_1, C_2 < \infty\) such that for all
\(0 < n_{\mathrm{prior}} < \delta\),

\[
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}
\bigl[\|\beta\|\bigr]
\le C_1,
\qquad
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}
\bigl[\|\beta\|^2\bigr]
\le C_2,
\]

and

\[
\sup_{0 < n_{\mathrm{prior}} < \delta}
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}[\tau]
< \infty,
\qquad
\sup_{0 < n_{\mathrm{prior}} < \delta}
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}[\tau^2]
< \infty.
\]

---

#### Claim A.1 (Continuity and compactness of NG hyperparameters)

Under Assumptions 1–3 of Theorem 3, the NG hyperparameters satisfy:

- \(n_{\mathrm{prior}} \mapsto \mu_{\mathrm{post}}(n_{\mathrm{prior}})\) is continuous on \((0,\infty)\) and
  \(\mu_{\mathrm{post}}(n_{\mathrm{prior}}) \to \hat\beta\) as \(n_{\mathrm{prior}} \to 0^{+}\).
- \(n_{\mathrm{prior}} \mapsto \Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}})\) is continuous on \((0,\infty)\) and
  \(\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}}) \to G^{-1}\) as \(n_{\mathrm{prior}} \to 0^{+}\).
- \(n_{\mathrm{prior}} \mapsto a_n(n_{\mathrm{prior}})\) and \(n_{\mathrm{prior}} \mapsto b_n(n_{\mathrm{prior}})\) are
  continuous on \((0,\infty)\) and converge to strictly positive limits as \(n_{\mathrm{prior}} \to 0^{+}\).

In particular, there exists \(\delta > 0\) such that for all \(0 < n_{\mathrm{prior}} < \delta\), the four
hyperparameters lie in compact subsets of their respective spaces.

*Proof of Claim A.1.*

By Theorem 1 and the prior setup, the NG hyperparameters can be written explicitly as functions of
\(n_{\mathrm{prior}} > 0\):

\[
\mu_{\mathrm{post}}(n_{\mathrm{prior}})
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}\,\mu
+
\frac{n_w}{n_{\mathrm{prior}}+n_w}\,\hat\beta,
\]

\[
\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}})
=
\frac{n_w}{n_{\mathrm{prior}}+n_w}\,G^{-1},
\]

\[
a_n(n_{\mathrm{prior}})
=
\frac{n_{\mathrm{prior}}+n_w+1}{2},
\qquad
b_n(n_{\mathrm{prior}})
=
\frac{1}{2}\bigl(n_{\mathrm{prior}}+n_w-1\bigr)\,
\frac{S_{\mathrm{marg}}}{n_w-p}.
\]

Here \(\mu, \hat\beta, G, S_{\mathrm{marg}}, n_w, p\) are fixed and do not depend on \(n_{\mathrm{prior}}\).

Each of these maps is a rational (in fact affine) function of \(n_{\mathrm{prior}}\) with denominator
\(n_{\mathrm{prior}}+n_w > 0\), so all four are continuous on \((0,\infty)\). Assumption 1
(\(G\) positive definite) ensures that \(G^{-1}\) exists and is finite, so
\(\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}})\) is well defined for all \(n_{\mathrm{prior}}>0\). Assumption 2
(\(n_w > p\)) implies \(n_w-p>0\), so the denominator in \(b_n(n_{\mathrm{prior}})\) is positive. Assumption 3
(\(\mathrm{RSS}_w>0\)) implies \(S_{\mathrm{marg}}>0\), so the rate \(b_n(n_{\mathrm{prior}})\) is strictly positive
for all \(n_{\mathrm{prior}}>0\).

Taking the limit \(n_{\mathrm{prior}} \to 0^{+}\) in the explicit formulas gives

\[
\mu_{\mathrm{post}}(n_{\mathrm{prior}}) \to \hat\beta,
\qquad
\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}}) \to G^{-1},
\]

\[
a_n(n_{\mathrm{prior}}) \to \frac{n_w+1}{2} > 0,
\qquad
b_n(n_{\mathrm{prior}}) \to \frac{1}{2}\frac{n_w-1}{n_w-p}\,S_{\mathrm{marg}} > 0,
\]

where the strict positivity of the limits of \(a_n\) and \(b_n\) uses Assumptions 2–3.

Since each map is continuous on \((0,\infty)\) and has a finite limit as \(n_{\mathrm{prior}} \to 0^{+}\), there
exists \(\delta > 0\) such that, for all \(0 < n_{\mathrm{prior}} < \delta\),

- \(\mu_{\mathrm{post}}(n_{\mathrm{prior}})\) lies in a compact subset of \(\mathbb{R}^p\),
- \(\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}})\) lies in a compact subset of the positive definite matrices,
- \(a_n(n_{\mathrm{prior}})\) and \(b_n(n_{\mathrm{prior}})\) lie in compact subsets of \((0,\infty)\).

This is exactly the continuity and compactness statement of Claim A.1.
\(\square\)


---


#### Proof of Lemma A

For each \(n_{\mathrm{prior}} > 0\), Theorem 1 gives

- \(\beta \mid \tau, y, n_{\mathrm{prior}} \sim N\bigl(\mu_{\mathrm{post}}(n_{\mathrm{prior}}), \tau^{-1}\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}})\bigr)\),  
- \(\tau \mid y, n_{\mathrm{prior}} \sim \Gamma\bigl(a_n(n_{\mathrm{prior}}), b_n(n_{\mathrm{prior}})\bigr)\).

By Claim A.1, there exists \(\delta > 0\) such that for all \(0 < n_{\mathrm{prior}} < \delta\),

- \(a_n(n_{\mathrm{prior}})\) and \(b_n(n_{\mathrm{prior}})\) lie in compact subsets of \((0,\infty)\),  
- \(\mu_{\mathrm{post}}(n_{\mathrm{prior}})\) lies in a compact subset of \(\mathbb{R}^p\),  
- \(\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}})\) lies in a compact subset of the positive definite matrices.

---

##### Bounds for \(\tau\)

For each \(n_{\mathrm{prior}}\), \(\tau \mid y, n_{\mathrm{prior}} \sim \Gamma(a_n(n_{\mathrm{prior}}), b_n(n_{\mathrm{prior}}))\), so

\[
\mathbb{E}[\tau \mid y, n_{\mathrm{prior}}]
=
\frac{a_n(n_{\mathrm{prior}})}{b_n(n_{\mathrm{prior}})},
\qquad
\mathbb{E}[\tau^2 \mid y, n_{\mathrm{prior}}]
=
\frac{a_n(n_{\mathrm{prior}})\bigl(a_n(n_{\mathrm{prior}})+1\bigr)}{b_n(n_{\mathrm{prior}})^2}.
\]

On \((0,\delta)\), both \(a_n(n_{\mathrm{prior}})\) and \(b_n(n_{\mathrm{prior}})\) stay in compact subsets of
\((0,\infty)\) by Claim A.1, so the maps

\[
n_{\mathrm{prior}} \mapsto \frac{a_n(n_{\mathrm{prior}})}{b_n(n_{\mathrm{prior}})},
\qquad
n_{\mathrm{prior}} \mapsto \frac{a_n(n_{\mathrm{prior}})\bigl(a_n(n_{\mathrm{prior}})+1\bigr)}{b_n(n_{\mathrm{prior}})^2}
\]

are continuous and bounded on \((0,\delta)\). Hence

\[
\sup_{0 < n_{\mathrm{prior}} < \delta}
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}[\tau]
< \infty,
\qquad
\sup_{0 < n_{\mathrm{prior}} < \delta}
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}[\tau^2]
< \infty.
\]

---

##### Bounds for \(\beta\)

The marginal distribution of \(\beta \mid y, n_{\mathrm{prior}}\) under \(\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}\) has

\[
\mathbb{E}[\beta \mid y, n_{\mathrm{prior}}]
=
\mu_{\mathrm{post}}(n_{\mathrm{prior}}),
\]

and

\[
\mathrm{Cov}(\beta \mid y, n_{\mathrm{prior}})
=
\mathbb{E}[\sigma^2 \mid y, n_{\mathrm{prior}}]\,
\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}}),
\]

where \(\sigma^2 = 1/\tau\) and, for \(a_n(n_{\mathrm{prior}}) > 1\),

\[
\mathbb{E}[\sigma^2 \mid y, n_{\mathrm{prior}}]
=
\frac{b_n(n_{\mathrm{prior}})}{a_n(n_{\mathrm{prior}})-1}.
\]

By Claim A.1, \(a_n(n_{\mathrm{prior}}) \to (n_w+1)/2 > 0\) as \(n_{\mathrm{prior}} \to 0^{+}\). Shrinking \(\delta\) if
necessary, we may assume \(a_n(n_{\mathrm{prior}}) > 1\) for all \(0 < n_{\mathrm{prior}} < \delta\). On this interval,
\(a_n(n_{\mathrm{prior}})\) and \(b_n(n_{\mathrm{prior}})\) lie in compact subsets of \((0,\infty)\), so

\[
n_{\mathrm{prior}} \mapsto \frac{b_n(n_{\mathrm{prior}})}{a_n(n_{\mathrm{prior}})-1}
\]

is continuous and bounded on \((0,\delta)\). By Claim A.1, \(\Sigma_{0,\mathrm{post}}(n_{\mathrm{prior}})\) lies in a
compact subset of the positive definite matrices, so its operator norm and trace are bounded on
\((0,\delta)\). Therefore

\[
\sup_{0 < n_{\mathrm{prior}} < \delta}
\mathrm{tr}\,\mathrm{Cov}(\beta \mid y, n_{\mathrm{prior}})
< \infty.
\]

Now

\[
\mathbb{E}\bigl[\|\beta\|^2 \mid y, n_{\mathrm{prior}}\bigr]
=
\bigl\|\mathbb{E}[\beta \mid y, n_{\mathrm{prior}}]\bigr\|^2
+
\mathrm{tr}\,\mathrm{Cov}(\beta \mid y, n_{\mathrm{prior}}).
\]

By Claim A.1, \(\mu_{\mathrm{post}}(n_{\mathrm{prior}})\) lies in a compact subset of \(\mathbb{R}^p\) for
\(0 < n_{\mathrm{prior}} < \delta\), so \(\|\mu_{\mathrm{post}}(n_{\mathrm{prior}})\|\) is bounded on \((0,\delta)\). Combined
with the bound on \(\mathrm{tr}\,\mathrm{Cov}(\beta \mid y, n_{\mathrm{prior}})\), this implies

\[
\sup_{0 < n_{\mathrm{prior}} < \delta}
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}
\bigl[\|\beta\|^2\bigr]
< \infty.
\]

Define

\[
C_2
:=
\sup_{0 < n_{\mathrm{prior}} < \delta}
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}
\bigl[\|\beta\|^2\bigr]
< \infty.
\]

Finally, by Cauchy–Schwarz,

\[
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}
\bigl[\|\beta\|\bigr]
\le
\Bigl(
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}
\bigl[\|\beta\|^2\bigr]
\Bigr)^{1/2}
\le
\sqrt{C_2}
=: C_1.
\]

This proves Lemma A.

### A.4 Lemma B: Ratio convergence and domination

Lemma B (Ratio convergence and domination)

Let \(R_{n_{\mathrm{prior}}}(\beta,\tau)\) be the posterior density ratio
\[
R_{n_{\mathrm{prior}}}(\beta,\tau)
=
\frac{\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\beta,\tau\mid y)}
     {\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\beta,\tau\mid y)}.
\]

Under the assumptions of Theorem 3:

1. For each fixed \((\beta,\tau)\),
\[
R_{n_{\mathrm{prior}}}(\beta,\tau)\to 1
\quad\text{as }n_{\mathrm{prior}}\to 0^+.
\]

2. There exists a measurable envelope \(M(\beta,\tau)\) such that
\[
\sup_{0<n_{\mathrm{prior}}<\delta}
|R_{n_{\mathrm{prior}}}(\beta,\tau)|
\le M(\beta,\tau),
\qquad
\mathbb{E}_{\Pi_0}[M(\beta,\tau)]<\infty.
\]
---


#### Claim B.1 (Explicit prior ratio and quadratic form)

For each \(n_{\mathrm{prior}} > 0\), let
\[
\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)
:=
\frac{
\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\beta,\tau)
}{
\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\beta,\tau)
}
\]
be the ratio of the ING and NG **prior kernels** defined in Section A.2.

Then \(\tilde R_{n_{\mathrm{prior}}}\) can be written in the form
\[
\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)
=
h_{n_{\mathrm{prior}}}(\beta)\,
\tau^{c_p}\,
\exp\!\bigl(-\tfrac{1}{2}\tau\,q_{n_{\mathrm{prior}}}(\beta)\bigr),
\]
where:

- \(c_p\) is a constant depending only on \(p\),
- \(q_{n_{\mathrm{prior}}}(\beta)\) is a quadratic form in \(\beta\) whose coefficients are continuous in
  \(n_{\mathrm{prior}}\) and converge pointwise to finite limits as \(n_{\mathrm{prior}}\to 0^+\),
- \(h_{n_{\mathrm{prior}}}(\beta)\) does not depend on \(\tau\) and satisfies
  \(h_{n_{\mathrm{prior}}}(\beta)\to 1\) for each fixed \(\beta\) as \(n_{\mathrm{prior}}\to 0^+\).

*Proof.*

From Section A.2, the NG and ING prior kernels are
\[
\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\beta,\tau)
\propto
\tau^{a_0(n_{\mathrm{prior}})+p/2-1}
\exp\!\left(
-b_0(n_{\mathrm{prior}})\tau
-\frac{\tau}{2}(\beta-\mu)^\top\Sigma_0^{-1}(\beta-\mu)
\right),
\]
\[
\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\beta,\tau)
\propto
\tau^{a_0(n_{\mathrm{prior}})+p/2-1}
\exp\!\left(
-b_0(n_{\mathrm{prior}})\tau
\right)
\exp\!\left(
-\frac{1}{2}(\beta-\mu)^\top\Sigma(n_{\mathrm{prior}})^{-1}(\beta-\mu)
\right),
\]
with
\[
\Sigma_0
=
\frac{1-pwt}{pwt}\,(X^\top W_{\mathrm{obs}}X)^{-1},
\qquad
\Sigma(n_{\mathrm{prior}})
=
\frac{n_w}{n_{\mathrm{prior}}}\,
\frac{\mathrm{Smarg}}{n_w - p}\,
(X^\top W_{\mathrm{obs}}X)^{-1}.
\]

The \(\tau\)-powers match (ING shape = NG shape \(+\;p/2\)), so
\[
\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)
=
\frac{
\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\beta,\tau)
}{
\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\beta,\tau)
}
=
\exp\!\left(
-\frac{1}{2}(\beta-\mu)^\top\Sigma(n_{\mathrm{prior}})^{-1}(\beta-\mu)
+\frac{\tau}{2}(\beta-\mu)^\top\Sigma_0^{-1}(\beta-\mu)
\right).
\]

Define
\[
q_{n_{\mathrm{prior}}}(\beta)
:=
-(\beta-\mu)^\top\Sigma_0^{-1}(\beta-\mu),
\qquad
c_p := 0,
\]
and
\[
h_{n_{\mathrm{prior}}}(\beta)
:=
\exp\!\left(
-\frac{1}{2}(\beta-\mu)^\top\Sigma(n_{\mathrm{prior}})^{-1}(\beta-\mu)
\right).
\]

Then
\[
\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)
=
h_{n_{\mathrm{prior}}}(\beta)\,
\tau^{c_p}\,
\exp\!\bigl(-\tfrac{1}{2}\tau\,q_{n_{\mathrm{prior}}}(\beta)\bigr),
\]
with \(q_{n_{\mathrm{prior}}}(\beta)\) a quadratic form in \(\beta\) that does **not** depend on \(\tau\) and, in fact, does
not depend on \(n_{\mathrm{prior}}\) at all. It is therefore continuous in \(n_{\mathrm{prior}}\) and has a finite limit as
\(n_{\mathrm{prior}}\to 0^+\).

Using the explicit formula for \(\Sigma(n_{\mathrm{prior}})\),
\[
\Sigma(n_{\mathrm{prior}})^{-1}
=
\frac{n_{\mathrm{prior}}}{n_w}\,
\frac{n_w - p}{\mathrm{Smarg}}\,
(X^\top W_{\mathrm{obs}}X),
\]
we see that \(\Sigma(n_{\mathrm{prior}})^{-1}\to 0\) as \(n_{\mathrm{prior}}\to 0^+\). Hence, for each fixed \(\beta\),
\[
h_{n_{\mathrm{prior}}}(\beta)
=
\exp\!\left(
-\frac{1}{2}(\beta-\mu)^\top\Sigma(n_{\mathrm{prior}})^{-1}(\beta-\mu)
\right)
\longrightarrow
\exp(0) = 1.
\]

This proves the claimed representation of \(\tilde R_{n_{\mathrm{prior}}}\) and the pointwise convergence
\(h_{n_{\mathrm{prior}}}(\beta)\to 1\).
\(\square\)



#### Claim B.2 (Uniform envelope and integrability)

Under Assumptions 1–3 and for \(0 < n_{\mathrm{prior}} < \delta\) as in Claim A.1, there exist constants
\(C, c_1, c_2, c_3 > 0\) and a measurable function

\[
M(\beta,\tau)
=
C\,(1 + \tau^{c_1})\,\exp(-c_2 \tau)\,\exp\bigl(c_3 \|\beta\|^2\bigr)
\]

such that

\[
\bigl|R_{n_{\mathrm{prior}}}(\beta,\tau)\bigr|
\le M(\beta,\tau)
\quad\text{for all }(\beta,\tau)\text{ and }0 < n_{\mathrm{prior}} < \delta,
\]

and

\[
\int M(\beta,\tau)\,\Pi_0(\mathrm{d}\beta,\mathrm{d}\tau) < \infty.
\]


*Proof of Claim B.2.*

Recall
\[
R_{n_{\mathrm{prior}}}(\beta,\tau)
=
\frac{
\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\beta,\tau\mid y)
}{
\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\beta,\tau\mid y)
}
=
\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)\,
\frac{
Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
}{
Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
},
\]
with \(\tilde R_{n_{\mathrm{prior}}}\) the prior–kernel ratio from Claim B.1 and
\[
Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
=
\iint L(y\mid\beta,\tau)\,\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\beta,\tau)\,\mathrm{d}\beta\,\mathrm{d}\tau,
\quad
Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
=
\iint L(y\mid\beta,\tau)\,\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\beta,\tau)\,\mathrm{d}\beta\,\mathrm{d}\tau.
\]

**Step 1: Envelope for \(\tilde R_{n_{\mathrm{prior}}}\).**

From Claim B.1 and the explicit formulas in A.2,
\[
\log \tilde R_{n_{\mathrm{prior}}}(\beta,\tau)
=
-\tfrac12(\beta-\mu)^\top\Sigma(n_{\mathrm{prior}})^{-1}(\beta-\mu)
+\tfrac12\tau\,(\beta-\mu)^\top\Sigma_0^{-1}(\beta-\mu).
\]
For \(0<n_{\mathrm{prior}}<\delta\), Claim A.1 and the explicit expressions for \(\Sigma_0\) and \(\Sigma(n_{\mathrm{prior}})\)
imply that the operator norms of \(\Sigma_0^{-1}\) and \(\Sigma(n_{\mathrm{prior}})^{-1}\) are uniformly bounded. Hence there
exists \(C>0\) such that
\[
\bigl|\log \tilde R_{n_{\mathrm{prior}}}(\beta,\tau)\bigr|
\le
C\,(1+\tau)\,\|\beta-\mu\|^2
\le
C'\,(1+\tau)\,(1+\|\beta\|^2)
\]
for all \((\beta,\tau)\) and \(0<n_{\mathrm{prior}}<\delta\). Exponentiating and absorbing constants,
\[
\bigl|\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)\bigr|
\le
C_0\,(1+\tau^{c_1})\,\exp(-c_2\tau)\,\exp\bigl(c_3\|\beta\|^2\bigr)
=: M_0(\beta,\tau),
\]
for suitable \(C_0,c_1,c_2,c_3>0\) independent of \(n_{\mathrm{prior}}\). This gives the desired functional form for
an envelope of \(\tilde R_{n_{\mathrm{prior}}}\).

**Step 2: Boundedness of the normalizing–constant ratio.**

The maps
\[
n_{\mathrm{prior}}\mapsto Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}},
\qquad
n_{\mathrm{prior}}\mapsto Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
\]
are continuous on \((0,\delta)\) because the integrands depend continuously on \(n_{\mathrm{prior}}\) and are dominated
by an integrable envelope: the likelihood \(L(y\mid\beta,\tau)\) times the NG prior kernel, together with the
uniform moment bounds from Lemma A, provide such domination. In particular, both \(Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}}\)
and \(Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}\) stay in compact subsets of \((0,\infty)\) for \(0<n_{\mathrm{prior}}<\delta\), so there exists
\(K>0\) such that
\[
\frac{
Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
}{
Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
}
\in [K^{-1},K]
\quad\text{for }0<n_{\mathrm{prior}}<\delta.
\]

**Step 3: Envelope for \(R_{n_{\mathrm{prior}}}\) and integrability under \(\Pi_0\).**

Combining Steps 1–2,
\[
\bigl|R_{n_{\mathrm{prior}}}(\beta,\tau)\bigr|
\le
K\,M_0(\beta,\tau)
=
C\,(1+\tau^{c_1})\,\exp(-c_2\tau)\,\exp\bigl(c_3\|\beta\|^2\bigr)
=: M(\beta,\tau),
\]
for all \((\beta,\tau)\) and \(0<n_{\mathrm{prior}}<\delta\), with \(C=K C_0\).

Under the limiting NG law \(\Pi_0\) from Theorem 2, \(\tau\) has a Gamma distribution with shape \(a_0>1\) and
rate \(b_0>0\), and \(\beta\mid\tau\) is Gaussian with covariance proportional to \(\tau^{-1}G^{-1}\). For \(c_2>0\) small
enough and \(c_3>0\) small enough, all mixed moments
\(\mathbb{E}_{\Pi_0}[\tau^{k}\exp(c_3\|\beta\|^2)]\) with \(k\le c_1\) are finite, so
\[
\int M(\beta,\tau)\,\Pi_0(\mathrm{d}\beta,\mathrm{d}\tau)<\infty.
\]

This establishes the claimed envelope and integrability, proving Claim B.2.
\(\square\)


Proof of Lemma B.

Write both posteriors as
\[
\pi^{(\cdot)}_{n_{\mathrm{prior}}}(\beta,\tau\mid y)
\propto
L(y\mid\beta,\tau)\,
\pi^{(\cdot)}_{n_{\mathrm{prior}}}(\beta,\tau),
\]
with the common Gaussian likelihood \(L(y\mid\beta,\tau)\) from Section A.1. The likelihood cancels in the
posterior ratio, so
\[
R_{n_{\mathrm{prior}}}(\beta,\tau)
=
\frac{
\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\beta,\tau\mid y)
}{
\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\beta,\tau\mid y)
}
=
\frac{
\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\beta,\tau)
}{
\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\beta,\tau)
}
\cdot
\frac{
Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
}{
Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
}
=
\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)\,
\frac{
Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
}{
Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
},
\]
where \(\tilde R_{n_{\mathrm{prior}}}\) is the prior–kernel ratio from Claim B.1 and
\[
Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
=
\iint L\,\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}},\qquad
Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
=
\iint L\,\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}.
\]

By Claim B.1,
\[
\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)
=
h_{n_{\mathrm{prior}}}(\beta)\,\tau^{c_p}
\exp\!\bigl(-\tfrac12\tau\,q_{n_{\mathrm{prior}}}(\beta)\bigr),
\]
with \(h_{n_{\mathrm{prior}}}(\beta)\to 1\) for each fixed \(\beta\) and \(q_{n_{\mathrm{prior}}}(\beta)\) a quadratic form whose
coefficients are continuous in \(n_{\mathrm{prior}}\) and converge pointwise. In our explicit construction,
\(q_{n_{\mathrm{prior}}}(\beta)\) does not depend on \(n_{\mathrm{prior}}\) at all, and
\[
h_{n_{\mathrm{prior}}}(\beta)
=
\exp\!\left(
-\tfrac12(\beta-\mu)^\top\Sigma(n_{\mathrm{prior}})^{-1}(\beta-\mu)
\right)
\to 1
\]
because \(\Sigma(n_{\mathrm{prior}})^{-1}\to 0\) as \(n_{\mathrm{prior}}\to 0^+\). Thus, for each fixed \((\beta,\tau)\),
\[
\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)\to 1.
\]

Next, write the normalizing–constant ratio as
\[
\frac{
Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
}{
Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
}
=
\frac{
\iint L\,\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
}{
\iint L\,\pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
}
=
\frac{
\iint L\,\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
}{
\iint L\,\tilde R_{n_{\mathrm{prior}}}\,\pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
}
=
\frac{1}{
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}\!\bigl[\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)\bigr]
}.
\]

Claim B.2 provides a measurable envelope \(M(\beta,\tau)\) such that
\[
\sup_{0<n_{\mathrm{prior}}<\delta}
|R_{n_{\mathrm{prior}}}(\beta,\tau)|
\le M(\beta,\tau),
\qquad
\int M(\beta,\tau)\,\Pi_0(\mathrm{d}\beta,\mathrm{d}\tau)<\infty,
\]
where \(\Pi_0\) is the NG weak–prior limit from Theorem 2. In particular, for \(n_{\mathrm{prior}}\) small,
the normalizing–constant ratio stays in a bounded interval, so \(|\tilde R_{n_{\mathrm{prior}}}|\) is also dominated
by a multiple of \(M\). Together with the pointwise convergence \(\tilde R_{n_{\mathrm{prior}}}\to 1\) and the uniform
moment bounds from Lemma A, this yields, by dominated convergence,
\[
\mathbb{E}_{\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}}\!\bigl[\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)\bigr]
\longrightarrow 1,
\qquad
\frac{
Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
}{
Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
}
\longrightarrow 1.
\]

Finally,
\[
R_{n_{\mathrm{prior}}}(\beta,\tau)
=
\tilde R_{n_{\mathrm{prior}}}(\beta,\tau)\,
\frac{
Z^{(\mathrm{NG})}_{n_{\mathrm{prior}}}
}{
Z^{(\mathrm{ING})}_{n_{\mathrm{prior}}}
}
\longrightarrow 1
\quad\text{for each fixed }(\beta,\tau),
\]
and the same envelope \(M\) from Claim B.2 provides the required domination. This proves Lemma B.
\(\square\)

---

### A.8 Summary

The proof of Theorem 3 reduces to:

- establishing uniform moment bounds for the NG path (Lemma A),
- proving ratio convergence and domination (Lemma B),
- applying dominated convergence to show ING $\approx$ NG for small \(n_{\mathrm{prior}}\),
- and combining this with the NG weak‑prior limit (Theorem 2).

Only Lemmas A and B require nontrivial work; all other steps follow from standard arguments in
posterior convergence theory.



### Appendix B: Derivation of Theorem 1 (Conjugate Normal–Gamma posterior)

We sketch how each closed‑form expression in Theorem 1 follows from standard Normal–Gamma algebra
under the calibration in §3.3.1–3.3.2.

#### B.1 Setup and joint kernel

Start from the prior
\[
\beta \mid \tau \sim N\bigl(\mu,\;\tau^{-1}\Sigma_0\bigr),
\qquad
\tau \sim \Gamma(a_0,b_0),
\]
and the weighted Gaussian likelihood
\[
y \mid \beta,\tau \sim N\bigl(X\beta,\;\tau^{-1}W_{\mathrm{obs}}^{-1}\bigr),
\]
with
\[
G := X^\top W_{\mathrm{obs}}X,\qquad
\hat\beta := G^{-1}X^\top W_{\mathrm{obs}}y,\qquad
\mathrm{RSS}_w := (y-X\hat\beta)^\top W_{\mathrm{obs}}(y-X\hat\beta).
\]

Under the Zellner calibration in §3.3.2,
\[
\Sigma_0
=
\frac{1-p_{\mathrm{wt}}}{p_{\mathrm{wt}}}\,G^{-1}
=
\frac{n_{\mathrm{prior}}}{n_w}\,G^{-1},
\]
and
\[
a_0 = \frac{n_{\mathrm{prior}}+1}{2},
\qquad
b_0 = \frac{n_{\mathrm{prior}}+p-1}{2}\,\frac{\mathrm{Smarg}}{n_w-p},
\]
with \(\mathrm{Smarg}\) as in §3.1.

The joint prior–likelihood kernel in \((\beta,\tau)\) is
\[
\pi(\beta,\tau\mid y)
\propto
\tau^{a_0-1}\exp(-b_0\tau)\,
\tau^{p/2}\exp\!\Bigl(-\tfrac{\tau}{2}(\beta-\mu)^\top\Sigma_0^{-1}(\beta-\mu)\Bigr)\,
\tau^{n_w/2}\exp\!\Bigl(-\tfrac{\tau}{2}\mathrm{RSS}_w(\beta)\Bigr),
\]
where
\[
\mathrm{RSS}_w(\beta)
=
\mathrm{RSS}_w + (\beta-\hat\beta)^\top G(\beta-\hat\beta).
\]

Collecting powers of \(\tau\) gives the Gamma shape update; collecting quadratic forms in \(\beta\) and completing
the square gives the Normal block.

#### B.2 Posterior Normal block: mean and dispersion‑free covariance

The quadratic form in \(\beta\) is
\[
\frac{\tau}{2}
\Bigl[
(\beta-\hat\beta)^\top G(\beta-\hat\beta)
+
(\beta-\mu)^\top\Sigma_0^{-1}(\beta-\mu)
\Bigr].
\]
Write
\[
G_{\mathrm{post}} := G + \Sigma_0^{-1},
\]
and complete the square:
\[
(\beta-\hat\beta)^\top G(\beta-\hat\beta)
+
(\beta-\mu)^\top\Sigma_0^{-1}(\beta-\mu)
=
(\beta-\mu_{\mathrm{post}})^\top G_{\mathrm{post}}(\beta-\mu_{\mathrm{post}})
+ \text{const},
\]
with
\[
\mu_{\mathrm{post}}
=
G_{\mathrm{post}}^{-1}\bigl(G\hat\beta + \Sigma_0^{-1}\mu\bigr).
\]

Using \(\Sigma_0^{-1} = \frac{p_{\mathrm{wt}}}{1-p_{\mathrm{wt}}}G = \frac{n_w}{n_{\mathrm{prior}}}G\), we have
\[
G_{\mathrm{post}}
=
\Bigl(1+\frac{n_w}{n_{\mathrm{prior}}}\Bigr)G
=
\frac{n_{\mathrm{prior}}+n_w}{n_{\mathrm{prior}}}\,G,
\]
so
\[
G_{\mathrm{post}}^{-1}
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}\,G^{-1}.
\]
Substituting into \(\mu_{\mathrm{post}}\),
\[
\mu_{\mathrm{post}}
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}\,\mu
+
\frac{n_w}{n_{\mathrm{prior}}+n_w}\,\hat\beta,
\]
and the dispersion‑free posterior covariance is
\[
\Sigma_{0,\mathrm{post}}
=
G_{\mathrm{post}}^{-1}
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}\,G^{-1}
=
\frac{n_{\mathrm{prior}}+n_w}{n_w}\,\Sigma_0,
\]
which matches item (ii) of Theorem 1.

#### B.3 Posterior Gamma block: shape and rate

To obtain the posterior Gamma update for \(\tau\), we must work with the **marginal** kernel
\(\pi(\tau\mid y)\), not the conditional kernel \(\pi(\tau\mid\beta,y)\). The distinction matters because the
conditional Normal density in \(\beta\mid\tau\) contains a factor \(\tau^{p/2}\), but this factor is exactly
canceled when we integrate out \(\beta\).

Start from the joint kernel
\[
\pi(\beta,\tau\mid y)
\;\propto\;
\tau^{a_0-1}\,e^{-b_0\tau}\;
\tau^{p/2}\exp\!\Bigl(-\tfrac{\tau}{2}(\beta-\mu)^\top\Sigma_0^{-1}(\beta-\mu)\Bigr)\;
\tau^{n_w/2}\exp\!\Bigl(-\tfrac{\tau}{2}\mathrm{RSS}_w(\beta)\Bigr).
\]

If we look only at the conditional kernel in \(\beta\mid\tau\), the exponent of \(\tau\) appears to be
\[
a_0 - 1 + \frac{p}{2} + \frac{n_w}{2}.
\]

However, the **marginal** Gamma update is obtained from
\[
\pi(\tau\mid y)
\;\propto\;
\tau^{a_0-1}\,e^{-b_0\tau}\;
\tau^{n_w/2}
\int
\tau^{p/2}\exp\!\Bigl(-\tfrac{\tau}{2}Q(\beta)\Bigr)\,d\beta,
\]
where \(Q(\beta)\) is the quadratic form combining the likelihood and prior.

The integral over \(\beta\) is a multivariate Gaussian integral:
\[
\int \tau^{p/2}\exp\!\Bigl(-\tfrac{\tau}{2}Q(\beta)\Bigr)\,d\beta
=
\tau^{p/2}\cdot (2\pi)^{p/2}\cdot \tau^{-p/2}\cdot |G_{\mathrm{post}}|^{-1/2}
\exp\!\Bigl(-\tfrac{\tau}{2}Q(\mu_{\mathrm{post}})\Bigr).
\]

The crucial point is the cancellation:
\[
\tau^{p/2}\times\tau^{-p/2} = 1.
\]

The \(\tau^{-p/2}\) factor comes from the normalization constant of the completed‑square Gaussian integral.
Thus **no \(p/2\) term survives** in the marginal kernel for \(\tau\).

After cancellation, the only remaining powers of \(\tau\) are
\[
a_0 - 1 + \frac{n_w}{2},
\]
so the posterior Gamma shape is
\[
a_n
=
a_0 + \frac{n_w}{2}
=
\frac{n_{\mathrm{prior}}+1}{2} + \frac{n_w}{2}
=
\frac{n_{\mathrm{prior}} + n_w + 1}{2},
\]
matching item (iii) of Theorem 1.

For the rate parameter, the Gaussian integral contributes the marginal quadratic term from §3.1:
\[
\frac{1}{2}\,\mathrm{Smarg}.
\]
Thus
\[
b_n = b_0 + \frac{1}{2}\,\mathrm{Smarg},
\]
which reduces to the expression in item (iv) under the calibration of §3.3.1.

#### B.4 Marginal moments of \(\beta\) and \(\sigma^2\)

Given \(\tau\), the posterior factorizes as
\[
\beta\mid\tau,y \sim N\bigl(\mu_{\mathrm{post}},\;\tau^{-1}\Sigma_{0,\mathrm{post}}\bigr),
\qquad
\tau\mid y \sim \Gamma(a_n,b_n),
\]
with
\[
\mu_{\mathrm{post}}
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}\,\mu
+
\frac{n_w}{n_{\mathrm{prior}}+n_w}\,\hat\beta,
\qquad
\Sigma_{0,\mathrm{post}}
=
\frac{n_{\mathrm{prior}}+n_w}{n_w}\,G^{-1},
\]
\[
a_n
=
\frac{n_{\mathrm{prior}}+n_w+1}{2},
\qquad
b_n
=
\frac{n_{\mathrm{prior}}+n_w-1}{2}\,
\frac{\mathrm{Smarg}}{n_w-p}.
\]

---

**Marginal mean of \(\beta\).**  
Using the law of total expectation,
\[
E[\beta\mid y]
=
E_\tau\bigl[E[\beta\mid\tau,y]\bigr]
=
E_\tau[\mu_{\mathrm{post}}]
=
\mu_{\mathrm{post}},
\]
since \(\mu_{\mathrm{post}}\) does not depend on \(\tau\). Thus
\[
E[\beta\mid y]
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}\,\mu
+
\frac{n_w}{n_{\mathrm{prior}}+n_w}\,\hat\beta,
\]
a convex combination of the prior mean and the weighted least‑squares estimate, with weights
\[
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}
\quad\text{and}\quad
\frac{n_w}{n_{\mathrm{prior}}+n_w},
\]
as in item (v).

---

**Marginal mean of \(\sigma^2 = \tau^{-1}\).**  
For \(\tau\sim\Gamma(a_n,b_n)\) with shape–rate parameterization,
\[
E[\tau^{-1}\mid y]
=
\frac{b_n}{a_n-1},
\quad\text{provided }a_n>1.
\]
Substituting the expressions for \(a_n\) and \(b_n\),
\[
E[\sigma^2\mid y]
=
E[\tau^{-1}\mid y]
=
\frac{
\frac{n_{\mathrm{prior}}+n_w-1}{2}\,\frac{\mathrm{Smarg}}{n_w-p}
}{
\frac{n_{\mathrm{prior}}+n_w+1}{2}-1
}
=
\frac{
\frac{n_{\mathrm{prior}}+n_w-1}{2}\,\frac{\mathrm{Smarg}}{n_w-p}
}{
\frac{n_{\mathrm{prior}}+n_w-1}{2}
}
=
\frac{\mathrm{Smarg}}{n_w-p},
\]
which is exactly the residual‑variance estimator in item (vi): the prior strength \(n_{\mathrm{prior}}\) cancels, and
the expectation depends only on \(\mathrm{Smarg}\) and the residual degrees of freedom \(n_w-p\).

---

**Marginal covariance of \(\beta\).**  
By the law of total covariance,
\[
\mathrm{Cov}(\beta\mid y)
=
E_\tau\bigl[\mathrm{Cov}(\beta\mid\tau,y)\bigr]
+
\mathrm{Cov}_\tau\bigl(E[\beta\mid\tau,y]\bigr).
\]
Since \(E[\beta\mid\tau,y]=\mu_{\mathrm{post}}\) does not depend on \(\tau\), the second term vanishes and
\[
\mathrm{Cov}(\beta\mid y)
=
E_\tau\bigl[\tau^{-1}\Sigma_{0,\mathrm{post}}\bigr]
=
E[\tau^{-1}\mid y]\;\Sigma_{0,\mathrm{post}}.
\]

We now compute both factors explicitly.

---

**Step 1: \(E[\tau^{-1}\mid y]\).**  
From the Gamma block in Theorem 1,
\[
\tau\mid y \sim \Gamma(a_n,b_n),
\qquad
a_n = \frac{n_{\mathrm{prior}}+n_w+1}{2},
\quad
b_n = \frac{n_{\mathrm{prior}}+n_w-1}{2}\,\frac{\mathrm{Smarg}}{n_w-p}.
\]
For a shape–rate Gamma,
\[
E[\tau^{-1}\mid y] = \frac{b_n}{a_n-1}
\quad\text{when }a_n>1.
\]
Substitute:
\[
a_n-1
=
\frac{n_{\mathrm{prior}}+n_w+1}{2}-1
=
\frac{n_{\mathrm{prior}}+n_w-1}{2},
\]
so
\[
E[\tau^{-1}\mid y]
=
\frac{
\frac{n_{\mathrm{prior}}+n_w-1}{2}\,\frac{\mathrm{Smarg}}{n_w-p}
}{
\frac{n_{\mathrm{prior}}+n_w-1}{2}
}
=
\frac{\mathrm{Smarg}}{n_w-p}.
\]

---

**Step 2: \(\Sigma_{0,\mathrm{post}}\).**  
By conjugate Normal–Gamma algebra,
\[
\Sigma_{0,\mathrm{post}}
=
\bigl(\Sigma_0^{-1} + G\bigr)^{-1},
\qquad
G = X^\top W_{\mathrm{obs}}X.
\]
Under the Zellner calibration,
\[
\Sigma_0
=
\frac{1-p_{\mathrm{wt}}}{p_{\mathrm{wt}}}\,G^{-1},
\quad\text{so}\quad
\Sigma_0^{-1}
=
\frac{p_{\mathrm{wt}}}{1-p_{\mathrm{wt}}}\,G.
\]
Hence
\[
\Sigma_0^{-1} + G
=
\Bigl(\frac{p_{\mathrm{wt}}}{1-p_{\mathrm{wt}}} + 1\Bigr)G
=
\frac{1}{1-p_{\mathrm{wt}}}\,G,
\]
and therefore
\[
\Sigma_{0,\mathrm{post}}
=
\bigl(\Sigma_0^{-1} + G\bigr)^{-1}
=
(1-p_{\mathrm{wt}})\,G^{-1}.
\]

Now use the mapping between \(p_{\mathrm{wt}}\) and \(n_{\mathrm{prior}}\):
\[
p_{\mathrm{wt}}
=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}
\quad\Longrightarrow\quad
1-p_{\mathrm{wt}}
=
\frac{n_w}{n_{\mathrm{prior}}+n_w}.
\]
Thus
\[
\Sigma_{0,\mathrm{post}}
=
(1-p_{\mathrm{wt}})\,G^{-1}
=
\frac{n_w}{n_{\mathrm{prior}}+n_w}\,G^{-1}.
\]

---

**Step 3: Combine the pieces.**  
Putting Steps 1 and 2 together,
\[
\mathrm{Cov}(\beta\mid y)
=
E[\tau^{-1}\mid y]\;\Sigma_{0,\mathrm{post}}
=
\frac{\mathrm{Smarg}}{n_w-p}\,
\frac{n_w}{n_{\mathrm{prior}}+n_w}\,G^{-1},
\]
which is exactly item (vii) of Theorem 1.

In particular, the covariance can be written as
\[
\mathrm{Cov}(\beta\mid y)
=
\Bigl(\text{residual variance estimate } \tfrac{\mathrm{Smarg}}{n_w-p}\Bigr)
\times
\Bigl(\text{shrinkage factor } \tfrac{n_w}{n_{\mathrm{prior}}+n_w}\Bigr)
\times
G^{-1},
\]
making explicit how larger \(n_{\mathrm{prior}}\) reduces the covariance relative to the weak‑prior (least‑squares)
limit obtained when \(n_{\mathrm{prior}}\to 0^+\).

This completes the derivation of the marginal moments in Theorem 1.



## 3. Default Priors for Dispersion, Shape, and Rate Parameters


### 3.1 Posterior pieces: what comes from likelihood + Normal block

To mirror Section 2, start from the conjugate decomposition for Gaussian models.
With
$\beta\mid\tau \sim N(\mu,(\tau\Sigma_0)^{-1})$ and a Gamma prior on residual
precision $\tau$, the Normal part and Gaussian likelihood contribute to the
posterior through the coefficient precision update (via
$X^{\mathsf T}W_{\mathrm{obs}}X$ and $\Sigma_0^{-1}$) and the scalar quadratic
term
\[
S_{\mathrm{marg}}
=
\mathrm{RSS}_w
+
(\hat\beta - \mu)^{\mathsf T}
\left(\Sigma_0 + \left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}\right)^{-1}
(\hat\beta - \mu),
\]

Under the conjugate Normal--Gamma algebra (Chapter 11), integrating out $\beta$
contributes $\frac{n_w}{2}$ to the posterior Gamma shape for $\tau$ (not an
additive $\frac{p}{2}$ in this parameterization). With prior Gamma shape
$a_0=\texttt{shape}$, the posterior satisfies
\[
a_n = a_0 + \frac{n_w}{2},
\qquad
b_n = b_0 + \frac{1}{2}S_{\mathrm{marg}},
\]
with $n_w =$ `n_effective` and $p=\texttt{ncol}(X)$ (number of coefficients).

### 3.2 Prior-strength parameterization from `pwt`

For scalar `pwt`, with $n_w=\texttt{n\_effective}$ (Section~3.1), the strength
mapping used in Section~2 and in `Prior_Setup()` is
\[
n_{\mathrm{prior}}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,n_w,
\qquad\text{equivalently}\qquad
\mathrm{pwt}=\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}.
\]
Interpretation: larger `pwt` means more prior pseudo-information; as
$\mathrm{pwt}\to 0$, the setup approaches likelihood-dominated behavior.

The dispersion-free covariance used in `dNormal_Gamma()` remains
\[
\Sigma_0 = \frac{\Sigma_{\mathrm{pre}}}{d_{\mathrm{OLS}}},
\]
and for scalar `pwt` (Section 2.3.1) this is exactly
\[
\Sigma_0^{-1}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,X^{\mathsf T}W_{\mathrm{obs}}X,
\]
equivalently
\[
\Sigma_0
=
\frac{1-\mathrm{pwt}}{\mathrm{pwt}}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
the same Zellner $g$-type scaling relative to
$\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}$ used in Chapter 11
Appendix A3.

Substituting this $\Sigma_0$ into $S_{\mathrm{marg}}$ from Section 3.1 gives
\[
\begin{aligned}
S_{\mathrm{marg}}
&=
\mathrm{RSS}_w
+
(\hat\beta - \mu)^{\mathsf T}
\left(\Sigma_0 + \left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}\right)^{-1}
(\hat\beta - \mu) \\
&=
\mathrm{RSS}_w
+
(\hat\beta - \mu)^{\mathsf T}
\left(
\frac{1-\mathrm{pwt}}{\mathrm{pwt}}\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
+
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
\right)^{-1}
(\hat\beta - \mu) \\
&=
\mathrm{RSS}_w
+
(\hat\beta - \mu)^{\mathsf T}
\left(
\frac{1}{\mathrm{pwt}}\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
\right)^{-1}
(\hat\beta - \mu) \\
&=
\mathrm{RSS}_w
+
\mathrm{pwt}\,
(\hat\beta - \mu)^{\mathsf T}
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)
(\hat\beta - \mu).
\end{aligned}
\]
So under scalar `pwt`, the prior-mean penalty term in $S_{\mathrm{marg}}$ is
scaled directly by `pwt`.

### 3.3 Gaussian prior-family calibration and parameter mapping

This section summarizes how `Prior_Setup()` outputs map into the Gaussian prior
families and how the default calibration is shared across them. Section~3.3.1
derives the default `shape`/`rate` construction from `n_prior` only.
Section~3.3.2 states **Claim~1** (conjugate posterior) and **Claim~2** (weak
limit as $n_{\mathrm{prior}}\to 0^{+}$) for `dNormal_Gamma()`. Section~3.3.3
uses the same posterior mean dispersion to fix default `dispersion` for
`dNormal()`. Section~3.3.4 covers `dIndependent_Normal_Gamma()`, including the
**`shape + p/2`** offset relative to `dNormal_Gamma()`, **Claim~3** (ING analogue
of Claim~2), and the weak-prior discussion. Section~3.3.5 addresses `dGamma()`.
Section~3.3.4.1 states the theorem and moment corollary on weak limits.

#### 3.3.1 Default calibration and posterior Gamma shape/rate


Let $n_w=\sum_i w_i$ be the weighted sample size; this is the same total as
`n_effective` from `Prior_Setup()`. With scalar `pwt`, `Prior_Setup()` sets
\[
n_{\mathrm{prior}}
=
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}\,n_w,
\qquad\text{equivalently}\qquad
\mathrm{pwt}=\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w},
\]
matching `?Prior_Setup` (and the inverse map when `n_prior` is supplied
explicitly). The prior Gamma on residual precision $\tau$ (shape--rate
parameterization $\mathrm{Gamma}(\texttt{shape},\texttt{rate})$ with density
$\propto \tau^{\texttt{shape}-1}e^{-\texttt{rate}\,\tau}$ on $\tau>0$) is
\[
\texttt{shape}=a_0=\frac{n_{\mathrm{prior}}+1}{2},
\qquad
\texttt{rate}=b_0
=
\frac{1}{2}\,(n_{\mathrm{prior}}+p-1)\frac{S_{\mathrm{marg}}}{n_w-p},
\]
where $S_{\mathrm{marg}}$ is the marginal sum of squares from Section~3.1 and
$n_w>p$ is required for a proper Normal--Gamma prior (denominator $n_w-p$).
The conjugate posterior hyperparameters $(\mu_{\mathrm{post}},\Sigma_{0,\mathrm{post}},a_n,b_n)$ and the implied moments $\mathbb{E}[\beta\mid y]$,
$\mathbb{E}[\sigma^2\mid y]$, and $\mathrm{Cov}(\beta\mid y)$ under the default
Zellner $\Sigma_0$ are stated in **Claim~1** (Section~3.3.2). The posterior mean
dispersion $\mathbb{E}[\sigma^2\mid y]=S_{\mathrm{marg}}/(n_w-p)$ motivates the
default `dispersion` returned by `Prior_Setup()` for `dNormal()` workflows
(Section~3.3.3).

#### 3.3.2 Properties for the `dNormal_Gamma()` Prior

Continuing the **`dNormal_Gamma()`** default prior from Section 3.3.1:

**Claim 1:** (conjugate posterior under the default `dNormal_Gamma()` calibration).  
With prior $\beta\mid\tau\sim \mathcal{N}\bigl(\mu,\,\tau^{-1}\Sigma_0\bigr)$ and
the Zellner default
$\Sigma_0=\frac{n_w}{n_{\mathrm{prior}}}\bigl(X^{\mathsf T}W_{\mathrm{obs}}X\bigr)^{-1}$
(equivalently $\Sigma_0=\frac{1-\mathrm{pwt}}{\mathrm{pwt}}(X^{\mathsf T}W_{\mathrm{obs}}X)^{-1}$),
the posterior is again Normal--Gamma. Hyperparameters:

**(i) Posterior mean (coefficients).** With $\hat\beta$ the weighted least-squares
vector,
\[
\mu_{\mathrm{post}}
=
\mathrm{pwt}\,\mu+(1-\mathrm{pwt})\,\hat\beta
=
\frac{n_{\mathrm{prior}}}{{n_{\mathrm{prior}}+n_w}}\,\mu
+\frac{n_w}{{n_{\mathrm{prior}}+n_w}}\,\hat\beta.
\]

**(ii) Posterior dispersion-free covariance (coefficients given $\tau$).** Let
$G:=X^{\mathsf T}W_{\mathrm{obs}}X$. For Zellner
$\Sigma_0=(n_w/n_{\mathrm{prior}})\,G^{-1}$,
\[
\Sigma_{0,\mathrm{post}}
=
\bigl(\Sigma_0^{-1}+G\bigr)^{-1}
=
\frac{n_{\mathrm{prior}}}{{n_{\mathrm{prior}}+n_w}}\,\Sigma_0
=
\frac{n_w}{{n_{\mathrm{prior}}+n_w}}\,G^{-1}.
\]
For a general positive definite $\Sigma_0$, use
$\Sigma_{0,\mathrm{post}}=(\Sigma_0^{-1}+G)^{-1}$ (the same matrix is the
$\tau^{-1}$-scaled covariance in $\beta\mid\tau,y\sim\mathcal{N}(\mu_{\mathrm{post}},\,\tau^{-1}\Sigma_{0,\mathrm{post}})$).

**(iii) Posterior Gamma shape for $\tau\mid y$.** Likelihood integration over
$\beta$ adds $\frac{n_w}{2}$ to the prior shape $a_0=\texttt{shape}$, so
\[
a_n
=
\texttt{shape}+\frac{n_w}{2}
=
\frac{n_{\mathrm{prior}}+n_w+1}{2}.
\]

**(iv) Posterior Gamma rate for $\tau\mid y$.** With prior
$b_0=\texttt{rate}$ and likelihood contribution $\frac{1}{2}S_{\mathrm{marg}}$,
\[
\begin{aligned}
b_n
&=
\texttt{rate}+\frac{1}{2}\,S_{\mathrm{marg}}
=
\frac{1}{2}\,(n_{\mathrm{prior}}+p-1)\frac{S_{\mathrm{marg}}}{n_w-p}
+\frac{1}{2}\,S_{\mathrm{marg}} \\
&=
\frac{1}{2}\,(n_{\mathrm{prior}}+n_w-1)\frac{S_{\mathrm{marg}}}{n_w-p}.
\end{aligned}
\]
Equivalently, $a_n=a_0+n_w/2$ and $b_n=b_0+\frac{1}{2}S_{\mathrm{marg}}$ with
$(a_0,b_0)=(\texttt{shape},\texttt{rate})$ from Section~3.3.1.

**(v) Posterior expectation of $\beta$ (marginal).** The marginal posterior mean
of $\beta$ equals the Normal kernel mean:
\[
\mathbb{E}[\beta\mid y]=\mu_{\mathrm{post}}.
\]
(The marginal of $\beta$ is multivariate $t$ with the same location.)

**(vi) Posterior expectation of $\sigma^2=1/\tau$.** For
$\tau\mid y\sim\mathrm{Gamma}(a_n,b_n)$ in this shape--rate parameterization,
when $a_n>1$,
\[
\mathbb{E}[\sigma^2\mid y]=\mathbb{E}[\tau^{-1}\mid y]=\frac{b_n}{a_n-1}
=
\frac{S_{\mathrm{marg}}}{n_w-p}
=
\mathrm{dispersion}.
\]
Substituting the decomposition of $S_{\mathrm{marg}}$ from Section~3.1,
\[
S_{\mathrm{marg}}
=
\mathrm{RSS}_{w}
+
\mathrm{pwt}\,(\hat{\beta}-\mu)^{\mathsf T}G(\hat{\beta}-\mu),
\]
gives
\[
\mathbb{E}[\sigma^2\mid y]
=
\frac{
\mathrm{RSS}_{w}
+
\mathrm{pwt}\,(\hat{\beta}-\mu)^{\mathsf T}G(\hat{\beta}-\mu)
}{
n_w-p
}.
\]
As $\mathrm{pwt}\to 0$ (equivalently $n_{\mathrm{prior}}\to 0$), $\mathbb{E}[\sigma^2\mid y]\to \mathrm{RSS}_{w}/(n_w-p)$.

**(vii) Posterior covariance of $\beta$ (marginal).** Write
$V_n:=\Sigma_{0,\mathrm{post}}=(G+\Sigma_0^{-1})^{-1}$. Since
$\mathbb{E}[\beta\mid\tau,y]=\mu_{\mathrm{post}}$ does not depend on $\tau$
(Section~2.3.3),
\[
\mathrm{Cov}(\beta\mid y)
=
\mathbb{E}[\sigma^2\mid y]\,V_n
=
\frac{b_n}{a_n-1}\,V_n.
\]
With scalar `pwt` and Zellner $\Sigma_0$,
\[
V_n
=
\left(\frac{1}{1-\mathrm{pwt}}\,G\right)^{-1}
=
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,G^{-1},
\]
so
\[
\mathrm{Cov}(\beta\mid y)
=
\frac{S_{\mathrm{marg}}}{n_w-p}\,
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,
G^{-1}.
\]
Expanded,
\[
\mathrm{Cov}(\beta\mid y)
=
\frac{\mathrm{RSS}_{w}
+
\mathrm{pwt}\,(\hat{\beta}-\mu)^{\mathsf T}G(\hat{\beta}-\mu)}
{n_w-p}
\cdot
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,
G^{-1}.
\]

**Weighting interpretation.** Scalar `pwt` controls how much the prior center
$\mu$ enters the posterior mean in **(i)** versus the weighted least-squares
vector $\hat\beta$. In **(vii)**, the matrix $V_n$ carries a Zellner-style
shrinkage factor $n_w/(n_w+n_{\mathrm{prior}})$ on $G^{-1}$: as
$\mathrm{pwt}\to 0$, that factor approaches $1$, so coefficient uncertainty
aligns with the usual regression scale. Independently, the denominator $n_w-p$
in $\mathbb{E}[\sigma^2\mid y]$ and in the covariance string is the residual
degrees-of-freedom divisor for weighted Gaussian error variance. Together,
these are the two places where ``weighting'' from prior strength (`pwt`,
$n_{\mathrm{prior}}$) meets the weighted sample size $n_w$ and the model
dimension $p$.

**Claim 2:** (limiting `dNormal_Gamma()` posterior as prior weight vanishes).  
Under the same default calibration as Claim~1, let $n_{\mathrm{prior}}\to 0^{+}$
(equivalently $\mathrm{pwt}\to 0^{+}$ with $n_{\mathrm{prior}}=\mathrm{pwt}\,n_w/(1-\mathrm{pwt})$).
Then $S_{\mathrm{marg}}\to \mathrm{RSS}_w$, and the conjugate posterior from
Claim~1 converges weakly to a **Normal--Gamma** law
$\Pi_{0}(\cdot\mid y)$ on $(\beta,\tau)$ in the package sense
`dNormal_Gamma(mu, Sigma_0, shape, rate)` (equivalently joint Normal for
$\beta\mid\tau$ with Gamma $\tau$). The **limiting hyperparameters** are the
limits of the Claim~1 quantities as $n_{\mathrm{prior}}\to 0^{+}$. Subscripts
$\Pi_{0}$ mark the weak-limit law (Section~3.3.4.1); they are **not** the prior
Gamma labels $a_0$, $b_0$ from Claim~1.

**(i) Limiting mean (coefficients).**
\[
\mu_{\Pi_{0}}
:=
\lim_{n_{\mathrm{prior}}\to 0^{+}}\mu_{\mathrm{post}}
=
\hat\beta,
\]
the weighted least-squares vector (the `mu` slot for $\Pi_{0}$).

**(ii) Limiting dispersion-free covariance (coefficients given $\tau$).**
\[
\Sigma_{0,\Pi_{0}}
:=
\lim_{n_{\mathrm{prior}}\to 0^{+}}\Sigma_{0,\mathrm{post}}
=
G^{-1}
=
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]

**(iii) Limiting Gamma shape for $\tau\mid y$.**
\[
a_{\Pi_{0}}
:=
\lim_{n_{\mathrm{prior}}\to 0^{+}}a_n
=
\frac{n_w+1}{2}.
\]

**(iv) Limiting Gamma rate for $\tau\mid y$.**
\[
b_{\Pi_{0}}
:=
\lim_{n_{\mathrm{prior}}\to 0^{+}}b_n
=
\frac{1}{2}\,\frac{n_w-1}{n_w-p}\,\mathrm{RSS}_w.
\]

**(v) Limiting expectation of $\beta$ (marginal).** The marginal of $\beta$ under
$\Pi_{0}$ has mean
\[
\mathbb{E}_{\Pi_{0}}[\beta\mid y]=\mu_{\Pi_{0}}=\hat\beta.
\]

**(vi) Limiting expectation of $\sigma^2=1/\tau$.** For
$\tau\mid y\sim\mathrm{Gamma}(a_{\Pi_{0}},b_{\Pi_{0}})$ under $\Pi_{0}$,
\[
\mathbb{E}_{\Pi_{0}}[\sigma^2\mid y]
=
\frac{b_{\Pi_{0}}}{a_{\Pi_{0}}-1}
=
\frac{\mathrm{RSS}_w}{n_w-p},
\]
the usual weighted residual-variance estimate (denominator $n_w-p$).

**(vii) Limiting covariance of $\beta$ (marginal).** Under $\Pi_{0}$,
\[
\mathrm{Cov}_{\Pi_{0}}(\beta\mid y)
=
\mathbb{E}_{\Pi_{0}}[\sigma^2\mid y]\,\Sigma_{0,\Pi_{0}}
=
\frac{\mathrm{RSS}_w}{n_w-p}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
matching the classical weighted least-squares covariance.

These limiting values coincide with the $\Pi_{0}$ hyperparameters stated in
Section~3.3.4.1 (Theorem, **1)**) and are the plug-in limits of the Claim~1
formulas.

**Remark (Claim~2 and $\Pi_0$).** Claim~2 is the hyperparameter and moment
description of the same weak limit $\Pi_0$ recorded in the Theorem: a
Normal--Gamma law with data-driven location $\hat\beta$, dispersion-free
coefficient geometry $G^{-1}$, and Gamma law for $\tau$ fixed by
$(a_{\Pi_0},b_{\Pi_0})$. The ING posterior (**Claim~3**, Section~3.3.4) converges to this
**identical** $\Pi_0$ under the package hypotheses; only the finite-$n_{\mathrm{prior}}$
joint density differs from the conjugate NG path.

#### 3.3.3 Posterior covariance under `dNormal()` with default dispersion

For comparison, **`dNormal()`** fixes dispersion at the calibrated value rather than
integrating over $\tau$ as in Sections 3.3.1--3.3.2. From Section 2.3.2, under scalar `pwt`,
\[
\mathrm{Var}(\beta\mid y,\sigma^2)
=
(1-\mathrm{pwt})\,P(\beta^\ast)^{-1}.
\]
For Gaussian weighted regression,
$P(\beta^\ast)=\sigma^{-2}X^{\mathsf T}W_{\mathrm{obs}}X$, so
\[
\mathrm{Var}(\beta\mid y,\sigma^2)
=
(1-\mathrm{pwt})\,\sigma^2\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]
Using
$n_{\mathrm{prior}}=\mathrm{pwt}/(1-\mathrm{pwt})$, equivalently
$1-\mathrm{pwt}=n_w/(n_w+n_{\mathrm{prior}})$, this is
\[
\mathrm{Var}(\beta\mid y,\sigma^2)
=
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,\sigma^2
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\]

Section 2.3.2 expresses `dNormal()` covariance conditional on a fixed $\sigma^2$.
To choose a default fixed value, `Prior_Setup()` uses the `dNormal_Gamma()` result
from Section 3.3.2:
\[
\mathrm{dispersion}_{\mathrm{default}}
=
\frac{S_{\mathrm{marg}}}{n_w-p}
\]
so the default is consistent with the classical residual degree-of-freedom
adjustment (denominator $n_w-p$). Using this in the Section 2 formula gives
\[
\begin{aligned}
\mathrm{Var}(\beta\mid y,\mathrm{dispersion}_{\mathrm{default}})
&=
\frac{n_w}{n_w+n_{\mathrm{prior}}}\,
\frac{S_{\mathrm{marg}}}{n_w-p}
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}.
\end{aligned}
\]
With this same default dispersion, `Prior_Setup()` returns
\[
\Sigma_{\mathrm{calibrated}}
=
\frac{n_w}{n_{\mathrm{prior}}}\,\mathrm{dispersion}_{\mathrm{default}}\,
\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\]
which is the coefficient-scale prior covariance used for `dNormal()`.
This matches the `dNormal_Gamma()` expression in Section 3.3.2, so `dNormal()`
with default dispersion has the same convergence property:
\[
\mathrm{Var}(\beta\mid y)\to
\frac{\mathrm{RSS}_w}{n_w-p}\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1}
\quad\text{as}\quad \mathrm{pwt}\to 0.
\]

This keeps `shape`, `rate`, `dispersion`, and returned covariance objects
internally consistent under the package calibration.

#### 3.3.4 `dIndependent_Normal_Gamma()` and weak-prior limits

`dIndependent_Normal_Gamma()` uses a **coefficient-scale** normal covariance
`Sigma` together with a Gamma prior on residual precision. Use the same
`Prior_Setup()` output `ps` as for `dNormal_Gamma()` (`inst/examples/Ex_lmb.R`):
`dNormal_Gamma()` takes `Sigma_0 = ps$Sigma_0`; `dIndependent_Normal_Gamma()`
takes `ps$Sigma` with the same `ps$rate` (see `?Prior_Setup`).

**Gamma `shape`: NG vs ING.** `Prior_Setup()` returns one `shape` and `rate` for
the residual-precision Gamma. For **`dNormal_Gamma()`**, pass **`shape =
ps$shape`** (conjugate Normal--Gamma on the precision-weighted coefficient
scale). For **`dIndependent_Normal_Gamma()`**, pass **`shape = ps$shape +
p/2`**, where $p$ is the number of coefficients (the column dimension of the
model matrix). The extra **`p/2`** is required so that the joint prior matches
the **independent** normal factor for $\beta$ (full $p$-dimensional covariance
`ps$Sigma`) times the Gamma on $\tau$ in the package parameterization; NG and
ING then share the same `rate` and the same weak-prior limit $\Pi_0$ (Chapter~11
and `?dIndependent_Normal_Gamma`). The posterior Gamma **shape** after seeing
data still updates from this prior `shape` slot; only the **limiting** law
$\Pi_0$ is stated in Claim~3 below.

**Claim 3:** (limiting `dIndependent_Normal_Gamma()` posterior as prior weight vanishes).  
Pass **`dIndependent_Normal_Gamma(ps$mu, ps$Sigma, shape = ps$shape + p/2, rate =
ps$rate)`** with $p$ as above. Let $n_{\mathrm{prior}}\to 0^{+}$ as in Claim~2.
Under hypotheses **(1)--(3)** of the Theorem in Section~3.3.4.1, the ING
posterior $\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\cdot\mid y)$ converges
weakly to the **same** Normal--Gamma law $\Pi_0(\cdot\mid y)$ as in Claim~2. The
limiting hyperparameters and marginal moments match Claim~2 **(i)--(vii)**
exactly:
\[
\mu_{\Pi_{0}}=\hat\beta,
\qquad
\Sigma_{0,\Pi_{0}}=G^{-1},
\qquad
a_{\Pi_{0}}=\frac{n_w+1}{2},
\qquad
b_{\Pi_{0}}=\frac{1}{2}\,\frac{n_w-1}{n_w-p}\,\mathrm{RSS}_w,
\]
\[
\mathbb{E}_{\Pi_{0}}[\beta\mid y]=\hat\beta,
\qquad
\mathbb{E}_{\Pi_{0}}[\sigma^2\mid y]=\frac{\mathrm{RSS}_w}{n_w-p},
\qquad
\mathrm{Cov}_{\Pi_{0}}(\beta\mid y)
=
\frac{\mathrm{RSS}_w}{n_w-p}\,G^{-1}.
\]
At finite $n_{\mathrm{prior}}$, ING is not the conjugate NG factorization, so
closed forms differ from Claim~1; Claim~3 concerns **only** the
$n_{\mathrm{prior}}\to 0^{+}$ limit. A proof that weak convergence implies
convergence of these moments for the ING path (uniform integrability and
Billingsley) is in **Appendix~6.2**.

The numbered conjecture items and `Ex_lmb` discussion below predate the formal
Theorem in Section~3.3.4.1; they align with Claim~3 and that Theorem when the
hypotheses hold.

**Package calibration (conjugate fragment).** With `shape` and `rate` from
`n_{\mathrm{prior}}` only,
`compute_gaussian_prior()` sets `rate` so that, **in the conjugate Normal--Gamma
update** (Section 3.3.1--3.3.2),
\[
\frac{b_n}{a_n-1}=\frac{S_{\mathrm{marg}}}{n_w-p},
\]
with the same $S_{\mathrm{marg}}$ and $n_w-p$. The **marginal**
$E(\sigma^2\mid y)$ under `dIndependent_Normal_Gamma()` need not equal this
closed form at finite $n_{\mathrm{prior}}$ because $\beta$ and $\tau$ are not
conjugate in that parameterization. **Claim~3** and Section~3.3.4.1 nonetheless
give the weak limit and limiting moments as $n_{\mathrm{prior}}\to 0^{+}$; the
numbered list below is retained as motivation for the `Ex_lmb` pairing.

**Conjecture (marginal limits for the `Ex_lmb` pairing).** The joint prior
factorizations differ between `dNormal_Gamma()` and `dIndependent_Normal_Gamma()`
(Chapter 11 and Section 2.3 of this appendix). It is nonetheless plausible
that, **when the two fits use the same `ps` as in `Ex_lmb.R` above** (same
formula, data, and scalar `pwt`, but different `pfamily`), the **marginal** posterior summaries align in the weak-prior
limit $\mathrm{pwt}\to 0^+$ (equivalently $n_{\mathrm{prior}}\to 0^+$) with
$n_w>p$ held fixed:

1. **Coefficients:** $E(\beta\mid y)\to \hat\beta$ (the weighted least-squares
   coefficient vector).
2. **Dispersion:** $E(\sigma^2\mid y)\to \mathrm{RSS}_w/(n_w-p)$, the classical
   weighted residual variance.
3. **Coefficient uncertainty:** the marginal posterior covariance of $\beta$
   converges to
   \[
   \frac{\mathrm{RSS}_w}{n_w-p}\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
   \]
   i.e.\ the usual weighted least-squares covariance.

In words: the example workflow compares conjugate `dNormal_Gamma()` on `ps` to
independent `dIndependent_Normal_Gamma()` on the same `ps` (with **`shape =
ps$shape + p/2`** for ING as above). Finite-$n_{\mathrm{prior}}$ joint posteriors
need not coincide; **Claim~3** and Section~3.3.4.1 formalize the shared weak
limit $\Pi_0$ and marginal moment limits under the stated hypotheses.

#### 3.3.4.1 Theorem, equivalence of limits, and moments

The objects below are defined relative to the weighted Gaussian sampling model
in Sections 3.1--3.3, `Prior_Setup()` / `compute_gaussian_prior()` calibration,
and the `dNormal_Gamma()` construction in this package. Let
$\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\cdot\mid y)$ denote the posterior law
of $(\beta,\sigma^2)$ under `dNormal_Gamma()` with `Prior_Setup()` hyperparameters
indexed by scalar $n_{\mathrm{prior}}>0$. Let
$\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\cdot\mid y)$ denote the posterior
under `dIndependent_Normal_Gamma()` with the same `ps` except **`shape =
ps$shape + p/2`** with $p=\dim(\beta)$. All limits are taken as
$n_{\mathrm{prior}}\to 0^+$ with data $(y,X)$ fixed and with the `Prior_Setup`
construction held fixed aside from $n_{\mathrm{prior}}$. The weak limit of the
posterior as $n_{\mathrm{prior}}\to 0^+$ is denoted $\Pi_{0}(\cdot\mid y)$ (subscript
$0$, not $\infty$, because the limit is taken as $n_{\mathrm{prior}}\to 0$).

**1) Theorem (equivalence of posterior limits; target statement).**  

Consider the default prior specifications produced by `Prior_Setup()` for the
weighted Gaussian model in Sections~3.1--3.3, and assume:

1. $X^{\mathsf T}W_{\mathrm{obs}}X$ is positive definite.

2. $n_w>p$.

3. $\mathrm{RSS}_w>0$.

Then there exists a **unique** probability measure $\Pi_{0}(\cdot\mid y)$ on
$\mathcal{S}$ such that, as $n_{\mathrm{prior}}\to 0^+$,
\[
\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\cdot\mid y)
\Rightarrow \Pi_{0}(\cdot\mid y),
\qquad
\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}(\cdot\mid y)
\Rightarrow \Pi_{0}(\cdot\mid y).
\]
Specifically, the shared limiting distribution $\Pi_{0}(\cdot\mid y)$ is a
**Normal--Gamma** law in the package parameterization
`dNormal_Gamma(mu, Sigma_0, shape, rate)` with hyperparameters
\[
\mu=\hat\beta,
\qquad
\Sigma_0=\left(X^{\mathsf T}W_{\mathrm{obs}}X\right)^{-1},
\qquad
\texttt{shape}=\frac{n_w+1}{2},
\qquad
\texttt{rate}=\frac{1}{2}\,\frac{n_w-1}{n_w-p}\,\mathrm{RSS}_w,
\]
where $\hat\beta$ is the weighted least-squares coefficient vector,
$\mathrm{RSS}_w$ is the weighted residual sum of squares from Section~3.1, and
$n_w$ and $p$ are as above. These values are the limits of the conjugate
posterior hyperparameters
$(\mu_{n_{\mathrm{prior}}},\Sigma_{0,n_{\mathrm{prior}}},a_n,b_n)$ in
Sections~3.3.1--3.3.2 as $n_{\mathrm{prior}}\to 0^+$ (with
$S_{\mathrm{marg}}\to\mathrm{RSS}_w$).

**2) Corollary (expectations and covariance).**  

Assume the **same** hypotheses **(1)--(3)** as in the Theorem in **1)**.

**Normal--Gamma path.** For every admissible $n_{\mathrm{prior}}>0$, the
posterior $\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\cdot\mid y)$ is the
**conjugate** Normal--Gamma law in the package parameterization. Hence
$\mathbb{E}[\beta\mid y]$, $\mathbb{E}[\sigma^{2}\mid y]$, and
$\mathrm{Cov}(\beta\mid y)$ are **explicit elementary functions** of the
posterior hyperparameters $(\mu_{n_{\mathrm{prior}}},\Sigma_{0,n_{\mathrm{prior}}},
a_{n_{\mathrm{prior}}},b_{n_{\mathrm{prior}}})$ from Sections~3.3.1--3.3.2 (the
same closed forms as in any textbook conjugate linear-model treatment). Under
**(1)--(3)**, those hyperparameters converge to the limiting
$\Pi_{0}$-hyperparameters from **1)** as $n_{\mathrm{prior}}\to 0^{+}$. The
corresponding posterior moments therefore converge to the **plug-in** values at
the limit **immediately**, by **continuity** of those formulas on the admissible
region, without appealing to weak convergence or uniform integrability for the NG
case. See **Appendix~6.1**.

**Independent Normal--Gamma path.** At finite $n_{\mathrm{prior}}$, the ING
posterior is **not** the same conjugate NG factorization, so the same shortcut
is unavailable. Weak convergence from **1)** does **not** automatically pass
unbounded functionals; **Appendix~6.2** completes the argument for the ING
moments using uniform integrability (via uniform $L^{r}$ bounds) and Billingsley.

Then
\[
\mathbb{E}_{\Pi^{(\mathrm{NG})}}[\beta\mid y]\to \mathbb{E}_{\Pi_{0}}[\beta\mid y],
\quad
\mathbb{E}_{\Pi^{(\mathrm{ING})}}[\beta\mid y]\to \mathbb{E}_{\Pi_{0}}[\beta\mid y],
\]
\[
\mathbb{E}_{\Pi^{(\mathrm{NG})}}[\sigma^2\mid y]\to \mathbb{E}_{\Pi_{0}}[\sigma^2\mid y],
\quad
\mathbb{E}_{\Pi^{(\mathrm{ING})}}[\sigma^2\mid y]\to \mathbb{E}_{\Pi_{0}}[\sigma^2\mid y],
\]
and
\[
\mathrm{Cov}_{\Pi^{(\mathrm{NG})}}(\beta\mid y)\to \mathrm{Cov}_{\Pi_{0}}(\beta\mid y),
\quad
\mathrm{Cov}_{\Pi^{(\mathrm{ING})}}(\beta\mid y)\to \mathrm{Cov}_{\Pi_{0}}(\beta\mid y),
\]
where $\mathrm{Cov}(\beta\mid y)=\mathbb{E}[\beta\beta^{\mathsf T}\mid y]-
\mathbb{E}[\beta\mid y]\mathbb{E}[\beta^{\mathsf T}\mid y]$. In particular,
the **NG** and **ING** limiting expectations and limiting posterior covariance
matrices **coincide** with each other and with the moments of $\Pi_{0}$.

**Remark.** For **NG**, **Appendix~6.1** records the logic (closed forms +
continuity). For **ING**, **Appendix~6.2** supplies the measure-theoretic step
from weak convergence to moments.

Proofs of the Theorem are deferred; supporting derivations for the kernels and
normalizing constants appear in **Appendix~5.1--5.2** at the end of this
chapter. A fully explicit description of $\Pi_{0}$ from the **ING**
factorization alone is a separate step.

#### 3.3.4.2 Un-normalized log posteriors and lemmas (toward the ING limit)

The definitions of $g_{\mathrm{NG}}$, $g_{\mathrm{ING}}$, $g_{0}$, $\Delta g$, and **Lemmas~1--2** (pointwise log kernels and normalizing constants) are given in **Appendix 5.1** at the end of this chapter. That appendix records the weighted Gaussian notation, the `Prior_Setup` calibrations, the un-normalized log-posterior kernels, and the proofs toward the ING weak-prior limit.

#### 3.3.4.3 Marginal $\beta$ after integrating $\tau$

The integration of $\tau$ out of the NG and ING kernels, the Gamma conditional rates $\rho_{\bullet}$, the explicit marginal difference and its factored form, the log-ratio identities, the Fubini identity for $\int_{\mathbb{R}^{p}} D_{n_{\mathrm{prior}}}(\beta)\,\mathrm{d}\beta$, and the surrounding limits are in **Appendix 5.2**.

#### 3.3.5 `dGamma()` prior on coefficients

For Gaussian models, `dGamma()` supplies a Gamma prior on precision for
coefficients in a different parameterization than the residual-precision Gamma
above. Mapping `Prior_Setup()` outputs (`shape`, `rate`, `dispersion`, `Sigma`)
into `dGamma()` workflows is deferred here; see `?dGamma`, `?Prior_Setup`, and
Chapter A02 for constructor slots and simulation entry points.

### 3.4 Returned objects and model mapping

For Gaussian workflows, `Prior_Setup()` returns a matched set:
- `Sigma_0` (dispersion-free, for `dNormal_Gamma()`)
- `shape`, `rate` (Gamma prior on residual precision)
- `dispersion` and `Sigma` (coefficient-scale companion quantities)

Always pass **`Sigma_0 = ps$Sigma_0`** into `dNormal_Gamma()` (not `ps$Sigma`):

```r
ps <- Prior_Setup(formula, data, family = gaussian(), n_prior = n0)
lmb(
  formula,
  data,
  pfamily = dNormal_Gamma(
    ps$mu,
    Sigma_0 = ps$Sigma_0,
    shape     = ps$shape,
    rate      = ps$rate
  )
)
```

For matrix-input workflows, Chapter A02: `prior_list` for `rNormalGamma_reg()`
uses `Sigma = ps$Sigma_0` (dispersion-free slot) together with `shape` and
`rate`.

### 3.5 Variants (`pwt` vector, `sd`, user `Sigma`)

With vector `pwt`, $\Sigma_{\mathrm{pre}}$ uses the Hadamard scaling from
Section 2.4; coefficient blending remains elementwise
$(1-\mathrm{pwt}_j)\hat\beta_j + \mathrm{pwt}_j\mu_j$.

With **`sd`**, `pwt` is derived from $\operatorname{diag}(V_0)$ (Section 2.4),
and coefficient-scale $\Sigma_{\mathrm{pre}}$ is passed as `Sigma` into
`compute_gaussian_prior()`. The helper then reconciles returned
`Sigma`, `Sigma_0`, and `dispersion` consistently.

With user-supplied `Sigma`, the helper enforces `Sigma_0 = Sigma / dispersion`
(using calibrated or user-provided `dispersion`), so `dNormal` and
`dNormal_Gamma` parameterizations stay aligned. When only `dispersion` is
provided, returned `Sigma` is `Sigma_0 * dispersion`.

## 4. Other Models With Unknown Dispersion

This appendix focused on the `Prior_Setup()` $\to$ `dNormal_Gamma()` Gaussian
path. Other compound specifications reuse parts of the same output but expect
**different** slots:

- **`dIndependent_Normal_Gamma()`** takes **undivided** `ps$Sigma` (coefficient
  scale) plus **`shape = ps$shape + p/2`** with **`p = ncol(ps$x)`** and the same `ps$rate`;
  do **not** substitute `Sigma_0` for `Sigma` there.
- **`dGamma()`** with fixed coefficients uses `shape`, `rate`, and typically
  `coefficients` from the prior list for residual precision updates.

See Chapter A02 for `prior_list` templates (`rNormalGamma_reg()` vs
`rindepNormalGamma_reg()`), Chapter 11 Section 3.3 for the independent
Normal--Gamma model, and Chapter A07 for truncated-Gamma / envelope extensions.

## 5. Appendix: Detailed derivations for Section 3.3.4.2 and Section 3.3.4.3

### 5.1 Un-normalized log posteriors and lemmas (toward the ING limit)


**References for notation.** The weighted exponential-family framework (weights
$w_i$, canonical parameters, and $b(\theta_i)$) is set out in **Chapter~05**
(foundations of GLMs and exponential families). The same weighted
log-likelihood notation feeds **Chapter~11**, which gives the general weighted
log-likelihood $\ell(\beta,\phi)$, the **Gaussian** special case in precision
form $\ell(\beta,\tau)$, and the **log-prior** and **log-posterior** for the
conjugate Normal--Gamma model (Section~3.2) and the independent Normal--Gamma
model (Section~3.3). Below, $\ell$, $\mathrm{RSS}(\beta)$, and the prior pieces
match Chapter~11 **exactly**; only the map from `Prior_Setup()` objects to
$(\mu_0,\Sigma_0,a_0,b_0)$ is specialized as in Section~3.3.4 of this appendix.

Work on the **precision** scale $\tau=\phi^{-1}=\sigma^{-2}>0$ and coefficients
$\beta\in\mathbb{R}^{p}$. Let $W=\mathrm{diag}(w_1,\ldots,w_n)$ denote the
observation weights (so $W=W_{\mathrm{obs}}$ in Sections~3.1--3.3),
\[
G \;=\; X^{\mathsf T}W_{\mathrm{obs}}X,
\]
and define the weighted residual sum of squares
\[
\mathrm{RSS}(\beta)
=
\sum_{i=1}^{n} w_i\bigl(y_i-x_i^{\mathsf T}\beta\bigr)^2
=
\bigl(y-X\beta\bigr)^{\mathsf T}W\bigl(y-X\beta\bigr),
\]
and $n_w=\sum_{i=1}^{n} w_i$, as in Chapter~11 (Gaussian special case). Let
$\hat\beta$ denote the weighted least-squares coefficient vector,
$\mathrm{RSS}_w=\mathrm{RSS}(\hat\beta)$, and let $\mu\in\mathbb{R}^{p}$ denote the
**prior mean** returned by `Prior_Setup()` (fixed as $n_{\mathrm{prior}}$
varies).

**`Prior_Setup` strength in terms of $n_{\mathrm{prior}}$ and $n_w$.**  
When `Prior_Setup()` is driven by a scalar **`n_prior`** on the same scale as
$n_w$ (i.e.\ `n_effective` $=n_w$), the implied scalar prior weight satisfies
\[
\begin{aligned}
\mathrm{pwt}
&=
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}, \\
1-\mathrm{pwt}
&=
\frac{n_w}{n_{\mathrm{prior}}+n_w}, \\
\frac{\mathrm{pwt}}{1-\mathrm{pwt}}
&=
\frac{n_{\mathrm{prior}}}{n_w}.
\end{aligned}
\]
Then the **calibrated** objects in Sections~3.3.1--3.3.3 can be written as
functions of $(n_{\mathrm{prior}},n_w)$:
\[
\begin{aligned}
S_{\mathrm{marg}}(n_{\mathrm{prior}})
&=
\mathrm{RSS}_w
\\
&\quad{}
+
\frac{n_{\mathrm{prior}}}{n_{\mathrm{prior}}+n_w}\,
(\hat\beta-\mu)^{\mathsf T}G(\hat\beta-\mu), \\
a_0(n_{\mathrm{prior}})
&=
\frac{n_{\mathrm{prior}}+1}{2}, \\
b_0(n_{\mathrm{prior}})
&=
\frac{1}{2}\,(n_{\mathrm{prior}}+p-1)\,
\frac{S_{\mathrm{marg}}(n_{\mathrm{prior}})}{n_w-p}, \\
\Sigma_0(n_{\mathrm{prior}})^{-1}
&=
\frac{n_{\mathrm{prior}}}{n_w}\,G
\quad\text{(dispersion-free \texttt{Sigma\_0} slot from Prior\_Setup).}
\end{aligned}
\]
\[
\begin{aligned}
\mathrm{dispersion}(n_{\mathrm{prior}})
&=
\frac{S_{\mathrm{marg}}(n_{\mathrm{prior}})}{n_w-p}, \\
\Sigma_{\mathrm{coeff}}(n_{\mathrm{prior}})
&=
\Sigma_0(n_{\mathrm{prior}})\,\mathrm{dispersion}(n_{\mathrm{prior}}) \\
&=
\frac{n_w}{n_{\mathrm{prior}}}\,
\frac{S_{\mathrm{marg}}(n_{\mathrm{prior}})}{n_w-p}\,G^{-1}
\quad\text{(\texttt{Sigma} coefficient scale for \texttt{dNormal()} / ING).}
\end{aligned}
\]
\[
\begin{aligned}
a_0^{\mathrm{ING}}(n_{\mathrm{prior}})
&=
a_0(n_{\mathrm{prior}})+\frac{p}{2} \\
&=
\frac{n_{\mathrm{prior}}+p+1}{2}.
\end{aligned}
\]
All of these are indexed by $n_{\mathrm{prior}}$ so that the dependence on the
weak-prior limit $n_{\mathrm{prior}}\to 0^+$ is explicit.

**Weighted Gaussian log-likelihood (Chapter~11, Section~2.2).**  
Up to an additive constant **not** depending on $(\beta,\tau)$,
\[
\ell(\beta,\tau;y,n_w)
=
\frac{\tau}{2}\,\mathrm{RSS}(\beta)
-\frac{n_w}{2}\,\log\!\left(\frac{2\pi}{\tau}\right),
\]
equivalently the precision-form display obtained by expanding
$\log(2\pi/\tau)$ and absorbing $\log(2\pi)$ into the constant.

**Definitions (un-normalized log posteriors as functions of
$(\beta,\tau,n_{\mathrm{prior}},n_w)$).**  
Absorb into a single additive constant (allowed to depend on $n_{\mathrm{prior}}$
but not on $(\beta,\tau)$) every term that does not depend on $(\beta,\tau)$.

The **conjugate Normal--Gamma** kernel is the Chapter~11 Section~3.2 log-posterior with
the substitutions above (including $\frac{p}{2}\log\tau$ from
$|\tau\Sigma_0^{-1}|^{1/2}$ in $\beta\mid\tau$):
\[
\begin{aligned}
g_{\mathrm{NG}}(\beta,\tau;n_{\mathrm{prior}},n_w)
&=
\ell(\beta,\tau;y,n_w)
+\bigl(a_0(n_{\mathrm{prior}})-1\bigr)\log\tau
-b_0(n_{\mathrm{prior}})\tau \\
&\quad
-\frac{\tau}{2}\,
(\beta-\mu)^{\mathsf T}\Sigma_0(n_{\mathrm{prior}})^{-1}(\beta-\mu)
+\frac{p}{2}\,\log\tau
\\
&=
\frac{\tau}{2}\,\mathrm{RSS}(\beta)
-\frac{n_w}{2}\,\log\!\left(\frac{2\pi}{\tau}\right)
+\frac{n_{\mathrm{prior}}+p-1}{2}\,\log\tau \\
&\quad
-\frac{1}{2}\,(n_{\mathrm{prior}}+p-1)\,
\frac{S_{\mathrm{marg}}(n_{\mathrm{prior}})}{n_w-p}\,\tau \\
&\quad
-\frac{\tau}{2}\,\frac{n_{\mathrm{prior}}}{n_w}\,
(\beta-\mu)^{\mathsf T}G(\beta-\mu).
\end{aligned}
\]

The **independent Normal--Gamma** kernel is the Chapter~11 Section~3.3 log-posterior
with the same $\ell$ and $b_0(n_{\mathrm{prior}})$, but **coefficient-scale**
normal covariance $\Sigma_{\mathrm{coeff}}(n_{\mathrm{prior}})$ and shape
$a_0^{\mathrm{ING}}(n_{\mathrm{prior}})$:
\[
\begin{aligned}
g_{\mathrm{ING}}(\beta,\tau;n_{\mathrm{prior}},n_w)
&=
\ell(\beta,\tau;y,n_w) \\
&\quad
-\frac{1}{2}\,
(\beta-\mu)^{\mathsf T}\Sigma_{\mathrm{coeff}}(n_{\mathrm{prior}})^{-1}(\beta-\mu)
\\
&\quad
+\bigl(a_0^{\mathrm{ING}}(n_{\mathrm{prior}})-1\bigr)\log\tau
-b_0(n_{\mathrm{prior}})\tau
\\
&=
\frac{\tau}{2}\,\mathrm{RSS}(\beta)
-\frac{n_w}{2}\,\log\!\left(\frac{2\pi}{\tau}\right)
+\frac{n_{\mathrm{prior}}+p-1}{2}\,\log\tau \\
&\quad
-\frac{1}{2}\,(n_{\mathrm{prior}}+p-1)\,
\frac{S_{\mathrm{marg}}(n_{\mathrm{prior}})}{n_w-p}\,\tau \\
&\quad
-\frac{1}{2}\,\frac{n_{\mathrm{prior}}(n_w-p)}{n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
(\beta-\mu)^{\mathsf T}G(\beta-\mu).
\end{aligned}
\]

Define the **difference kernel** (only terms that do not cancel) by
\[
\Delta g(\beta,\tau;n_{\mathrm{prior}},n_w)
\;=\;
g_{\mathrm{ING}}(\beta,\tau;n_{\mathrm{prior}},n_w)
-
g_{\mathrm{NG}}(\beta,\tau;n_{\mathrm{prior}},n_w).
\]
Then $\ell(\beta,\tau;y,n_w)$, $-b_0(n_{\mathrm{prior}})\tau$, and all
$\log\tau$ terms **cancel**, because
$a_0^{\mathrm{ING}}(n_{\mathrm{prior}})-a_0(n_{\mathrm{prior}})=p/2$ offsets
the extra $\frac{p}{2}\log\tau$ in $g_{\mathrm{NG}}$. What remains is the
mismatch between the **coefficient-scale** normal ($\Sigma_{\mathrm{coeff}}$)
and the **precision-scaled** conjugate normal ($\tau\Sigma_0$):
\[
\begin{aligned}
\Delta g(\beta,\tau;n_{\mathrm{prior}},n_w)
&=
-\frac{1}{2}\,
(\beta-\mu)^{\mathsf T}\Sigma_{\mathrm{coeff}}(n_{\mathrm{prior}})^{-1}(\beta-\mu)
+\frac{\tau}{2}\,
(\beta-\mu)^{\mathsf T}\Sigma_0(n_{\mathrm{prior}})^{-1}(\beta-\mu)
\\
&=
\frac{1}{2}\,
(\beta-\mu)^{\mathsf T}
\Bigl(
\tau\,\Sigma_0(n_{\mathrm{prior}})^{-1}
-\Sigma_{\mathrm{coeff}}(n_{\mathrm{prior}})^{-1}
\Bigr)
(\beta-\mu)
\\
&=
\frac{n_{\mathrm{prior}}}{2}\,
(\beta-\mu)^{\mathsf T}G(\beta-\mu)
\left(
\frac{\tau}{n_w}
-\frac{n_w-p}{n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}
\right).
\end{aligned}
\]

The **limiting Normal--Gamma** kernel $g_{0}$ is obtained from
$g_{\mathrm{NG}}$ by **the same expressions**, with
$n_{\mathrm{prior}}=0$ and hence $S_{\mathrm{marg}}(0)=\mathrm{RSS}_w$ (no
$\mathrm{pwt}$-weighted penalty). Equivalently,
$g_{0}(\beta,\tau;n_w)=g_{\mathrm{NG}}(\beta,\tau;0,n_w)$ as long as the
calibration formulas above are evaluated at $n_{\mathrm{prior}}=0$.
\[
\begin{aligned}
g_{0}(\beta,\tau;n_w)
&=
\ell(\beta,\tau;y,n_w)
+\bigl(a_0(0)-1\bigr)\log\tau
-b_0(0)\tau \\
&\quad
-\frac{\tau}{2}\,
(\beta-\mu)^{\mathsf T}\Sigma_0(0)^{-1}(\beta-\mu)
+\frac{p}{2}\,\log\tau
\\
&=
\ell(\beta,\tau;y,n_w)
+\frac{p-1}{2}\,\log\tau
-b_0(0)\tau
\\
&=
\frac{\tau}{2}\,\mathrm{RSS}(\beta)
-\frac{n_w}{2}\,\log\!\left(\frac{2\pi}{\tau}\right)
+\frac{0-1}{2}\,\log\tau
+\frac{p}{2}\,\log\tau \\
&\quad
-\frac{1}{2}\,(0+p-1)\,
\frac{S_{\mathrm{marg}}(0)}{n_w-p}\,\tau \\
&\quad
-\frac{\tau}{2}\,\frac{0}{n_w}\,
(\beta-\mu)^{\mathsf T}G(\beta-\mu) \\
&=
\frac{\tau}{2}\,\mathrm{RSS}(\beta)
-\frac{n_w}{2}\,\log\!\left(\frac{2\pi}{\tau}\right)
+\frac{p-1}{2}\,\log\tau
-\frac{1}{2}\,(p-1)\,
\frac{\mathrm{RSS}_w}{n_w-p}\,\tau
\end{aligned}
\]
with $S_{\mathrm{marg}}(0)=\mathrm{RSS}_w$. One line above, $\Sigma_0(0)^{-1}=0$
so the $(\beta-\mu)$ quadratic vanishes and
$(a_0(0)-1)\log\tau+\frac{p}{2}\log\tau
=-\tfrac{1}{2}\log\tau+\frac{p}{2}\log\tau
=\frac{p-1}{2}\log\tau$.
The last line of the expanded block is identically zero; the final equality
drops it and replaces $\frac{0-1}{2}\log\tau+\frac{p}{2}\log\tau$ by
$\frac{p-1}{2}\log\tau$, and $(0+p-1)\,S_{\mathrm{marg}}(0)/(n_w-p)$ by
$(p-1)\,\mathrm{RSS}_w/(n_w-p)$. Here
$b_0(0)=\tfrac{1}{2}(p-1)\mathrm{RSS}_w/(n_w-p)$. The
Theorem in Section~3.3.4.1 records the **posterior** Normal--Gamma limit in
`dNormal_Gamma(mu, Sigma_0, shape, rate)` notation (different algebraic
packaging of the same weak limit); proofs relate that parameterization to this
$n_{\mathrm{prior}}=0$ evaluation.

Here $g_{0}$ depends on **$n_w$** and the data through
$(\hat\beta,\mathrm{RSS}_w,G)$ but **not** on $n_{\mathrm{prior}}$; it is the
log-kernel obtained by evaluating the $g_{\mathrm{NG}}$ template at
$n_{\mathrm{prior}}=0$. The same expression may be rearranged into the
**collapsed** Normal--Gamma joint density (Chapter~11 Section~3.2 prior display); proofs
may use either form up to $(\beta,\tau)$-independent constants.

Thus $\exp\{g_{\mathrm{NG}}\}$, $\exp\{g_{\mathrm{ING}}\}$, and $\exp\{g_{0}\}$
are **un-normalized** kernels on $(\beta,\tau)$ for the respective laws under
this convention. Chapter~A07 truncates $\tau$ only to implement accept--reject
sampling; the integrals and limits below use the **full** domain
$\mathbb{R}^{p}\times(0,\infty)$ appropriate to the continuous posteriors.

### 5.2 Marginal $\beta$ after integrating $\tau$: Gamma conditionals and the NG/ING difference


Fix $n_{\mathrm{prior}}>0$ and suppress it in $b_0$, $\Sigma_0$, and when convenient
write $S_{\mathrm{marg}}=S_{\mathrm{marg}}(n_{\mathrm{prior}})$. Define
$Q_{G}(\beta)=(\beta-\mu)^{\mathsf T}G(\beta-\mu)$.

From Appendix~5.1,
$\Sigma_{\mathrm{coeff}}(n_{\mathrm{prior}})
=\frac{n_w}{n_{\mathrm{prior}}}\,
\frac{S_{\mathrm{marg}}(n_{\mathrm{prior}})}{n_w-p}\,G^{-1}$, hence
\[
\Sigma_{\mathrm{coeff}}(n_{\mathrm{prior}})^{-1}
=
\frac{n_{\mathrm{prior}}}{n_w}\,
\frac{n_w-p}{S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
G.
\]
Therefore every ING prior quadratic appearing below can be written **without**
$\Sigma_{\mathrm{coeff}}$ as
\[
\frac{1}{2}\,
(\beta-\mu)^{\mathsf T}\Sigma_{\mathrm{coeff}}(n_{\mathrm{prior}})^{-1}(\beta-\mu)
=
\frac{n_{\mathrm{prior}}(n_w-p)}
{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
Q_{G}(\beta),
\]
a **scalar** multiple of $Q_{G}(\beta)$ with coefficient
$\frac{n_{\mathrm{prior}}(n_w-p)}{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}
=O(n_{\mathrm{prior}})$ as $n_{\mathrm{prior}}\to 0^{+}$ because
$S_{\mathrm{marg}}(n_{\mathrm{prior}})\to\mathrm{RSS}_w>0$.

Work from the **explicit** blocks for $g_{\mathrm{NG}}$ and $g_{\mathrm{ING}}$ in
Appendix~5.1 (same $\ell$, same $b_0$, same $\log\tau$ power, different
$\beta$ factor). Expand
$-\frac{n_w}{2}\log\bigl(\frac{2\pi}{\tau}\bigr)
=\frac{n_w}{2}\log\tau-\frac{n_w}{2}\log(2\pi)$.
Define the **Gamma shape** in $\tau\mid\beta$ as an explicit function of
$n_{\mathrm{prior}}$ (it is **not** constant as the prior weight varies):
\[
s(n_{\mathrm{prior}})
:=
\frac{n_w+n_{\mathrm{prior}}+p+1}{2},
\qquad
s(n_{\mathrm{prior}})-1
=
\frac{n_w+n_{\mathrm{prior}}+p-1}{2}.
\]
In particular
$s(n_{\mathrm{prior}})=\frac{n_w+p+1}{2}+\frac{n_{\mathrm{prior}}}{2}$, so the
shape moves linearly with $n_{\mathrm{prior}}$. Algebraically,
$s(n_{\mathrm{prior}})
=a_0(n_{\mathrm{prior}})+\frac{n_w}{2}+\frac{p}{2}$
with $a_0(n_{\mathrm{prior}})=(n_{\mathrm{prior}}+1)/2$ from Appendix~5.1: the
$\log\tau$ power adds contributions from $\ell$, from the prior on $\tau$, and
from the $\beta\mid\tau$ normalizing factor in the NG factorization (and the ING
shape $a_0^{\mathrm{ING}}$ is chosen so the **same** $s(n_{\mathrm{prior}})$
appears). Below, shorthand $\tau^{s-1}$, $\Gamma(s)$, and $\rho^{-s}$ always mean
$\tau^{s(n_{\mathrm{prior}})-1}$,
$\Gamma\bigl(s(n_{\mathrm{prior}})\bigr)$, and $\rho^{-s(n_{\mathrm{prior}})}$.

For every $\beta$, the $\tau$-dependent part of $g_{\mathrm{NG}}$ is
\[
\frac{\tau}{2}\mathrm{RSS}(\beta)
-b_0\tau
-\frac{\tau}{2}\,\frac{n_{\mathrm{prior}}}{n_w}\,Q_{G}(\beta)
+\bigl(s(n_{\mathrm{prior}})-1\bigr)\log\tau
=
-\tau\,\rho_{\mathrm{NG}}(\beta)+\bigl(s(n_{\mathrm{prior}})-1\bigr)\log\tau,
\]
hence
\[
\rho_{\mathrm{NG}}(\beta)
=
b_0+\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)-\frac{1}{2}\mathrm{RSS}(\beta).
\]
The $\tau$-dependent part of $g_{\mathrm{ING}}$ is
$-\tau\,\rho_{\mathrm{ING}}(\beta)+\bigl(s(n_{\mathrm{prior}})-1\bigr)\log\tau$ with
\[
\rho_{\mathrm{ING}}(\beta)
=
b_0-\frac{1}{2}\mathrm{RSS}(\beta).
\]
Substitute
$b_0=b_0(n_{\mathrm{prior}})
=\frac{1}{2}(n_{\mathrm{prior}}+p-1)\,
S_{\mathrm{marg}}(n_{\mathrm{prior}})/(n_w-p)$
(Appendix~5.1). Then the rates depend on $n_{\mathrm{prior}}$ **only** through
$b_0(n_{\mathrm{prior}})$, through $S_{\mathrm{marg}}(n_{\mathrm{prior}})$ inside
$b_0$, and through the explicit $\frac{n_{\mathrm{prior}}}{2n_w}$ multiplying
$Q_{G}(\beta)$ in $\rho_{\mathrm{NG}}$:
\[
\begin{aligned}
\rho_{\mathrm{ING}}(\beta;n_{\mathrm{prior}})
&=
\frac{(n_{\mathrm{prior}}+p-1)\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}
{2(n_w-p)}
-\frac{1}{2}\mathrm{RSS}(\beta), \\
\rho_{\mathrm{NG}}(\beta;n_{\mathrm{prior}})
&=
\frac{(n_{\mathrm{prior}}+p-1)\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}
{2(n_w-p)}
+\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)
-\frac{1}{2}\mathrm{RSS}(\beta).
\end{aligned}
\]
In particular
$\rho_{\mathrm{NG}}(\beta;n_{\mathrm{prior}})
-\rho_{\mathrm{ING}}(\beta;n_{\mathrm{prior}})
=\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)$ for all $\beta$.

**Separately**, the remaining **$\tau$-free** term in $g_{\mathrm{ING}}$ is the ING
prior quadratic. Using
$\Sigma_{\mathrm{coeff}}(n_{\mathrm{prior}})^{-1}$ above, it equals
\[
-\frac{n_{\mathrm{prior}}(n_w-p)}
{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
Q_{G}(\beta),
\]
from the third and fourth lines of the $g_{\mathrm{ING}}$ display (Appendix~5.1).

Let $C_{\mathrm{absorb}}$ denote any additive constant **allowed in
Appendix~5.1** that does not depend on $(\beta,\tau)$. The only other
$\tau$-free contribution from $\ell$ is $-\frac{n_w}{2}\log(2\pi)$. Define the
**common** scalar
\[
K
:=
\exp\Bigl\{C_{\mathrm{absorb}}-\frac{n_w}{2}\log(2\pi)\Bigr\}.
\]
Then, for every $\beta$ for which $\rho_{\mathrm{NG}}(\beta)>0$ and
$\rho_{\mathrm{ING}}(\beta)>0$ (equivalently the Gamma integrals below converge),
\[
\begin{aligned}
\exp\{g_{\mathrm{NG}}(\beta,\tau)\}
&=
K\,\tau^{\,s(n_{\mathrm{prior}})-1}\exp\{-\tau\,\rho_{\mathrm{NG}}(\beta)\}, \\
\exp\{g_{\mathrm{ING}}(\beta,\tau)\}
&=
K\,\tau^{\,s(n_{\mathrm{prior}})-1}\exp\{-\tau\,\rho_{\mathrm{ING}}(\beta)\}\,
\exp\Bigl\{
-\frac{n_{\mathrm{prior}}(n_w-p)}
{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
Q_{G}(\beta)
\Bigr\}.
\end{aligned}
\]

**Gamma conditional for $\tau\mid\beta,y$.**  
With the usual **rate** parameterization
$\mathrm{Gamma}\bigl(s(n_{\mathrm{prior}}),\rho\bigr)$ having density
$\propto\tau^{\,s(n_{\mathrm{prior}})-1}e^{-\rho\tau}$ on $\tau>0$,
\[
\tau\mid\beta,y\sim\mathrm{Gamma}\bigl(s(n_{\mathrm{prior}}),\rho_{\bullet}(\beta)\bigr),
\qquad
\bullet\in\{\mathrm{NG},\mathrm{ING}\},
\]
with the **same** shape $s(n_{\mathrm{prior}})$ in both models and rates
$\rho_{\mathrm{NG}}$, $\rho_{\mathrm{ING}}$ above. In particular
$\rho_{\mathrm{NG}}(\beta)-\rho_{\mathrm{ING}}(\beta)
=\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)$.

**Exact $\tau$ integrals (unnormalized $\beta$ marginals).**  
Since
$\int_{0}^{\infty}\tau^{\,s(n_{\mathrm{prior}})-1}e^{-\rho\tau}\,\mathrm{d}\tau
=\Gamma\bigl(s(n_{\mathrm{prior}})\bigr)/\rho^{\,s(n_{\mathrm{prior}})}$
for $\rho>0$,
\[
\begin{aligned}
\widetilde p_{\mathrm{NG}}(\beta\mid y)
&:=
\int_{0}^{\infty}
\exp\{g_{\mathrm{NG}}(\beta,\tau)\}\,\mathrm{d}\tau
=
K\,\Gamma\bigl(s(n_{\mathrm{prior}})\bigr)\,
\rho_{\mathrm{NG}}(\beta)^{-\,s(n_{\mathrm{prior}})}, \\
\widetilde p_{\mathrm{ING}}(\beta\mid y)
&:=
\int_{0}^{\infty}
\exp\{g_{\mathrm{ING}}(\beta,\tau)\}\,\mathrm{d}\tau
=
K\,\Gamma\bigl(s(n_{\mathrm{prior}})\bigr)\,
\rho_{\mathrm{ING}}(\beta)^{-\,s(n_{\mathrm{prior}})}\,
\exp\Bigl\{
-\frac{n_{\mathrm{prior}}(n_w-p)}
{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
Q_{G}(\beta)
\Bigr\}.
\end{aligned}
\]
Here $\widetilde p_{\mathrm{NG}}(\cdot\mid y)$ is the usual **multivariate $t$**
kernel obtained by expanding $\mathrm{RSS}(\beta)$ into a quadratic in $\beta$; the
ING marginal differs at finite $n_{\mathrm{prior}}$ by the explicit Gaussian
factor.

**Exact difference.**  
The **same** prefactor $K\,\Gamma\bigl(s(n_{\mathrm{prior}})\bigr)$ cancels in ratios; the **difference** of the
two $\tau$ integrals is the pointwise identity obtained by substituting the
rates $\rho_{\bullet}(\beta;n_{\mathrm{prior}})$ above, written below directly in
terms of $A_{n_{\mathrm{prior}}}$, $\mathrm{RSS}(\beta)$, and $Q_{G}(\beta)$
(recall $s(n_{\mathrm{prior}})=(n_w+n_{\mathrm{prior}}+p+1)/2$). Write
$A_{n_{\mathrm{prior}}}:=\frac{(n_{\mathrm{prior}}+p-1)\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}
{2(n_w-p)}$ for the $b_0$-part common to both rates. Then
\[
\begin{aligned}
{}&
\int_{0}^{\infty}\exp\{g_{\mathrm{ING}}(\beta,\tau)\}\,\mathrm{d}\tau
-
\int_{0}^{\infty}\exp\{g_{\mathrm{NG}}(\beta,\tau)\}\,\mathrm{d}\tau \\
&\qquad
=
K\,\Gamma\bigl(s(n_{\mathrm{prior}})\bigr)\Biggl[
\Biggl(
A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}
\Biggr)^{-\,s(n_{\mathrm{prior}})}
\exp\Bigl\{
-\frac{n_{\mathrm{prior}}(n_w-p)}
{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
Q_{G}(\beta)
\Bigr\}
\\
&\qquad\qquad
-
\Biggl(
A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}
+\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)
\Biggr)^{-\,s(n_{\mathrm{prior}})}
\Biggr].
\end{aligned}
\]
The same difference admits the **factored** form obtained by pulling the **NG**
rate power
$\bigl(A_{n_{\mathrm{prior}}}-\frac{1}{2}\mathrm{RSS}(\beta)
+\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)\bigr)^{-\,s(n_{\mathrm{prior}})}$
to the front:
\[
\begin{aligned}
{}&
\int_{0}^{\infty}\exp\{g_{\mathrm{ING}}(\beta,\tau)\}\,\mathrm{d}\tau
-
\int_{0}^{\infty}\exp\{g_{\mathrm{NG}}(\beta,\tau)\}\,\mathrm{d}\tau \\
&\qquad
=
K\,\Gamma\bigl(s(n_{\mathrm{prior}})\bigr)\,
\Biggl(
A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}
+\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)
\Biggr)^{-\,s(n_{\mathrm{prior}})}
\\
&\qquad\qquad
\times
\Biggl[
\Biggl(
\frac{
A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}
+\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)
}{
A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}
}
\Biggr)^{\,s(n_{\mathrm{prior}})}
\exp\Bigl\{
-\frac{n_{\mathrm{prior}}(n_w-p)}
{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
Q_{G}(\beta)
\Bigr\}
-1
\Biggr].
\end{aligned}
\]
The outer power is the **NG** rate
$A_{n_{\mathrm{prior}}}-\frac{1}{2}\mathrm{RSS}(\beta)
+\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)$; the inner bracket compares
that rate to the **ING** piece
$A_{n_{\mathrm{prior}}}-\frac{1}{2}\mathrm{RSS}(\beta)$ and the ING Gaussian
correction, and is the form most convenient for tracking
$n_{\mathrm{prior}}\to 0^{+}$ (the ratio tends to $1$ and the exponential to
$1$, so the bracket tends to $0$).

Here $\mathrm{RSS}(\beta)$ and $Q_{G}(\beta)$ do **not** depend on
$n_{\mathrm{prior}}$; every $n_{\mathrm{prior}}$ dependence in the bracket is
through $s(n_{\mathrm{prior}})$ (hence also through $\Gamma\bigl(s(n_{\mathrm{prior}})\bigr)$),
through $S_{\mathrm{marg}}(n_{\mathrm{prior}})$ in
$A_{n_{\mathrm{prior}}}$ and in the Gaussian exponent, and through the extra
$\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)$ in the **NG** rate only.

**Log ratio** (no proportionality ambiguity), in the same expanded notation with
$A_{n_{\mathrm{prior}}}$ fixed as above:
\[
\begin{aligned}
\log \widetilde p_{\mathrm{ING}}(\beta\mid y)
-
\log \widetilde p_{\mathrm{NG}}(\beta\mid y)
&=
-\frac{n_{\mathrm{prior}}(n_w-p)}
{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
Q_{G}(\beta) \\
&\quad
-
s(n_{\mathrm{prior}})\,\log
\frac{
A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}
}{
A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}
+\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)
}.
\end{aligned}
\]
Since the ratio in the logarithm equals
$1/\bigl(1+\frac{n_{\mathrm{prior}}}{2n_w}\,
Q_{G}(\beta)/\bigl(A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}\bigr)\bigr)$,
\[
\begin{aligned}
\log \widetilde p_{\mathrm{ING}}(\beta\mid y)
-
\log \widetilde p_{\mathrm{NG}}(\beta\mid y)
&=
-\frac{n_{\mathrm{prior}}(n_w-p)}
{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
Q_{G}(\beta) \\
&\quad
+
s(n_{\mathrm{prior}})\,\log\Biggl(
1+\frac{n_{\mathrm{prior}}}{2n_w}\,
\frac{Q_{G}(\beta)}
{A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}}
\Biggr)
\end{aligned}
\]
whenever $A_{n_{\mathrm{prior}}}-\frac{\mathrm{RSS}(\beta)}{2}>0$. As
$n_{\mathrm{prior}}\to 0^{+}$,
$S_{\mathrm{marg}}(n_{\mathrm{prior}})\to\mathrm{RSS}_w$ and the Gaussian
prefactor satisfies
$\exp\bigl\{-\frac{n_{\mathrm{prior}}(n_w-p)}
{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,Q_{G}(\beta)\bigr\}\to 1$
for each fixed $\beta$; equivalently
$\Sigma_{\mathrm{coeff}}(n_{\mathrm{prior}})^{-1}
=\frac{n_{\mathrm{prior}}(n_w-p)}{n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,G\to 0$
entrywise. Also $\Sigma_0(n_{\mathrm{prior}})^{-1}
=\frac{n_{\mathrm{prior}}}{n_w}\,G\to 0$, and the ratio
$\bigl(A_{n_{\mathrm{prior}}}-\frac{\mathrm{RSS}(\beta)}{2}\bigr)
/\bigl(A_{n_{\mathrm{prior}}}-\frac{\mathrm{RSS}(\beta)}{2}
+\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)\bigr)\to 1$
pointwise over $\beta$, matching the **same** weak limit recorded in
Section~3.3.4.1.

**$\beta$-integral of the displayed marginal difference.**  
For each $\beta$, define the **pointwise** difference (same quantity as in the
display immediately above)
\[
D_{n_{\mathrm{prior}}}(\beta)
:=
\int_{0}^{\infty}\exp\{g_{\mathrm{ING}}(\beta,\tau)\}\,\mathrm{d}\tau
-
\int_{0}^{\infty}\exp\{g_{\mathrm{NG}}(\beta,\tau)\}\,\mathrm{d}\tau
=
K\,\Gamma\bigl(s(n_{\mathrm{prior}})\bigr)\Biggl[
\Biggl(
A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}
\Biggr)^{-\,s(n_{\mathrm{prior}})}
\exp\Bigl\{
-\frac{n_{\mathrm{prior}}(n_w-p)}
{2\,n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}\,
Q_{G}(\beta)
\Bigr\}
-
\Biggl(
A_{n_{\mathrm{prior}}}
-\frac{\mathrm{RSS}(\beta)}{2}
+\frac{n_{\mathrm{prior}}}{2n_w}\,Q_{G}(\beta)
\Biggr)^{-\,s(n_{\mathrm{prior}})}
\Biggr],
\]
the same bracket as in the first **Exact difference** display above; whenever
the Gamma--rate factorization is valid (equivalently: wherever both rate
combinations are $>0$); elsewhere $D_{n_{\mathrm{prior}}}(\beta)$ is still
defined by the $\tau$ integrals on the left-hand side.

To integrate $D_{n_{\mathrm{prior}}}(\beta)$ over $\beta$, use **Fubini's
theorem** on the **difference** of the two nonnegative joint kernels. Under the
structural conditions in Section~3.3.4.1, $\exp\{g_{\mathrm{ING}}\}$ and
$\exp\{g_{\mathrm{NG}}\}$ are integrable on
$\mathcal{D}=\mathbb{R}^{p}\times(0,\infty)$ (proper **ING** and **NG**
posteriors), hence
$\exp\{g_{\mathrm{ING}}\}-\exp\{g_{\mathrm{NG}}\}\in L^{1}(\mathcal{D})$ because
$|\exp\{g_{\mathrm{ING}}\}-\exp\{g_{\mathrm{NG}}\}|
\le
\exp\{g_{\mathrm{ING}}\}+\exp\{g_{\mathrm{NG}}\}$. Therefore
\[
\begin{aligned}
\int_{\mathbb{R}^{p}} D_{n_{\mathrm{prior}}}(\beta)\,\mathrm{d}\beta
&=
\int_{\mathbb{R}^{p}}
\int_{0}^{\infty}
\Bigl(
\exp\{g_{\mathrm{ING}}(\beta,\tau)\}
-
\exp\{g_{\mathrm{NG}}(\beta,\tau)\}
\Bigr)\,\mathrm{d}\tau\,\mathrm{d}\beta \\
&=
\iint_{\mathcal{D}}
\exp\{g_{\mathrm{ING}}(\beta,\tau)\}\,\mathrm{d}\beta\,\mathrm{d}\tau
-
\iint_{\mathcal{D}}
\exp\{g_{\mathrm{NG}}(\beta,\tau)\}\,\mathrm{d}\beta\,\mathrm{d}\tau \\
&=
Z_{\mathrm{ING}}(n_{\mathrm{prior}})
-
Z_{\mathrm{NG}}(n_{\mathrm{prior}}).
\end{aligned}
\]
The first equality is Fubini applied to the integrable difference; the second is
the definition of $Z_{\mathrm{ING}}$ and $Z_{\mathrm{NG}}$ in Lemma~2 below.
**No further manipulation of the closed form in $K$, $\rho_{\bullet}$, and
$Q_{G}$ is needed:** the $\beta$-integral of that bracketed expression is exactly
$Z_{\mathrm{ING}}-Z_{\mathrm{NG}}$ because it is the same
$D_{n_{\mathrm{prior}}}(\beta)$ integrated over $\beta$.

By Lemma~2,
$Z_{\mathrm{ING}}(n_{\mathrm{prior}})-Z_{\mathrm{NG}}(n_{\mathrm{prior}})\to 0$ as
$n_{\mathrm{prior}}\to 0^{+}$, hence
\[
\int_{\mathbb{R}^{p}} D_{n_{\mathrm{prior}}}(\beta)\,\mathrm{d}\beta\longrightarrow 0.
\]
(**Total variation**
$\int_{\mathbb{R}^{p}}|D_{n_{\mathrm{prior}}}(\beta)|\,\mathrm{d}\beta$ is a
different quantity from this signed integral.)

**Lemma 1 (pointwise log kernels).**  
Assume the structural conditions in the Theorem (Section~3.3.4.1). Then, as
$n_{\mathrm{prior}}\to 0^{+}$:

1. **Normal--Gamma versus the limit.** For each fixed $(\beta,\tau)$ in the
   domain,
   \[
   g_{\mathrm{NG}}(\beta,\tau;n_{\mathrm{prior}},n_w)\;\longrightarrow\;g_{0}(\beta,\tau;n_w)
   \qquad\text{as}\qquad n_{\mathrm{prior}}\to 0^{+}.
   \]

2. **Independent Normal--Gamma versus Normal--Gamma.** For each fixed
   $(\beta,\tau)$ in the domain,
   $\Delta g(\beta,\tau;n_{\mathrm{prior}},n_w)\to 0$ (equivalently
   $g_{\mathrm{ING}}-g_{\mathrm{NG}}\to 0$), as
   $n_{\mathrm{prior}}\to 0^{+}$.
   Consequently, for each such $(\beta,\tau)$,
   \[
   \exp\{g_{\mathrm{ING}}(\beta,\tau;n_{\mathrm{prior}},n_w)\}\;\longrightarrow\;\exp\{g_{0}(\beta,\tau;n_w)\}
   \qquad\text{as}\qquad n_{\mathrm{prior}}\to 0^{+}.
   \]

*Proof.* **1)** By construction,
$g_{\mathrm{NG}}(\beta,\tau;n_{\mathrm{prior}},n_w)$ uses the same closed form
as $g_{0}(\beta,\tau;n_w)$ once $n_{\mathrm{prior}}=0$ and
$S_{\mathrm{marg}}(0)=\mathrm{RSS}_w$ (Appendix~5.1); letting
$n_{\mathrm{prior}}\to 0^{+}$ sends the calibrated hyperparameters to that
evaluation, so
$g_{\mathrm{NG}}(\beta,\tau;n_{\mathrm{prior}},n_w)\to g_{0}(\beta,\tau;n_w)$
pointwise for each fixed $(\beta,\tau)$. (Equivalently, this is the conjugate
Normal--Gamma weak-prior limit.)

**2)** By definition of $\Delta g$,
\[
g_{\mathrm{ING}}(\beta,\tau;n_{\mathrm{prior}},n_w)
=
g_{\mathrm{NG}}(\beta,\tau;n_{\mathrm{prior}},n_w)
+
\Delta g(\beta,\tau;n_{\mathrm{prior}},n_w),
\]
hence
\[
\exp\{g_{\mathrm{ING}}(\beta,\tau;n_{\mathrm{prior}},n_w)\}
=
\exp\{g_{\mathrm{NG}}(\beta,\tau;n_{\mathrm{prior}},n_w)\}\,
\exp\{\Delta g(\beta,\tau;n_{\mathrm{prior}},n_w)\}.
\]
From the explicit formula for $\Delta g$ in Appendix~5.1,
\[
\Delta g(\beta,\tau;n_{\mathrm{prior}},n_w)
=
\frac{n_{\mathrm{prior}}}{2}\,
(\beta-\mu)^{\mathsf T}G(\beta-\mu)
\left(
\frac{\tau}{n_w}
-\frac{n_w-p}{n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}
\right),
\]
which is a product of $n_{\mathrm{prior}}$ and a factor that stays finite for
fixed $(\beta,\tau)$ and fixed data as $n_{\mathrm{prior}}\to 0^{+}$ (since
$S_{\mathrm{marg}}(n_{\mathrm{prior}})\to\mathrm{RSS}_w>0$). Therefore
$\Delta g(\beta,\tau;n_{\mathrm{prior}},n_w)\to 0$ and
$\exp\{\Delta g(\beta,\tau;n_{\mathrm{prior}},n_w)\}\to 1$. Combining with
**1)**,
\[
\exp\{g_{\mathrm{ING}}\}
=
\exp\{g_{\mathrm{NG}}\}\,\exp\{\Delta g\}
\;\longrightarrow\;
\exp\{g_{0}\}\cdot 1
=
\exp\{g_{0}\}
\]
pointwise for each fixed $(\beta,\tau)$.

**Lemma 2 (normalizing constants).**  
Let $\mathcal{D}=\mathbb{R}^{p}\times(0,\infty)$ and define
\[
\begin{aligned}
Z_{\mathrm{NG}}(n_{\mathrm{prior}})
&=
\iint_{\mathcal{D}}
\exp\{g_{\mathrm{NG}}(\beta,\tau;n_{\mathrm{prior}},n_w)\}\,
\mathrm{d}\beta\,\mathrm{d}\tau, \\
Z_{\mathrm{ING}}(n_{\mathrm{prior}})
&=
\iint_{\mathcal{D}}
\exp\{g_{\mathrm{ING}}(\beta,\tau;n_{\mathrm{prior}},n_w)\}\,
\mathrm{d}\beta\,\mathrm{d}\tau, \\
Z_{0}
&=
\iint_{\mathcal{D}}
\exp\{g_{0}(\beta,\tau;n_w)\}\,
\mathrm{d}\beta\,\mathrm{d}\tau.
\end{aligned}
\]
Under the structural conditions in the Theorem (Section~3.3.4.1), as
$n_{\mathrm{prior}}\to 0^{+}$,
\[
Z_{\mathrm{NG}}(n_{\mathrm{prior}})\longrightarrow Z_{0},
\qquad
Z_{\mathrm{ING}}(n_{\mathrm{prior}})\longrightarrow Z_{0}.
\]

*Proof.* **Finiteness.** For every $n_{\mathrm{prior}}>0$ admitted by
`compute_gaussian_prior()`, the **NG** and **ING** posteriors are **proper**
probability measures on $\mathcal{D}$, so $Z_{\mathrm{NG}}(n_{\mathrm{prior}})$ and
$Z_{\mathrm{ING}}(n_{\mathrm{prior}})$ are finite and positive. The limiting
Normal--Gamma law $\Pi_{0}$ (Section~3.3.4.1) is a proper distribution on
$(\beta,\tau)$ under the same assumptions ($n_w>p$, $\mathrm{RSS}_w>0$, $G$
positive definite), so $Z_{0}<\infty$.

**Convergence $Z_{\mathrm{NG}}\to Z_{0}$.** The kernel
$\exp\{g_{\mathrm{NG}}(\beta,\tau;n_{\mathrm{prior}},n_w)\}$ is, up to an overall
multiplicative constant absorbed in $g_{\mathrm{NG}}$, the **conjugate**
Normal--Gamma posterior density with hyperparameters
$(\mu,\Sigma_0(n_{\mathrm{prior}}),a_n,b_n)$ that vary **continuously** in
$n_{\mathrm{prior}}$ on $(0,\varepsilon)$. The normalizing constant of a
proper Normal--Gamma$(\mu,\Sigma_0,a,b)$ law on $\mathcal{D}$ is a **continuous**
function of $(\mu,\Sigma_0,a,b)$ on the open parameter set where the density is
integrable. As $n_{\mathrm{prior}}\to 0^{+}$, those hyperparameters converge to
the limiting values implicit in $g_{0}$, hence
$Z_{\mathrm{NG}}(n_{\mathrm{prior}})\to Z_{0}$.

**Convergence $Z_{\mathrm{ING}}\to Z_{\mathrm{NG}}$ (hence $Z_{\mathrm{ING}}\to
Z_{0}$).** This is the substantive step. Write
\[
Z_{\mathrm{ING}}-Z_{\mathrm{NG}}
=
\iint_{\mathcal{D}}
\exp\{g_{\mathrm{NG}}\}\,\bigl(\exp\{\Delta g\}-1\bigr)\,
\mathrm{d}\beta\,\mathrm{d}\tau.
\]
Let $\pi_{\mathrm{NG}}$ denote the **NG** posterior probability measure with
density $\exp\{g_{\mathrm{NG}}\}/Z_{\mathrm{NG}}$. Appendix~5.1 gives
\[
\Delta g(\beta,\tau;n_{\mathrm{prior}},n_w)
=
n_{\mathrm{prior}}\,H(\beta,\tau;n_{\mathrm{prior}}),
\quad
H(\beta,\tau;n_{\mathrm{prior}})
=
\frac{1}{2}\,
(\beta-\mu)^{\mathsf T}G(\beta-\mu)
\left(
\frac{\tau}{n_w}
-\frac{n_w-p}{n_w\,S_{\mathrm{marg}}(n_{\mathrm{prior}})}
\right),
\]
with $S_{\mathrm{marg}}(n_{\mathrm{prior}})\to\mathrm{RSS}_w>0$. Under
$\pi_{\mathrm{NG}}$, the coefficients have **finite moments of all orders**:
conditionally on $\tau$, $\beta$ is Gaussian; marginally, $\tau$ is Gamma. In
particular,
\[
M(n_{\mathrm{prior}})
:=
\mathbb{E}_{\pi_{\mathrm{NG}}}\!\left[
\bigl\|G^{1/2}(\beta-\mu)\bigr\|^{2}\,(1+\tau)
\right]
\]
is finite for each admissible $n_{\mathrm{prior}}>0$ and remains bounded as
$n_{\mathrm{prior}}\to 0^{+}$ (continuous dependence of the conjugate posterior on
$n_{\mathrm{prior}}$). Hence, with a constant $C$ depending only on
$(n_w,p,\mathrm{RSS}_w)$ and the fixed prior mean $\mu$,
\[
|H(\beta,\tau;n_{\mathrm{prior}})|
\le
C\,\bigl\|G^{1/2}(\beta-\mu)\bigr\|^{2}\,(1+\tau)
\qquad\text{for all }(\beta,\tau)\in\mathcal{D}
\]
and all sufficiently small $n_{\mathrm{prior}}$. Therefore
$\mathbb{E}_{\pi_{\mathrm{NG}}}[|H|]\le C'\,M(n_{\mathrm{prior}})$ is finite and
$O(1)$ as $n_{\mathrm{prior}}\to 0^{+}$.

For every $(\beta,\tau)$ one has the **pointwise** bound
\[
\bigl|\exp\{g_{\mathrm{NG}}\}\,(\exp\{\Delta g\}-1)\bigr|
\le
\exp\{g_{\mathrm{ING}}\}+\exp\{g_{\mathrm{NG}}\},
\]
because if $\Delta g\ge 0$ then $0\le \exp\{\Delta g\}-1\le \exp\{\Delta g\}$ and
$\exp\{g_{\mathrm{NG}}\}(\exp\{\Delta g\}-1)\le \exp\{g_{\mathrm{ING}}\}$, while
if $\Delta g<0$ then $|\exp\{\Delta g\}-1|=1-\exp\{\Delta g\}\le 1$ and
$\exp\{g_{\mathrm{NG}}\}\,|\exp\{\Delta g\}-1|\le \exp\{g_{\mathrm{NG}}\}$.

On $\{|\Delta g|\le 1\}$, the mean-value theorem gives
$|\exp\{\Delta g\}-1|\le e\,|\Delta g|=e\,n_{\mathrm{prior}}\,|H|$, hence
\[
\iint_{\{|\Delta g|\le 1\}}
\exp\{g_{\mathrm{NG}}\}\,\bigl|\exp\{\Delta g\}-1\bigr|
\le
e\,n_{\mathrm{prior}}\,Z_{\mathrm{NG}}\,
\mathbb{E}_{\pi_{\mathrm{NG}}}\!\bigl[|H|\bigr].
\]
On $\{|\Delta g|>1\}$, using the pointwise bound above,
\[
\iint_{\{|\Delta g|>1\}}
\exp\{g_{\mathrm{NG}}\}\,\bigl|\exp\{\Delta g\}-1\bigr|
\le
\iint_{\{|\Delta g|>1\}} \exp\{g_{\mathrm{ING}}\}
\;+\;
\iint_{\{|\Delta g|>1\}} \exp\{g_{\mathrm{NG}}\}.
\]
Here $\Delta g=n_{\mathrm{prior}}H$, so $\{|\Delta g|>1\}=\{|H|>1/n_{\mathrm{prior}}\}$.
For any $n_{\mathrm{prior}}>0$, $\mathbb{E}_{\pi_{\mathrm{NG}}}[H^{2}]$ and
$\mathbb{E}_{\pi_{\mathrm{ING}}}[H^{2}]$ are finite (same moment argument as for
$M(n_{\mathrm{prior}})$ above). Chebyshev's inequality gives
\[
\pi_{\mathrm{NG}}\bigl(|H|>1/n_{\mathrm{prior}}\bigr)
\le
n_{\mathrm{prior}}^{2}\,\mathbb{E}_{\pi_{\mathrm{NG}}}\bigl[H^{2}\bigr],
\qquad
\pi_{\mathrm{ING}}\bigl(|H|>1/n_{\mathrm{prior}}\bigr)
\le
n_{\mathrm{prior}}^{2}\,\mathbb{E}_{\pi_{\mathrm{ING}}}\bigl[H^{2}\bigr],
\]
so
\[
\iint_{\{|\Delta g|>1\}} \exp\{g_{\mathrm{NG}}\}
=
Z_{\mathrm{NG}}\,\pi_{\mathrm{NG}}\bigl(|H|>1/n_{\mathrm{prior}}\bigr)
\le
n_{\mathrm{prior}}^{2}\,Z_{\mathrm{NG}}\,\mathbb{E}_{\pi_{\mathrm{NG}}}\bigl[H^{2}\bigr]
\]
and, with the same event,
\[
\iint_{\{|\Delta g|>1\}} \exp\{g_{\mathrm{ING}}\}
\le
n_{\mathrm{prior}}^{2}\,Z_{\mathrm{ING}}\,\mathbb{E}_{\pi_{\mathrm{ING}}}\bigl[H^{2}\bigr].
\]
Combining the $\{|\Delta g|\le 1\}$ and $\{|\Delta g|>1\}$ pieces,
\[
\bigl|Z_{\mathrm{ING}}-Z_{\mathrm{NG}}\bigr|
\le
e\,n_{\mathrm{prior}}\,Z_{\mathrm{NG}}\,
\mathbb{E}_{\pi_{\mathrm{NG}}}\!\bigl[|H|\bigr]
\;+\;
n_{\mathrm{prior}}^{2}\Bigl(
Z_{\mathrm{ING}}\,\mathbb{E}_{\pi_{\mathrm{ING}}}\bigl[H^{2}\bigr]
+
Z_{\mathrm{NG}}\,\mathbb{E}_{\pi_{\mathrm{NG}}}\bigl[H^{2}\bigr]
\Bigr)
=
O(n_{\mathrm{prior}})
\]
as $n_{\mathrm{prior}}\to 0^{+}$ (the $n_{\mathrm{prior}}$ term from the first
integral dominates the $n_{\mathrm{prior}}^{2}$ tail). Since $Z_{\mathrm{NG}}$
is bounded away from $0$, $Z_{\mathrm{ING}}/Z_{\mathrm{NG}}\to 1$ and
$Z_{\mathrm{ING}}-Z_{\mathrm{NG}}\to 0$. Together with $Z_{\mathrm{NG}}\to Z_{0}$,
\[
Z_{\mathrm{ING}}(n_{\mathrm{prior}})
=
\bigl(Z_{\mathrm{ING}}-Z_{\mathrm{NG}}\bigr)
+
Z_{\mathrm{NG}}
\longrightarrow
0+Z_{0}
=
Z_{0}.
\]

Together, Lemmas~1--2 supply one route to the **ING** weak limit: normalized
posterior densities are $\exp\{g_{\bullet}\}/Z_{\bullet}$; convergence of
$Z_{\mathrm{NG}}$ and $Z_{\mathrm{ING}}$ as above and pointwise convergence of
kernels yield convergence of the normalized densities (e.g.\ Scheffe's lemma for
total variation).

## 6. Appendix: Moment limits in the Corollary (Section 3.3.4.1)

Claims~1 and~2 in Section~3.3.2 are **conjugate** Normal--Gamma identities and
limits as $n_{\mathrm{prior}}\to 0^{+}$; standard algebra and continuity in the
hyperparameters suffice, so this appendix does not expand them further.
**Claim~3** (Section~3.3.4) matches the same limiting $\Pi_0$ for the ING
posterior; the nontrivial step is passing from weak convergence to moments when
closed-form conjugate formulas are unavailable at finite $n_{\mathrm{prior}}$.
That argument is **Appendix~6.2**.

### 6.1 Normal--Gamma: conjugate formulas and continuity (Claims~1--2)

Under $\Pi^{(\mathrm{NG})}_{n_{\mathrm{prior}}}(\cdot\mid y)$, the pair
$(\beta,\tau)$ with $\tau=1/\sigma^{2}$ has the **conjugate** Normal--Gamma
structure (Chapter~11 and Sections~3.3.1--3.3.2 here). Consequently:

- $\mathbb{E}[\beta\mid y]$ and $\mathrm{Cov}(\beta\mid y)$ are **standard
  closed forms**: the marginal of $\beta$ is **multivariate $t$** with location
  $\mu_{n_{\mathrm{prior}}}$, scale matrix determined by $\Sigma_{0,n_{\mathrm{prior}}}$
  and the posterior Gamma shape for $\tau$, and degrees of freedom
  $\nu=2\,s-p=n_{w}+n_{\mathrm{prior}}+1$ in the `Prior_Setup` calibration
  (**Appendix~5.2**, with $s=(n_{w}+n_{\mathrm{prior}}+p+1)/2$). In particular
  $\mathbb{E}[\beta\mid y]=\mu_{n_{\mathrm{prior}}}$ and, for $\nu>2$,
  $\mathrm{Cov}(\beta\mid y)=\frac{\nu}{\nu-2}$ times the scale matrix in that
  marginal $t$ density (the standard conjugate Normal--Gamma expression in terms
  of $\Sigma_{0,n_{\mathrm{prior}}}$ and the residual-precision update).
- $\mathbb{E}[\sigma^{2}\mid y]=\mathbb{E}[\tau^{-1}\mid y]$ is the
  **inverse-Gamma** (equivalently $\tau$--Gamma) posterior mean: **rate divided
  by (shape minus one)** in the package's rate parameterization, whenever the
  posterior shape exceeds one.

Each of these expressions is an **elementary function** of the posterior
hyperparameters $\mu_{n_{\mathrm{prior}}}$, $\Sigma_{0,n_{\mathrm{prior}}}$,
and the posterior Gamma shape and rate parameters from Sections~3.3.1--3.3.2. Under hypotheses **(1)--(3)** of the
Theorem in Section~3.3.4.1, those hyperparameters converge to the limiting
$\Pi_{0}$ values as $n_{\mathrm{prior}}\to 0^{+}$. On the region where the
closed forms are well-defined (e.g.\ $\nu>2$ for second moments of $\beta$,
posterior shape${}>1$ for $\mathbb{E}[\sigma^{2}]$), the map from hyperparameters
to moments is **continuous**. Hence the **NG** limits in the Corollary are the
plug-in moments at $\Pi_{0}$, **without** using weak convergence or uniform
integrability on this path.

### 6.2 Independent Normal--Gamma: proof of Claim~3 (uniform integrability)

This subsection justifies convergence of posterior means and covariances along
the ING path as in **Claim~3** and the Corollary in Section~3.3.4.1. At finite
$n_{\mathrm{prior}}$, the ING posterior does **not** coincide with
the conjugate NG factorization, so the same **single** list of textbook NG
formulas does not apply directly. To pass from **weak convergence**
$\Pi^{(\mathrm{ING})}_{n_{\mathrm{prior}}}\Rightarrow\Pi_{0}$ to convergence of
$\mathbb{E}[\beta\mid y]$, $\mathbb{E}[\sigma^{2}\mid y]$, and
$\mathrm{Cov}(\beta\mid y)$, use **uniform integrability** and Billingsley.

**Lemma 6.1.** If $\sup_{n}\mathbb{E}\bigl[\|X_{n}\|^{r}\bigr]<\infty$ for some
$r>1$, then $\{X_{n}\}$ is uniformly integrable.

*Proof.* $\mathbb{E}\bigl[\|X_{n}\|
\mathbf{1}\{\|X_{n}\|>M\}\bigr]
\le M^{1-r}\,\mathbb{E}\bigl[\|X_{n}\|^{r}\bigr]$. $\square$

**Lemma 6.2.** If $r>2$, the same hypothesis implies uniform integrability of
the products $\beta_{j}\beta_{k}$ needed for $\mathrm{Cov}(\beta\mid y)$ (e.g.\
$|\beta_{j}\beta_{k}|\le \|\beta\|^{2}$ and Lyapunov/Cauchy--Schwarz).

**Theorem 6.3** (Billingsley, *Convergence of Probability Measures*, 2nd ed.,
Theorem~25.11 and Corollary). If $X_{n}\Rightarrow X$ and $\{X_{n}\}$ is uniformly
integrable, then $\mathbb{E}[X_{n}]\to\mathbb{E}[X]$; the same applies to
coordinates and, with Lemma~6.2, to second moments and covariances when the limit
has finite moments.

**Uniform $L^{r}$ bounds for $\beta$ under ING.** Let $\widetilde p_{\bullet}$
be the unnormalized $\beta$ marginals after integrating $\tau$ (Appendix~5.2).
Since $Q_{G}(\beta)=(\beta-\mu)^{\mathsf T}G(\beta-\mu)\ge 0$,
$\widetilde p_{\mathrm{ING}}\le \widetilde p_{\mathrm{NG}}$ pointwise, hence
$Z_{\mathrm{ING}}\le Z_{\mathrm{NG}}$ and
\[
\pi_{\mathrm{ING}}(\beta)
\le
\frac{Z_{\mathrm{NG}}}{Z_{\mathrm{ING}}}\,\pi_{\mathrm{NG}}(\beta).
\]
For nonnegative $h$,
$\mathbb{E}_{\Pi^{(\mathrm{ING})}}[h(\beta)]
\le \frac{Z_{\mathrm{NG}}}{Z_{\mathrm{ING}}}\,
\mathbb{E}_{\Pi^{(\mathrm{NG})}}[h(\beta)]$.
Appendix~5.2 gives $Z_{\mathrm{NG}}/Z_{\mathrm{ING}}\to 1$, so this ratio is
bounded on $(0,\varepsilon]$ for $\varepsilon$ small. Together with **Appendix~6.1**,
the NG marginal $t$ moments of $\|\beta\|^{r}$ are bounded uniformly in
$n_{\mathrm{prior}}$ on that interval when $\nu>r$; hence the same holds for ING.

**Dispersion.** For $\sigma^{2}=\tau^{-1}$, use the same program: finite
posterior moments of $\sigma^{2}$ of fixed order along $n_{\mathrm{prior}}\in(0,\varepsilon]$
follow from **continuity** of the Gamma/inverse-Gamma moment maps in the
posterior shape and rate under **(1)--(3)**, together with the domination step
above where it applies, or direct finiteness at each $n_{\mathrm{prior}}$ plus
the boundedness implied by convergence of laws to $\Pi_{0}$ with finite
variance.

**Conclusion.** Lemmas~6.1--6.2 supply uniform integrability for the ING
functionals; Theorem~6.3 and weak convergence from the Theorem in Section~3.3.4.1
yield the **ING** limits in the Corollary. Combined with **Appendix~6.1**, both paths
agree with the moments of $\Pi_{0}$.

