---
title: "Chapter 01: Getting started"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 01: Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

# Chapter 01: Getting started

This chapter fits one model end to end. The data is `lme4::sleepstudy`:
reaction times for 18 subjects over ten days of restricted sleep. Everyone
slows down; the interesting question is by how much, and how much that
varies from person to person.

# The data

```{r data}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)
str(dat)
```

`Days` is centered. Centering is not required, but it makes the intercept
mean reaction time at the *average* day rather than at day zero, which
gives the random intercept and random slope a cleaner interpretation and
reduces their correlation — useful here, because the package assumes they
are uncorrelated.

# The formula

```{r formula}
form <- Reaction ~ Days_c + (1 + Days_c || Subject)
```

Read this as: reaction time depends on days of deprivation, and both the
baseline level and the per-day effect vary by subject.

The double bar `||` is required. `lme4` would let you write
`(1 + Days_c | Subject)`, which additionally estimates a correlation
between each subject's intercept and slope. **lmebayes** does not support
that; it needs a diagonal random-effect covariance so the Block 2 update
factorizes across coefficients. Chapter 02 explains why, and Chapter 13
discusses what you give up.

# Inspecting the design

Before fitting anything, `model_setup()` shows how the formula is turned
into matrices.

```{r design}
design <- model_setup(form, data = dat)
design
```

Three things to check here.

**Level 1** is the within-group model. Each subject gets their own
intercept and their own `Days_c` slope, so there are two group-effect
coefficients per subject.

**Level 2** is the across-group model. Both lines read `~ 1`, meaning
neither coefficient has group-level predictors: subjects are exchangeable,
and all we estimate at the population level is the average intercept and
the average slope. Chapter 04 shows what happens when you add school-level
covariates.

**Rank** matters. `Full-rank D_j: 18 of 18 groups` says every subject has
enough distinct `Days_c` values to identify their own slope. Subjects who
fail this check are flagged rather than removed: they stay in the fit, but
in the deficient direction they carry no likelihood information and are
effectively drawn from their prior. What they *are* excluded from is the
across-group rank check and the prior calibration. Chapter 04 shows a
dataset where this happens, and
`vignette("Chapter-B08", package = "lmebayesCore")` gives the conditions in
full.

# Setting up priors

`Prior_Setup_GLMM()` fits a reference `lmer()` model and uses it to
calibrate priors on a scale that matches the data.

```{r prior}
ps <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01)
```

`pop.pwt = 0.01` sets the *prior weight*: the population-level prior should
carry about as much information as one percent of the observations. Small
values mean weak priors and results close to `lmer()`; larger values pull
the population effects toward the prior mean. Chapter 03 unpacks every
field of this object.

```{r pfamily}
pf <- pfamily_list(ps)
pf
```

`pfamily_list()` converts the calibration into one prior object per
random-effect coefficient — here, one for the intercept and one for
`Days_c`. Both are `dNormal`, meaning the corresponding variance component
`tau^2_k` is held fixed at the `lmer` estimate. Chapter 07 shows how to
sample it instead.

# Fitting

```{r fit, results = "hide"}
set.seed(1)
fit <- lmerb(
  form,
  data             = dat,
  pfamily_list     = pf,
  dispersion_ranef = ps$group.dispersion,
  n                = 500L,
  progbar          = FALSE
)
```

`n = 500` is the number of posterior draws. `dispersion_ranef` supplies the
observation variance `sigma^2`; here it is the fixed `lmer` estimate.

```{r fit-print}
fit
```

The header says `500 draws, exact iid`. Because `sigma^2` is fixed and
both `tau^2_k` are known, the joint posterior over the subject
coefficients and the population coefficients is exactly multivariate
normal, so the sampler draws from it directly. There is no burn-in, no
autocorrelation, and no convergence diagnostic to run — the 500 draws are
genuinely independent. Chapter 06 covers the case where Gibbs sweeps are
needed.

The table compares three quantities per parameter:

| Column | Meaning |
|---|---|
| `fixef.mode` | Exact posterior mean, computed by iterated conditional modes before any sampling |
| `fixef.means` | Average of the 500 draws |
| `draws SD` | Posterior standard deviation from the draws |

The first two should agree to within Monte Carlo error — that is the
package's own correctness check, and a useful one to glance at. Here the
average subject reacts about 298 ms at mid-study and slows by roughly 10 ms
per additional day of deprivation.

# Reading more

`summary()` gives the full report: prior versus reference values, posterior
percentiles, tail probabilities, and a per-group section.

```{r summary}
summary(fit)
```

The `Pr(Prior_tail)` column is a directional tail probability: how far into
the prior's tail the posterior has moved. A value near 0.5 means the data
did not move that coefficient much; the near-zero value on `Days_c` says
the sleep-deprivation effect is strongly supported relative to the prior,
which centered it at zero.

# Standard accessors

The usual `lme4` extractors work, though `fixef()` and `ranef()` are `lme4`
generics — load **lme4** or call them namespace-qualified.

```{r accessors}
head(lme4::ranef(fit)$Subject)   # per-subject coefficients, chain means
```

`fixef()` has three modes, and the default is not the Bayesian one:

```{r accessors-hyper}
lme4::fixef(fit)                     # "mer": the reference lmer estimates
lme4::fixef(fit, type = "hyper")     # Block 2 posterior mode
lme4::fixef(fit, type = "hyper.mean")# Block 2 posterior mean from the draws
```

The default returns the reference `lmer()` fit's fixed effects so that
`fixef()` behaves like it does on a `merMod` object. For posterior
quantities ask for `"hyper"` or `"hyper.mean"`, or read
`fit$popef.mode` and `fit$popef.means` directly.

Note a naming subtlety specific to this package. What `ranef()` returns is
the *full* per-subject coefficient vector, not `lme4`'s mean-zero
deviation. Subject 308's intercept of roughly 339 is their actual expected
reaction time, not an offset from the population mean of 298. Chapter 02
explains the parameterization, and Chapter 05 catalogues every accessor.

# Comparing with lme4

```{r lmer-compare}
m_lmer <- lme4::lmer(form, data = dat, REML = TRUE)
round(
  cbind(
    lmer  = lme4::fixef(m_lmer),
    lmerb = unlist(fit$popef.means)
  ),
  3
)
```

With a prior weight of one percent, the two agree closely — as they should.
The Bayesian fit adds full posterior uncertainty rather than a point
estimate with a standard error, and it does so without asymptotic
approximation. Chapter 13 works through where the two approaches diverge
and why.

# What to read next

- **Chapter 02** — why `||`, and how the parameterization maps to `lme4`.
- **Chapter 03** — the prior workflow in detail.
- **Chapter 04** — a real analysis with group-level covariates.

# See also

- `?lmerb`, `?model_setup`, `?Prior_Setup_GLMM`
- `vignette("Chapter-B00", package = "lmebayesCore")` — the engine roadmap.
