---
title: "Chapter 08: Observation dispersion"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 08: Observation dispersion}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

```{r load-stored}
ch08 <- readRDS(system.file(
  "extdata", "Chapter08_group_dispersion.rds", package = "lmebayes"
))
```

# Chapter 08: Observation dispersion

Chapter 07 relaxed the assumption that the *between-group* variances
\(\tau^2_k\) are known. This chapter does the same for the *within-group*
variance \(\sigma^2\) — the residual scatter of individual observations
around their group's fitted line.

Three levels of ambition:

| Specification | \(\sigma^2\) | Argument |
|---|---|---|
| Fixed scalar | One number, held at the `lmer` estimate | `dispersion_ranef = ps$group.dispersion` |
| Sampled scalar | One number with a posterior | `dispersion_ranef = dGamma(...)` |
| Per-group | One \(\sigma^2_j\) per group, each sampled | `dispformula = ~group`, `dGamma_list()` |

The per-group case is the interesting one, and it is what this chapter
demonstrates. Because it runs an accept-reject dispersion draw for every
group on every sweep, the fit takes about a minute, so it is precomputed by
`data-raw/make_vignette_outputs.R` and loaded here.

# The default: one fixed variance

```{r default}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)
form <- Reaction ~ Days_c + (1 + Days_c || Subject)

ps <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01)
ps$group.dispersion
```

One number, `r round(ch08$scalar_dispersion, 1)`, applied to all 180
observations. It assumes every subject's reaction times scatter by the same
amount around their own line — which, for a sleep-deprivation study, is
implausible on its face. Some people get erratic when tired and some do
not, and that heterogeneity *is* a finding, not a nuisance.

# Per-group dispersion

Ask `Prior_Setup_GLMM()` for a per-group calibration with `dispformula`:

```{r prior-group, eval = FALSE}
ps_group <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01,
                             dispformula = ~Subject)
dl <- dGamma_list(ps_group)
length(dl)
dl[[1]]
```

```{r prior-group-out, echo = FALSE}
cat(length(ch08$dGamma_list_names), "\n")
cat(ch08$dGamma_list_first, sep = "\n")
```

`dGamma_list()` returns one `dGamma` prior per group. Each carries a
`shape` and `rate` for a Gamma prior on the *precision* \(1/\sigma^2_j\)
(hence `Inv_Dispersion = TRUE`), plus a truncation window `disp_lower` /
`disp_upper` calibrated from that group's own data.

The windows differ across groups by more than an order of magnitude,
because they are quantiles of each group's limiting posterior and the
subjects genuinely differ. Calling `dGamma_list()` on a `dispformula = ~1`
setup is an error rather than a silent fallback:

```{r dgamma-error, error = TRUE}
dGamma_list(ps)
```

# Fitting

```{r fit-code, eval = FALSE}
set.seed(2024)
fit <- lmerb(
  form, data = dat,
  pfamily_list     = pfamily_list(ps_group),
  dispersion_ranef = dl,
  dispformula      = ~Subject,
  n                = 200L,
  progbar          = FALSE
)
fit
```

```{r fit-out, echo = FALSE}
cat(ch08$fit_print, sep = "\n")
```

The engine is `two-block Gibbs` with `m_convergence` = `r ch08$m_convergence`.
Sampling \(\sigma^2\) breaks the joint normality that made the earlier fits
exact, so sweeps are required — the same mechanism as Chapter 07, applied
at the observation level instead of the group level.

Note the `Residual` line now reads `NA`: there is no single residual
standard deviation to report, because there are 18 of them.

# Reading the dispersion

`summary_sigma2()` exists for exactly this case. On a scalar-dispersion fit
it stops with a message; here it produces a full report.

```{r sigma2-code, eval = FALSE}
summary_sigma2(fit)
```

```{r sigma2-out, echo = FALSE}
cat(ch08$sigma2_summary, sep = "\n")
```

Three blocks, one per question.

**Prior and reference** lists, per subject, the Gamma prior in use, the
prior mean, the truncation window, the number of observations, and the
pooled `lmer` residual variance for comparison. `n_data = 10` everywhere —
ten observations is very little information about a variance, which is why
the prior weight matters here more than anywhere else in the package.

**Overview** gives the posterior mean and standard deviation of each
\(\sigma^2_j\), Monte Carlo error, and `Cand/draw`. That last column is
accept-reject efficiency: values in the tens or hundreds mean the envelope
is loose for that group, usually because its posterior sits near a
truncation boundary.

**Percentiles** shows the spread directly, and the spread is the finding.
The posterior means run from roughly 80 to 3600 — subject 308's reaction
times scatter more than an order of magnitude more than subject 309's. A
single pooled \(\sigma^2\) of `r round(ch08$scalar_dispersion, 0)` splits
the difference and describes neither.

That heterogeneity also propagates: groups with large \(\sigma^2_j\) carry
less weight in the population estimate and get shrunk harder toward it,
which is the statistically correct behaviour and is not something a pooled
model can express.

# Sampling a single scalar

Between the two extremes there is a middle option: one \(\sigma^2\) shared
by all groups, but sampled rather than fixed. Pass a single `dGamma()`
object instead of a list:

```{r scalar-gamma, eval = FALSE}
fit_scalar_sampled <- lmerb(
  form, data = dat,
  pfamily_list     = pfamily_list(ps),
  dispersion_ranef = dGamma(shape = 2, rate = 2 * ps$group.dispersion,
                            Inv_Dispersion = TRUE),
  n = 200L, progbar = FALSE
)
```

This is the usual Bayesian treatment of a residual variance, and it is what
you want when you believe the groups share a scatter but do not want to
pretend you know its value. It costs sweeps, like any non-Gaussian target,
but far less than the per-group version.

# Choosing

| Situation | Use |
|---|---|
| Many observations per group, homogeneous scatter | Fixed scalar |
| Few groups, homogeneous scatter, honest intervals wanted | Sampled scalar |
| Groups visibly differ in residual spread | `dispformula = ~group` with `dGamma_list()` |

Diagnosing which you are in is easy before fitting: compute the residual
variance per group from an `lmer` fit and look at the ratio of largest to
smallest. Anything beyond a factor of three or four is worth modeling.

```{r diagnose}
m <- lme4::lmer(form, data = dat, REML = TRUE)
v <- tapply(residuals(m), dat$Subject, var)
round(c(min = min(v), median = median(v), max = max(v),
        ratio = max(v) / min(v)), 1)
```

A ratio of that size is not something a single pooled variance can absorb.

# See also

- **Chapter 07** — the same relaxation applied to \(\tau^2_k\).
- `vignette("Chapter-B07", package = "lmebayesCore")` — the per-group
  dispersion algebra and the marginal precision blocks.
- `?dGamma_list`, `?summary_sigma2`
