---
title: "Chapter 12: Row-block models"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 12: Row-block models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

# Chapter 12: Row-block models

`lmbBlock()` and `glmbBlock()` sit one level below the mixed-model
interface. They partition the rows of a dataset into blocks and fit an
**independent** Bayesian regression within each block — no hierarchy, no
shared population level, no shrinkage between blocks.

That makes them the "no pooling" corner of Chapter 02, done properly: each
block gets a full posterior rather than a point estimate, and the fits are
independent by construction rather than by accident.

# When you want this

**As a baseline.** Fitting the same data with and without a hierarchy is
the clearest way to see what partial pooling actually did.

**When the blocks are not exchangeable.** The mixed model assumes group
coefficients are draws from a common population. If your blocks are
different experiments, different instruments, or different time periods
with no reason to share information, that assumption is wrong and pooling
them is worse than not.

**As a conceptual sibling of Block 1.** A Block 1 update in `lmerb()` is
also a set of independent per-group regressions — the difference is that
there the per-group priors come from the population level and are refreshed
every sweep, whereas here they are fixed once. The two are not the same
code path, though: `lmbBlock()` fits each block with `glmbayes::lmb()`,
while Block 1 uses the compiled block engines in **lmebayesCore**.
`vignette("Chapter-B09", package = "lmebayesCore")` traces both.

# A worked example

```{r data}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)

## Three subjects, to keep the output readable.
sub <- droplevels(subset(dat, Subject %in% c("308", "309", "310")))
table(sub$Subject)
```

Priors come from `Prior_SetupGroup()`, the row-block counterpart of
`Prior_Setup_GLMM()`. Note the argument is `group`:

```{r prior}
psb <- Prior_SetupGroup(Reaction ~ Days_c, group = sub$Subject,
                        data = sub, pwt = 0.05)
psb
```

The object is a list of ordinary `Prior_Setup()` results, one per block.
Each block is calibrated against **its own** data — there is no borrowing,
which is exactly the point.

```{r pfamily}
pfb <- pfamily_list(psb)
length(pfb)
names(pfb)
```

```{r fit, results = "hide"}
set.seed(1)
fitb <- lmbBlock(Reaction ~ Days_c, block = sub$Subject, data = sub,
                 pfamily_list = pfb, n = 200L)
```

```{r fit-class}
class(fitb)
names(fitb)
```

The result is a `blmb` object: a named list of `lmb` fits, one per block,
with `call`, `formula`, `block`, `group_info`, `k`, `p`, `pred_names`, and
`pfamily_lists` attached as attributes. Each element is an ordinary
**glmbayes** `lmb` fit — the block wrapper partitions rows and loops, and
`lmb()` does the sampling — so everything you know about `lmb()` applies:

```{r one-block}
summary(fitb[["308"]])
```

# Comparing blocks

Because the elements are ordinary fits, comparison is a loop:

```{r compare}
tab <- t(vapply(fitb, function(f) colMeans(f$coefficients), numeric(2)))
round(tab, 3)
```

```{r compare-sd}
round(t(vapply(fitb, function(f) apply(f$coefficients, 2, sd), numeric(2))), 3)
```

# Versus the hierarchical fit

The same three subjects, fitted hierarchically:

```{r hier, results = "hide"}
form <- Reaction ~ Days_c + (1 + Days_c || Subject)
ps <- Prior_Setup_GLMM(form, data = sub, pop.pwt = 0.05)
set.seed(1)
fith <- lmerb(form, data = sub,
              pfamily_list     = pfamily_list(ps),
              dispersion_ranef = ps$group.dispersion,
              n = 200L, progbar = FALSE)
```

```{r hier-compare}
hier <- as.matrix(lme4::ranef(fith)$Subject)
round(cbind(
  block_intercept = tab[, "(Intercept)"],
  hier_intercept  = hier[rownames(tab), "(Intercept)"],
  block_slope     = tab[, "Days_c"],
  hier_slope      = hier[rownames(tab), "Days_c"]
), 3)
```

The hierarchical estimates sit between the block estimates and their common
average — that is shrinkage, visible directly. With only three subjects and
ten observations each, the pull is substantial. It would be weaker with
more data per subject and stronger with less, following the ratio in
Chapter 02.

Which is right depends on a question the data cannot answer: are these
three subjects draws from a population you care about? If yes, the
hierarchical estimates are better — they borrow strength. If they are three
unrelated experiments, the block estimates are the honest ones.

# Generalized responses

`glmbBlock()` is the same idea with a `family` argument:

```{r glmb-block, eval = FALSE}
psg <- Prior_SetupGroup(y ~ x, group = dat$block, data = dat,
                        family = poisson(), pwt = 0.05)
fitg <- glmbBlock(y ~ x, block = dat$block, data = dat,
                  family = poisson(),
                  pfamily_list = pfamily_list(psg), n = 200L)
```

Each block runs `glmbayes::glmb()` independently, so the draws are iid
within block and independent across blocks. There are no Gibbs sweeps
anywhere: without a shared population level there is nothing to alternate
between, and consequently no `m_convergence`, no `tv_tol`, and no pilot
stage. Convergence is not a question that arises.

# What you give up

| | `lmbBlock()` | `lmerb()` |
|---|---|---|
| Blocks share information | No | Yes |
| Population-level parameters | None | Estimated |
| Small / noisy blocks | Estimated on their own | Shrunk toward the population |
| Group-level covariates | Not expressible | Level-2 design |
| Sampling | iid within block | iid draws via certified sweeps |
| Blocks with rank-deficient designs | Fail individually | Excluded, others unaffected |

The most consequential row is the third. A block with two observations gets
a posterior dominated by its prior, and that prior was calibrated from
those same two observations. The hierarchical model handles this case
properly by letting the other blocks inform it.

# See also

- **Chapter 02** — pooling, and why shrinkage looks the way it does.
- `vignette("Chapter-B09", package = "lmebayesCore")` — the row-block
  engines, and how Block 1 reuses them.
- `vignette("Chapter-03", package = "glmbayes")` — `lmb()`, the per-block
  sampler.
- `?lmbBlock`, `?glmbBlock`, `?Prior_SetupGroup`
