---
title: "Chapter 06: Convergence and tv_tol"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 06: Convergence and tv_tol}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

```{r setup}
library(lmebayes)
```

# Chapter 06: Convergence and `tv_tol`

Every fit so far has printed `exact iid` and reported `m_convergence = 1`.
This chapter is about what happens when the sampler actually has to work,
and about the claim that makes this package unusual: **you do not diagnose
convergence, you specify it**.

# The contract

`tv_tol` is a total-variation tolerance. Setting `tv_tol = 0.01` means:
*each stored draw is within 0.01 in total variation of the exact joint
posterior*. The sampler computes how many inner Gibbs sweeps that requires
and runs exactly that many, reporting the number as `m_convergence`.

That is a different contract from the usual MCMC workflow. There is no
trace plot to eyeball, no R-hat to compare against 1.01, no effective
sample size. Those tools exist because the number of burn-in iterations is
unknown; here it is computed, from the model and the data, before sampling
starts.

The machinery is @Nygren2020: for a two-block Gibbs sampler on a Gaussian
target, the total-variation distance after \(l\) sweeps has a closed-form
bound in the eigenvalues of a rate matrix. Inverting that bound gives the
sweep count. `vignette("Chapter-B02", package = "lmebayesCore")` derives
the bound and `vignette("Chapter-B03", package = "lmebayesCore")` shows how
it is inverted.

# When sweeps are needed at all

```{r data}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)
form <- Reaction ~ Days_c + (1 + Days_c || Subject)
ps <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01)
```

With `sigma^2` fixed and every `tau^2_k` known, the joint posterior over
group and population coefficients is exactly multivariate normal, and
`sim_method = "DEFAULT"` samples it directly:

```{r default, results = "hide"}
set.seed(1)
fit_iid <- lmerb(form, data = dat,
                 pfamily_list     = pfamily_list(ps),
                 dispersion_ranef = ps$group.dispersion,
                 n = 200L, progbar = FALSE)
```

```{r default-show}
c(engine = fit_iid$sim_method_used, m = fit_iid$m_convergence)
```

Gibbs sweeps become necessary when the target stops being jointly normal:
sampled variance components (Chapter 07), sampled observation dispersion
(Chapter 08), or a non-Gaussian likelihood (Chapters 09–10). You can also
force the Gibbs path on a target where the exact answer is known, which is
how the sampler is validated:

```{r gibbs, results = "hide"}
set.seed(1)
fit_gibbs <- lmerb(form, data = dat,
                   pfamily_list     = pfamily_list(ps),
                   dispersion_ranef = ps$group.dispersion,
                   n = 200L, progbar = FALSE,
                   sim_method = "TWO_BLOCK_GIBBS")
```

```{r gibbs-show}
c(engine = fit_gibbs$sim_method_used, m = fit_gibbs$m_convergence)
```

Three sweeps per stored draw. Both fits target the same posterior, so their
answers should agree:

```{r compare}
round(rbind(
  iid   = unlist(fit_iid$popef.means),
  gibbs = unlist(fit_gibbs$popef.means),
  exact = unlist(fit_iid$popef.mode)
), 3)
```

# What `m_convergence` responds to

Tightening the tolerance costs sweeps, but only logarithmically:

```{r tv-tol, results = "hide"}
tols <- c(0.1, 0.01, 1e-4, 1e-6)
ms <- vapply(tols, function(tol) {
  f <- lmerb(form, data = dat,
             pfamily_list     = pfamily_list(ps),
             dispersion_ranef = ps$group.dispersion,
             n = 20L, progbar = FALSE,
             sim_method = "TWO_BLOCK_GIBBS", tv_tol = tol)
  f$m_convergence
}, integer(1))
```

```{r tv-tol-table}
data.frame(tv_tol = tols, m_convergence = ms)
```

Four orders of magnitude of extra accuracy for two extra sweeps. That is
what makes the "certify every draw" contract affordable, and it has a
practical consequence: **there is rarely a reason to loosen `tv_tol`**. If
you want the whole `n`-draw sample certified at level \(\alpha\) rather
than each draw individually, set `tv_tol = alpha / n` — a union bound — and
pay a handful of sweeps for it.

The other driver is the model itself. Sweeps go up when group coefficients
and population coefficients are tightly coupled, which happens when groups
are small relative to the random-effect variance. This is the same quantity
that governs shrinkage (Chapter 02): heavily shrunk models mix slowly.

# The sweep history

When the Gibbs path runs, the fit stores cross-chain summaries after every
inner sweep.

```{r sweep-history}
names(fit_gibbs$sweep_history)
fit_gibbs$sweep_history$main
```

Read this as a convergence check you did not have to design. At sweep `m`,
the sampler has 200 independent chains all at sweep `m`; the `mean` and
`sd` columns are the cross-chain mean and standard deviation. Because the
chains are independent, that is a genuine snapshot of the sampler's
distribution after `m` sweeps — not a running average along one correlated
chain.

The `mode` row is the exact answer. Every chain starts there, which is why
the means barely move: starting at the posterior mean sets the mean-error
term of the convergence bound to exactly zero, leaving only the variance
term to be certified. That design choice is what keeps `m_convergence` in
the single digits.

The `pilot` element is populated when a pilot stage runs; see Chapter 07.

# Convergence plots

Two plotting helpers turn the history into pictures.

```{r plot-mean, fig.alt = "Standardized cross-chain mean by inner sweep"}
plot_mean_convergence(fit_gibbs$sweep_history$main)
```

This plots \((\text{mean}_m - \text{mean}_{\text{final}}) / \text{SD}\)
against sweep. Values hugging zero mean the chain distribution's *center*
was already correct — expected here, since chains start at the exact mean.
A visible drift would signal that the start point is not where the theory
assumes it is.

```{r plot-var, fig.alt = "Ratio of cross-chain variance to final variance by inner sweep"}
plot_var_convergence(fit_gibbs$sweep_history$main)
```

This plots \(\text{Var}_m / \text{Var}_{\text{final}}\). Chains that all
start at the same point have zero spread at sweep 0 and *grow* into the
posterior variance, so the interesting direction is upward approach to 1.
This is the quantity the Theorem 3 bound actually certifies, which is why
it is the more informative of the two plots for this sampler.

Both functions take a sweep-history object, not a fit — pass
`fit$sweep_history$main` or `$pilot`. Details in
`vignette("Chapter-B10", package = "lmebayesCore")`.

# Verifying the certificate empirically

The sampler's claim is falsifiable on this model, because the exact
posterior is known. Compare the Gibbs draws against the analytic mean:

```{r verify}
n_draws <- nrow(fit_gibbs$popef[["Days_c"]])
draws   <- fit_gibbs$popef[["Days_c"]][, 1]
exact   <- fit_gibbs$popef.mode[["Days_c"]][[1]]
z <- (mean(draws) - exact) / (sd(draws) / sqrt(n_draws))
round(c(draws_mean = mean(draws), exact_mean = exact, z = z), 4)
```

A \(z\) of moderate size is what independent draws from the right
distribution look like. This check is cheap and worth running whenever you
change sampler settings — it is the same test the package's own test suite
applies.

# What convergence does *not* mean

`m_convergence` certifies distance to the posterior **of the model you
specified**. It says nothing about whether that model is a good description
of the data. Prior sensitivity (Chapter 03), the uncorrelated-random-effects
restriction (Chapter 02), and comparison against `lme4` (Chapter 13) are
separate questions, and no amount of sampling accuracy answers them.

One warning you may see is `lambda* > 0.95`: the chain is still
geometrically ergodic, but coupling is strong enough that the model
deserves a second look. Usual causes are very small groups or a variance
component being driven toward zero.

# See also

- **Chapter 07** — the pilot stage, which is where the mean-error term of
  the bound stops being zero.
- `vignette("Chapter-B02", package = "lmebayesCore")` — the ergodicity
  theory.
- `vignette("Chapter-B03", package = "lmebayesCore")` — from the rate to a
  sweep count.
- `?plot_mean_convergence`, `?plot_var_convergence`
