---
title: "Chapter 10: Binomial GLMMs"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 10: Binomial GLMMs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

```{r load-stored}
ch10 <- readRDS(system.file(
  "extdata", "Chapter10_book_banning_binom.rds", package = "lmebayes"
))
```

# Chapter 10: Binomial GLMMs

Binary outcomes with group structure. The data is `book_banning` from
**bayesrules** [@JohnsonLauEtAl2022]: challenges to books in US libraries,
with an indicator for whether the book was removed, grouped by state.

The mechanics are the same as Chapter 09 — `glmerb()` with
`family = binomial()` — so this chapter concentrates on what is different
about binary data: separation, estimability, and why the sweep count is
larger.

# Setting up

```{r data-code, eval = FALSE}
data(book_banning, package = "bayesrules")
bb <- book_banning[, c("state", "removed", "violent")]
bb <- bb[complete.cases(bb), ]
bb$removed_i <- as.integer(bb$removed == 1L | bb$removed == "1")
bb$violent_i <- as.integer(
  bb$violent == TRUE | bb$violent == 1L | bb$violent == "TRUE"
)
bb$state <- factor(bb$state)

## Twelve largest states, to keep the stored fit small.
keep <- names(sort(table(bb$state), decreasing = TRUE))[seq_len(12L)]
bb <- droplevels(subset(bb, state %in% keep))

form <- removed_i ~ violent_i + (1 + violent_i || state)
model_setup(form, data = bb, family = binomial())
```

```{r design-out, echo = FALSE}
cat(ch10$design, sep = "\n")
```

Note the explicit conversion to integers. Logical and factor predictors get
expanded by `model.matrix()` in ways that can produce correlated
random-effect terms, which this package rejects; converting to 0/1 integers
keeps the `||` structure intact. If you hit
`Correlated random effects are not supported`, check for logical or factor
variables inside the random-effects term first.

# Full rank is not the same as estimable

The rank report shows something that did not appear in earlier chapters:

```
Full-rank D_j: 12 of 12 groups
Full-rank & estimable: 11 of 12 full-rank (11 of 12 total state)
  full-rank but not estimable: MI
```

Michigan's design matrix has full column rank — it contains both violent
and non-violent challenges — but the model is still not estimable there.
That is **separation**: within Michigan, the outcome is perfectly predicted
by `violent_i`, so the maximum-likelihood coefficient runs off to infinity.
The likelihood is log-concave but has no interior maximum in that
direction.

Separation is a binary-data problem specifically; it cannot happen with
Gaussian or (in this form) Poisson responses. `model_setup()` runs a
separate estimability check for exactly this reason, and it is the reason
the check is reported on its own line rather than folded into the rank
test.

Michigan is not thrown away. It keeps its coefficients, drawn from the
prior in the direction the likelihood cannot pin down — which is exactly
what a proper prior is for. What it loses is standing in the *level-2*
checks: an infinite-in-the-limit estimate carries no usable information
about the population parameters, so `model_setup()` restricts the
across-group rank condition to the 11 estimable states, and
`Prior_Setup_GLMM()` calibrates from those. Details in
`vignette("Chapter-B08", package = "lmebayesCore")`.

# Fitting

```{r fit-code, eval = FALSE}
ps <- Prior_Setup_GLMM(form, data = bb, family = binomial(), pop.pwt = 0.01)

set.seed(2024)
fit <- glmerb(
  form,
  data         = bb,
  family       = binomial(),
  pfamily_list = pfamily_list(ps),
  n            = 200L,
  progbar      = FALSE
)
fit
```

```{r fit-out, echo = FALSE}
cat(ch10$fit_print, sep = "\n")
```

# Reading it on the probability scale

Coefficients are log-odds. Two useful transformations:

```{r interpret}
pm <- unlist(ch10$popef_means)
round(rbind(log_odds = pm, odds_ratio = exp(pm)), 3)

plogis(pm[[1]])            # P(removed), non-violent
plogis(pm[[1]] + pm[[2]])  # P(removed), violent
```

In an average state, a challenged non-violent book has roughly a
`r round(100 * stats::plogis(unlist(ch10$popef_means)[[1]]))` percent chance
of removal, rising modestly for books challenged as violent. The posterior SD
on the `violent_i` effect is larger than the effect itself, so this is weak
evidence of any population-level difference.

Note that `plogis()` of the *mean* log-odds is not the mean probability —
the transformation is nonlinear. For a proper posterior on the probability
scale, transform the draws and then summarize:

```{r prob-draws, eval = FALSE}
p_draws <- plogis(fit$popef[["(Intercept)"]][, 1])
quantile(p_draws, c(0.025, 0.5, 0.975))
```

# Per-state effects

```{r ranef-out, echo = FALSE}
cat(ch10$ranef_head, sep = "\n")
```

These are log-odds intercepts per state, and they are full coefficients
rather than deviations (Chapter 02). Oregon at about -2.87 and Virginia at
about -0.09 correspond to removal probabilities of roughly 5 percent and 48
percent — a large practical difference that a state-level fixed effect
would have estimated far more noisily, and that complete pooling would have
erased entirely.

# Why the sweep count is higher

```{r convergence}
c(m_convergence = ch10$m_convergence)
```

`r ch10$m_convergence` sweeps, against 3 for the Poisson model of Chapter
09 and 1 for the Gaussian fits.

The reason is information per observation. A binary response carries at
most about 0.25 units of Fisher information per observation, and much less
when the probability is near 0 or 1 — which it is here, since most
challenges do not result in removal. With only 12 states and weak
per-observation information, the group coefficients and the population
coefficients are tightly coupled, the rate \(\lambda^*\) is close to one,
and more sweeps are needed.

This is the same relationship as in Chapter 02: **models that shrink hard
mix slowly**. Binary outcomes with few groups and rare events sit at the
difficult end of that spectrum, and the sweep count reports it honestly
rather than hiding it.

# The full summary

```{r summary-code, eval = FALSE}
summary(fit)
```

```{r summary-out, echo = FALSE}
cat(ch10$fit_summary, sep = "\n")
```

Both random-effect standard deviations are near 0.85 on the log-odds scale,
which is substantial: a one-standard-deviation state differs from average
by a factor of about `r round(exp(0.86), 1)` in odds. As in Chapter 09, the
between-group variation exceeds the population-level effect, which is the
signature of a model that needed random slopes.

# Practical notes for binary data

- **Convert to integers** before fitting; logical and factor variables in
  the random-effects term commonly trigger the correlated-effects error.
- **Check the estimability line**, not just the rank line. Separation is
  silent otherwise.
- **Expect more sweeps.** If `m_convergence` climbs into the tens, the
  cause is usually too few groups or too rare an outcome, not a sampler
  problem.
- **Watch for a `lambda* > 0.95` warning.** The chain is still valid, but
  the model is close to the boundary where the group and population levels
  cannot be separated.

# See also

- **Chapter 09** — the Poisson case and the shared GLMM machinery.
- **Chapter 11** — mode-only fits, which avoid the sampling cost entirely.
- `vignette("Chapter-B08", package = "lmebayesCore")` — the rank and
  estimability conditions.
- `?glmerb`, `?model_setup`
