---
title: "Chapter 00: Introduction and roadmap"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 00: Introduction and roadmap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

# Chapter 00: Introduction and roadmap

**lmebayes** fits Bayesian mixed-effects models — linear with `lmerb()`,
generalized with `glmerb()` — using a two-block Gibbs sampler whose burn-in
length is *derived from theory rather than guessed at*. You give the
sampler a total-variation tolerance; it computes how many inner sweeps are
needed to meet that tolerance and runs exactly that many.

That single design decision shapes everything else in this series. There
is no trace-plot ritual, no R-hat, no effective sample size: the `n` draws
a fit returns are mutually independent, each within `tv_tol` of the exact
joint posterior. The convergence theory that makes this possible is due to
@Nygren2020 and is derived in the **lmebayesCore** appendices.

## What this package covers

`lmerb()` and `glmerb()` accept `lme4`-style formulas and return objects
with familiar accessors — `fixef()`, `ranef()`, `coef()`, `summary()`.
The model class is deliberately narrower than `lme4`'s:

| Supported | Not supported |
|---|---|
| One grouping factor | Crossed or nested grouping factors |
| Uncorrelated random-effect terms (`||`) | Correlated terms (`|`) |
| Gaussian, Poisson, binomial, Gamma | Non-log-concave likelihoods |
| Group-level covariates on any random coefficient | — |
| Fixed or sampled variance components | — |
| Scalar, per-group, or modeled observation dispersion | — |

The restrictions are what buy the convergence certificate. A single
grouping factor with uncorrelated terms makes the Block 2 update factorize
across coefficients, and log-concavity is what the envelope accept-reject
sampler of @Nygren2006 requires.

## The five parts

### Part 1 — Foundations

- **Chapter 00 — Introduction and roadmap.** This chapter.
- **Chapter 01 — Getting started.** Your first `lmerb()` fit on
  `lme4::sleepstudy`: install, fit, read the output, compare to `lmer()`.
- **Chapter 02 — From `lm`/`glm` to mixed models.** Partial pooling, the
  centered parameterization this package uses, and how it maps onto
  `lme4`'s fixed-plus-deviation form.

### Part 2 — Specifying a model

- **Chapter 03 — The prior workflow.** `model_setup()` →
  `Prior_Setup_GLMM()` → `pfamily_list()`, and what every field means.
- **Chapter 04 — Linear mixed models with `lmerb()`.** A full analysis on
  `bayesrules::big_word_club`: random intercept and slopes, group-level
  covariates, cross-level moderation.
- **Chapter 05 — Reading the output.** `print()`, `summary()`, `fixef()`,
  `ranef()`, `coef()`, `summary_sigma2()`, and the raw draw containers.

### Part 3 — Convergence and variance components

- **Chapter 06 — Convergence and `tv_tol`.** What `m_convergence` means,
  how it is derived, and how to check it empirically with
  `plot_mean_convergence()` and `plot_var_convergence()`.
- **Chapter 07 — Estimated variance components and the pilot stage.**
  Moving `tau^2` from fixed to sampled with
  `dIndependent_Normal_Gamma()`.
- **Chapter 08 — Observation dispersion.** Scalar `sigma^2`, per-group
  dispersion with `dGamma_list()`, and `dispformula`.

### Part 4 — Generalized linear mixed models

- **Chapter 09 — Poisson GLMMs.** Counts with a log link, on
  `bayesrules::airbnb`.
- **Chapter 10 — Binomial GLMMs.** `bayesrules::book_banning`, logistic
  random slopes.
- **Chapter 11 — Point estimates without simulation.** `simulate = FALSE`,
  iterated conditional modes, and when the mode equals the mean.

### Part 5 — Extensions and comparisons

- **Chapter 12 — Row-block models.** `lmbBlock()` / `glmbBlock()`, where
  the row partition is user-specified rather than a grouping factor.
- **Chapter 13 — Comparing with `lme4` and `glmmTMB`.** Where results
  should agree, where they should not, and why.

## Background reading in the wider ecosystem

**lmebayes** sits on top of two other packages, and each carries its own
vignette series.

| Series | Package | What it covers |
|---|---|---|
| `Chapter-00` … `Chapter-13` | **lmebayes** | This curriculum |
| `Chapter-B00` … `Chapter-B10` | **lmebayesCore** | Engine internals, ergodicity theory, derivations |
| `Chapter-00` … `Chapter-18`, `A01`–`A12` | **glmbayes** | The iid GLM stack everything is built from |

If mixed models are new to you, read Chapters 01–04 here and stop. If you
want to know *why* a fit reports the `m_convergence` it does, read
`Chapter-B02` and `Chapter-B03` in **lmebayesCore**:

```r
vignette("Chapter-B02", package = "lmebayesCore")
vignette("Chapter-B03", package = "lmebayesCore")
```

If the accept-reject machinery behind every non-Gaussian draw is
unfamiliar, read **glmbayes** `Chapter-A05` (likelihood subgradient
densities) and `Chapter-A08` (envelope functions):

- <https://knygren.r-universe.dev/articles/glmbayes/Chapter-A05.html>
- <https://knygren.r-universe.dev/articles/glmbayes/Chapter-A08.html>

## A first look

The complete workflow is four calls. Formula, design, priors, fit.

```{r first-look}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)

form <- Reaction ~ Days_c + (1 + Days_c || Subject)

ps <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01)

set.seed(1)
fit <- lmerb(
  form,
  data             = dat,
  pfamily_list     = pfamily_list(ps),
  dispersion_ranef = ps$group.dispersion,
  n                = 200L,
  progbar          = FALSE
)

fit
```

`Prior_Setup_GLMM()` calibrates priors from a reference `lmer()` fit —
`pop.pwt = 0.01` means "give the prior the weight of one percent of the
data". `pfamily_list()` turns that calibration into one prior object per
random-effect coefficient. `lmerb()` does the rest.

Note what the header reports: `200 draws, exact iid`. With the observation
variance and both variance components held fixed, the joint posterior is
exactly multivariate normal, so the sampler draws from it directly and no
Gibbs sweeps are needed at all. Chapter 06 covers what happens when they
are.

## Which chapter answers your question

| Question | Chapter |
|---|---|
| How do I fit my first model? | 01 |
| Why is `(1 + x \|\| g)` required instead of `(1 + x \| g)`? | 02 |
| What do all these prior fields mean? | 03 |
| How do I add a group-level covariate? | 04 |
| Where are the posterior draws stored? | 05 |
| What is `m_convergence` and can I trust it? | 06 |
| How do I put a prior on the random-effect variance? | 07 |
| My groups have different residual variances. | 08 |
| My response is a count / a proportion. | 09, 10 |
| I only want point estimates, no draws. | 11 |
| My "groups" are arbitrary row blocks, not a factor. | 12 |
| Why does this disagree with `lmer()`? | 13 |

## Companion textbook

Several chapters use datasets from @JohnsonLauEtAl2022 (*Bayes Rules!*),
available in the suggested package **bayesrules**. Chapter 04 uses
`big_word_club`, Chapter 09 uses `airbnb`, and Chapter 10 uses
`book_banning`. Those chapters gate their data-dependent chunks so the
vignettes still knit without **bayesrules** installed.

# See also

- **Chapter 01** — getting started.
- `?lmerb`, `?glmerb` — the function references.
- `vignette("Chapter-B00", package = "lmebayesCore")` — the engine's own
  roadmap.
