---
title: "Chapter 13: Comparison with lme4 and glmmTMB"
author: "Kjell Nygren"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
bibliography: REFERENCES.bib
reference-section-title: References
vignette: >
  %\VignetteIndexEntry{Chapter 13: Comparison with lme4 and glmmTMB}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lmebayes)
```

```{r load-stored}
ch13 <- readRDS(system.file(
  "extdata", "Chapter13_mer_comparison.rds", package = "lmebayes"
))
```

# Chapter 13: Comparison with `lme4` and `glmmTMB`

This chapter is about scope. **lmebayes** does a narrower set of models than
`lme4` [@Bates2015] or `glmmTMB` [@Brooks2017], and does something
different with them. Knowing where the line falls is the difference between
using it well and fighting it.

Because **glmmTMB** is only suggested, the comparison is precomputed by
`data-raw/make_vignette_outputs.R`.

# Same model, three packages

```{r code, eval = FALSE}
data(sleepstudy, package = "lme4")
dat <- sleepstudy
dat$Days_c <- dat$Days - mean(dat$Days)
form <- Reaction ~ Days_c + (1 + Days_c || Subject)

m_lmer <- lme4::lmer(form, data = dat, REML = TRUE)
m_tmb  <- glmmTMB::glmmTMB(form, data = dat, family = gaussian())

ps <- Prior_Setup_GLMM(form, data = dat, pop.pwt = 0.01)
set.seed(2024)
fit <- lmerb(form, data = dat,
             pfamily_list     = pfamily_list(ps),
             dispersion_ranef = ps$group.dispersion,
             n = 1000L, progbar = FALSE)
```

```{r compare-fixef, echo = FALSE}
knitr::kable(ch13$fixef_compare, digits = 4)
```

All three agree to two decimal places. With a prior weight of one percent
they should: the prior is nearly flat, so the posterior mean lands on the
maximum-likelihood estimate.

The agreement is worth stating precisely, because it is easy to
over-interpret. It shows that the Bayesian machinery reproduces the right
answer in the case where a right answer is agreed on. It does not show the
three approaches are interchangeable.

# What is actually different

`lmer()` and `glmmTMB()` report a point estimate and a standard error
derived from the curvature of the likelihood at the optimum. That standard
error is asymptotic: it is exact as the number of groups goes to infinity,
and approximate otherwise. With 18 subjects, "otherwise" is where you are.

**lmebayes** reports draws from the posterior. Uncertainty comes from the
spread of those draws, with no asymptotic step:

```{r lmerb-print, echo = FALSE}
cat(ch13$lmerb_print, sep = "\n")
```

For this model the two happen to agree closely, because the posterior is
Gaussian and the sample is large enough. Where they diverge is small
numbers of groups, variance components near zero, and non-Gaussian
likelihoods — precisely the situations where the asymptotic approximation
is weakest.

The second difference is the one that has no counterpart at all. When the
header says `exact iid`, the 1000 draws are independent samples from the
exact posterior. When it says `two-block Gibbs`, each draw is certified
within `tv_tol` in total variation. Neither `lme4` nor `glmmTMB` makes a
distributional claim of that kind, because they are not sampling. Compared
against other *Bayesian* packages the claim is still unusual: Stan and JAGS
give correlated draws with diagnostics that are heuristic
(@GelmanRubin1992), not certificates.

# The feature matrix

| | `lme4` | `glmmTMB` | **lmebayes** |
|---|---|---|---|
| Crossed random effects | Yes | Yes | **No** |
| Nested random effects | Yes | Yes | **No** |
| Correlated random effects (`\|`) | Yes | Yes | **No** — `\|\|` only |
| Fixed effect without matching random slope | Yes | Yes | **No** |
| Families | Many | Many + zero-inflation | Gaussian, Poisson, binomial, Gamma |
| Dispersion models | No | Yes (`dispformula`) | Yes, per group |
| Priors | None | None | Calibrated, required |
| Output | Point estimate + SE | Point estimate + SE | Posterior draws |
| Uncertainty | Asymptotic | Asymptotic | Exact, finite-sample |
| Convergence | Optimizer | Optimizer | Certified per draw |

The three "No" rows in the middle are the cost of the two-block structure.
Crossed and nested factors break the conditional independence that makes
Block 1 factorize; correlated random effects break the coefficient-wise
factorization of Block 2; and the fixed-effect rule follows from the
centered parameterization (Chapter 02).

# Variance components

```{r varcomp, echo = FALSE}
knitr::kable(ch13$varcomp, digits = 4)
```

`lmer()` gives one number per component. In Chapter 07 those same
components got full posteriors, with credible intervals that were
noticeably wide — 18 subjects is little information about how subjects
vary. A REML point estimate does not carry that uncertainty forward into
the fixed-effect standard errors; the Bayesian fit does, which is the main
practical reason to prefer it here.

# Choosing

**Use `lme4`** for exploration, for crossed or nested designs, and for any
model with correlated random effects. It is fast, well understood, and the
reference implementation.

**Use `glmmTMB`** for zero-inflation, for families outside the log-concave
class, and for complex dispersion models.

**Use lmebayes** when the model fits its class — one grouping factor,
uncorrelated random effects, a log-concave family — and you want exact
finite-sample uncertainty rather than an asymptotic approximation. It is
strongest exactly where the alternatives are weakest: few groups, small
groups, variance components near the boundary.

A reasonable workflow uses both. Fit in `lme4` first: it is fast, it tells
you whether the random-effect correlation you are about to discard is
large, and its estimates are what `Prior_Setup_GLMM()` calibrates against
anyway. Then refit in **lmebayes** for inference.

```{r workflow, eval = FALSE}
m <- lme4::lmer(Reaction ~ Days_c + (1 + Days_c | Subject), data = dat)
print(lme4::VarCorr(m), comp = "Std.Dev.")   # is the correlation large?
```

If the correlation is near zero, the `||` restriction costs nothing. If it
is large, that is a substantive finding and this package cannot represent
it.

# Interface differences to remember

```{r interface, eval = FALSE}
## lme4
m <- lmer(y ~ x + (1 + x | g), data = d)
fixef(m); ranef(m); coef(m)

## lmebayes
ps  <- Prior_Setup_GLMM(y ~ x + (1 + x || g), data = d, pop.pwt = 0.01)
fit <- lmerb(y ~ x + (1 + x || g), data = d,
             pfamily_list     = pfamily_list(ps),
             dispersion_ranef = ps$group.dispersion,
             n = 1000)
fixef(fit, type = "hyper.mean")   # fixef(fit) returns the lmer reference
ranef(fit)                        # full coefficients, not deviations
```

Three traps, all covered earlier: `ranef()` returns full coefficients
rather than mean-zero deviations (Chapter 02); `fixef()` defaults to the
reference `lmer` fit rather than the posterior (Chapter 05); and the prior
is a required argument, not an option (Chapter 03).

# See also

- **Chapter 02** — why the restrictions exist.
- **Chapter 07** — variance components with real uncertainty.
- `vignette("Chapter-B02", package = "lmebayesCore")` — the convergence
  theory behind the certified-draw claim.
