Yaml

---
title: "Chapter 09: Models for the Gamma Family"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Chapter 09: Models for the Gamma Family}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(glmbayes)
```

## 1. Introductory Discussion

Gamma regression models are used when the response variable is strictly positive and right‑skewed.  
Typical applications include:

- insurance claim severities  
- cost or expenditure data  
- reaction times and waiting times  
- positive continuous measurements with heteroskedasticity

In classical statistics, these models are fit using

glm(..., family = Gamma(link = ...)).

In glmbayes, the Bayesian analogue is

glmb(..., family = Gamma(link = ...), pfamily = dNormal(...))

or, when dispersion is unknown and modeled,

glmb(..., family = Gamma(link = ...), pfamily = dNormal_Gamma(...))

or related pfamilies.

This chapter:

1. reviews the Gamma GLM structure and link functions  
2. shows how to fit classical and Bayesian Gamma regression models  
3. uses a realistic insurance example (carinsca)  
4. demonstrates how to estimate dispersion using gamma.dispersion()  
5. compares classical and Bayesian summaries, including DIC


## 2. The Gamma Likelihood

Gamma GLMs assume a positive continuous response with variance proportional to the square of the mean:

Y ~ Gamma(shape = k, mean = mu)

Var(Y) = mu^2 * phi

where phi is the dispersion parameter.

The exponential‑family form is

f(y | theta, phi) = exp( (y * theta - b(theta)) / a(phi) + c(y, phi) )

with

theta = -1 / mu  
b(theta) = -log(-theta)  
V(mu) = mu^2.

This structure yields log‑concavity under the canonical inverse link and under the commonly used log link, which is important for the envelope‑based sampler in glmbayes.


## 3. Link Functions for Gamma GLMs

The link function relates the mean mu to the linear predictor eta = X * beta:

g(mu) = eta.

The Gamma family in base R and glmbayes supports:

| Link      | Formula           | Notes                                   |
|-----------|-------------------|-----------------------------------------|
| inverse   | eta = 1 / mu      | canonical link                          |
| identity  | eta = mu          | must ensure mu > 0                      |
| log       | eta = log(mu)     | most common; ensures positivity of mu   |

In practice, the log link is usually preferred because:

- it enforces mu > 0 automatically  
- it gives multiplicative interpretations for coefficients  
- it tends to be numerically stable  
- it preserves log‑concavity of the likelihood


## 4. Gamma Regression Example: Insurance Claims

We now consider a Gamma regression example based on the carinsca dataset.  
The goal is to model average claim cost as a function of rating variables Merit and Class, using claim counts as weights.

### 4.1 Data Setup

We begin by preparing the data and setting appropriate factor levels and contrasts.

    data(carinsca)
    carinsca$Merit <- ordered(carinsca$Merit)
    carinsca$Class <- factor(carinsca$Class)
    options(contrasts = c("contr.treatment", "contr.treatment"))

    Claims  <- carinsca$Claims
    Insured <- carinsca$Insured
    Merit   <- carinsca$Merit
    Class   <- carinsca$Class
    Cost    <- carinsca$Cost

The response Cost/Claims represents the **average claim cost** (severity) per claim.  
The weights Claims reflect the number of claims on which each average is based, improving efficiency and aligning with standard GLM practice for rate or average models.


### 4.2 Classical Gamma Regression Using glm()

We fit a classical Gamma GLM with log link:

    out <- glm(Cost/Claims ~ Merit + Class,
               family  = Gamma(link = "log"),
               weights = Claims,
               x       = TRUE)

    summary(out)

The summary shows:

- regression coefficients on the log mean‑cost scale  
- standard errors and z‑statistics  
- deviance and AIC  
- an estimated dispersion parameter (phi)

Under the log link, exp(beta_j) can be interpreted as a multiplicative factor on the expected cost ratio for a one‑unit change in the corresponding covariate (or a change in factor level).


## 5. Bayesian Gamma Regression with glmb()

To fit the Bayesian analogue, we proceed in two steps:

1. estimate dispersion phi from the classical model,  
2. construct a prior for the regression coefficients using Prior_Setup(),  
3. call glmb() with a dNormal prior that uses the fixed dispersion.

### 5.1 Estimating Dispersion via gamma.dispersion()

The gamma.dispersion() helper estimates the dispersion parameter from a fitted Gamma GLM:

    disp <- gamma.dispersion(out)

This value will be treated as known in our Bayesian model, so that the Bayesian and classical fits are directly comparable in terms of how they treat phi.


### 5.2 Prior Setup Using Prior_Setup()

Next we use Prior_Setup() to construct a Zellner‑type g‑prior for the coefficients, aligned with the Gamma log‑link model and weighting structure:

    ps <- Prior_Setup(Cost/Claims ~ Merit + Class,
                      family  = Gamma(link = "log"),
                      weights = Claims)

    mu <- ps$mu
    V  <- ps$Sigma

The output from ps (if printed) will include:

- prior means (intercept from a null model, effects centered at zero by default)  
- prior standard deviations (scaled using the pwt argument and the Fisher information)  
- any additional hyperparameters relevant for dispersion modeling

With the default pwt = 0.01, the prior is weakly informative, and the posterior estimates typically remain close to the MLEs.


### 5.3 Fitting the Bayesian Gamma Model

We now call glmb(), specifying the same model and likelihood as in the classical fit, but adding a dNormal prior family that fixes dispersion at the value estimated above:

    out3 <- glmb(Cost/Claims ~ Merit + Class,
                 family  = Gamma(link = "log"),
                 pfamily = dNormal(mu = mu, Sigma = V, dispersion = disp),
                 weights = Claims)

The arguments mirror the glm() call, with pfamily providing the prior:


- family = Gamma(link = "log") specifies the likelihood and link  
- pfamily = dNormal(...) specifies a multivariate normal prior for beta with known dispersion  
- weights = Claims carries over the claim counts

### 5.4 Summarizing and Comparing the Fits

Classical summary:

    summary(out)

Bayesian summary:

    summary(out3)

The Bayesian summary will report:

- posterior modes and posterior means for each coefficient  
- posterior standard deviations and Monte Carlo errors  
- tail probabilities relative to the prior mean  
- posterior percentiles (credible intervals)  
- effective number of parameters pD  
- expected residual deviance D  
- DIC

Because dispersion is fixed at the same value in both models, differences between out and out3 reflect primarily the influence of the prior on beta, not differences in phi.


## 6. Interpretation and Model Diagnostics

### 6.1 Coefficients

Under the log link, coefficient interpretations are multiplicative:

- exp(beta_j) is the multiplicative change in expected cost for a one‑unit change in the covariate (or for a change from baseline to a given factor level).

With a weak prior, posterior means in out3 will be close to the MLEs in out, but with slightly more regularization and smaller posterior standard deviations.

### 6.2 Dispersion

By fixing dispersion at disp from the classical model, the Bayesian and classical fits share the same assumed level of variability.  
This allows:

- direct comparison of log‑likelihood based quantities,  
- a clean assessment of how the prior affects coefficient estimates.

If desired, a more advanced model can estimate dispersion by switching to a pfamily such as dNormal_Gamma, but that is covered in more detail in the dispersion‑focused vignettes.


### 6.3 DIC and Model Fit

The Bayesian summary of out3 reports:

- the effective number of parameters pD,  
- the expected residual deviance D,  
- the DIC value.

These can be compared to AIC and deviance from the classical model, especially when considering alternative specifications (for example, reduced models, different covariate sets, or alternative link functions).


## 7. Concluding Discussion

Gamma regression provides a flexible framework for modeling positive, skewed response variables.  
In this chapter, we have:

- introduced the Gamma GLM and its link functions,  
- shown how to fit a classical Gamma regression using glm(),  
- demonstrated how to estimate dispersion with gamma.dispersion(),  
- used Prior_Setup() to construct a default Zellner‑type prior,  
- fit a Bayesian Gamma regression with glmb() using a dNormal prior with fixed dispersion,  
- and compared classical and Bayesian summaries in a realistic insurance setting.

In later vignettes, we extend these ideas to:

- Gamma models with unknown dispersion and hierarchical priors,  
- Gamma regression combined with other components in larger GLM systems,  
- and specialized accept‑reject schemes for dispersion parameters.
